\section{Riemann integrability of continuous functions}\label{i:sec:11.5}

\begin{thm}\label{i:11.5.1}
  Let \(I\) be a bounded interval, and let \(f\) be a function which is uniformly continuous on \(I\).
  Then \(f\) is Riemann integrable.
\end{thm}

\begin{proof}
  From \cref{i:9.9.15} we see that \(f\) is bounded.
  Now we have to show that \(\underline{\int}_I f = \overline{\int}_I f\).

  If \(I\) is a point or the empty set then the theorem is trivial, so let us assume that \(I\) is one of the four intervals \([a, b]\), \((a, b)\), \((a, b]\), or \([a, b)\) for some real numbers \(a < b\).

      Let \(\varepsilon > 0\) be arbitrary.
      By uniform continuity, there exists a \(\delta > 0\) such that \(\abs{f(x) - f(y)} < \varepsilon\) whenever \(x, y \in I\) are such that \(\abs{x - y} < \delta\).
      By the Archimedean principle, there exists an integer \(N > 0\) such that \((b - a) / N < \delta\).

      Note that we can partition \(I\) into \(N\) intervals \(J_1, \dots, J_N\), each of length \((b - a) / N\).
      (How? One has to treat each of the cases \([a, b]\), \((a, b)\), \((a, b]\), \([a, b)\) slightly differently.)
  By \cref{i:11.3.12}, we thus have
  \[
    \overline{\int}_I f \leq \sum_{k = 1}^N \big(\sup_{x \in J_k} f(x)\big) \abs{J_k}
  \]
  and
  \[
    \underline{\int}_I f \geq \sum_{k = 1}^N \big(\inf_{x \in J_k} f(x)\big) \abs{J_k}
  \]
  so in particular
  \[
    \overline{\int}_I f - \underline{\int}_I f \leq \sum_{k = 1}^N \big(\sup_{x \in J_k} f(x) - \inf_{x \in J_k} f(x)\big) \abs{J_k}.
  \]
  However, we have \(\abs{f(x) - f(y)} < \varepsilon\) for all \(x, y \in J_k\), since \(\abs{J_k} = (b - a) / N < \delta\).
  In particular we have
  \[
    f(x) < f(y) + \varepsilon \text{ for all } x, y \in J_k.
  \]
  Taking suprema in \(x\), we obtain
  \[
    \sup_{x \in J_k} f(x) \leq f(y) + \varepsilon \text{ for all } y \in J_k,
  \]
  and then taking infima in \(y\) we obtain
  \[
    \sup_{x \in J_k} f(x) \leq \inf_{y \in J_k} f(y) + \varepsilon.
  \]
  Inserting this bound into our previous inequality, we obtain
  \[
    \overline{\int}_I f - \underline{\int}_I f \leq \sum_{k = 1}^N \varepsilon \abs{J_k},
  \]
  but by \cref{i:11.1.13} we thus have
  \[
    \overline{\int}_I f - \underline{\int}_I f \leq \varepsilon (b - a).
  \]
  But \(\varepsilon > 0\) was arbitrary, while \((b - a)\) is fixed.
  Thus \(\overline{\int}_I f - \underline{\int}_I f\) cannot be positive.
  By \cref{i:11.3.3} and the definition of Riemann integrability we thus have that \(f\) is Riemann integrable.
\end{proof}

\begin{cor}\label{i:11.5.2}
  Let \([a, b]\) be a closed interval, and let \(f : [a, b] \to \R\) be continuous.
  Then \(f\) is Riemann integrable.
\end{cor}

\begin{proof}
  Combining \cref{i:11.5.1} with \cref{i:9.9.16} we are done.
\end{proof}

\begin{note}
  Note that \cref{i:11.5.2} is not true if \([a, b]\) is replaced by any other sort of interval, since it is not even guaranteed then that continuous functions are bounded.
  For instance, the function \(f : (0, 1) \to \R\) defined by \(f(x) \coloneqq 1 / x\) is continuous but not Riemann integrable.
  However, if we assume that a function is both continuous \emph{and} bounded, we can recover Riemann integrability (see \cref{i:11.5.3}).
\end{note}

\begin{prop}\label{i:11.5.3}
  Let \(I\) be a bounded interval, and let \(f : I \to \R\) be both continuous and bounded.
  Then \(f\) is Riemann integrable on \(I\).
\end{prop}

\begin{proof}
  If \(I\) is a point or an empty set then the claim is trivial;
  if \(I\) is a closed interval the claim follows from \cref{i:11.5.2}.
  So let us assume that \(I\) is of the form \((a, b]\), \((a, b)\), or \([a, b)\) for some \(a < b\).

  We have a bound \(M\) for \(f\), so that \(-M \leq f(x) \leq M\) for all \(x \in I\).
  Now let \(0 < \varepsilon < (b - a) / 2\) be a small number.
  The function \(f\) when restricted to the interval \([a + \varepsilon, b - \varepsilon]\) is continuous, and hence Riemann integrable by \cref{i:11.5.2}.
  In particular, we can find a piecewise constant function \(h : [a + \varepsilon, b - \varepsilon] \to \R\) which majorizes \(f\) on \([a + \varepsilon, b - \varepsilon]\) such that
  \[
    \int_{[a + \varepsilon, b - \varepsilon]} h \leq \int_{[a + \varepsilon, b - \varepsilon]} f + \varepsilon.
  \]
  Define \(\tilde{h} : I \to \R\) by
  \[
    \tilde{h}(x) \coloneqq \begin{dcases}
      h(x) & \text{if } x \in [a + \varepsilon, b - \varepsilon]             \\
      M    & \text{if } x \in I \setminus [a + \varepsilon, b - \varepsilon]
    \end{dcases}
  \]
  Clearly \(\tilde{h}\) is piecewise constant on \(I\) and majorizes \(f\);
  by \cref{i:11.2.16} we have
  \[
    \int_I \tilde{h} = \varepsilon M + \int_{[a + \varepsilon, b - \varepsilon]} h + \varepsilon M \leq \int_{[a + \varepsilon, b - \varepsilon]} f + (2M + 1) \varepsilon.
  \]
  In particular we have
  \[
    \overline{\int}_I f \leq \int_{[a + \varepsilon, b - \varepsilon]} f + (2M + 1) \varepsilon.
  \]
  This is true since \(\tilde{h}\) majorize \(f\).
  A similar argument gives
  \[
    \underline{\int}_I f \geq \int_{[a + \varepsilon, b - \varepsilon]} f - (2M + 1) \varepsilon.
  \]
  and hence
  \[
    \overline{\int}_I f - \underline{\int}_I f \leq (4M + 2) \varepsilon.
  \]
  But \(\varepsilon\) is arbitrary, and so we can argue as in the proof of \cref{i:11.5.1} to conclude Riemann integrability.
\end{proof}

\begin{note}
  From \cref{i:11.5.1}, \cref{i:11.5.2} and \cref{i:11.5.3} we see that if we can show a function \(f\) being \emph{uniformly continuous} (not just continuous) on some bounded interval \(I\), then \(f\) is Riemann integrable on \(I\).
\end{note}

\begin{defn}\label{i:11.5.4}
  Let \(I\) be a bounded interval, and let \(f : I \to \R\).
  We say that \(f\) is \emph{piecewise continuous on \(I\)} iff there exists a partition \(\mathbf{P}\) of \(I\) such that \(f|_J\) is continuous on \(J\) for all \(J \in \mathbf{P}\).
\end{defn}

\setcounter{thm}{5}
\begin{prop}\label{i:11.5.6}
  Let \(I\) be a bounded interval, and let \(f : I \to \R\) be both piecewise continuous and bounded.
  Then \(f\) is Riemann integrable.
\end{prop}

\begin{proof}
  Since \(f\) is piecewise continuous on \(I\), by \cref{i:11.5.4} \(\exists \mathbf{P}\) such that \(\mathbf{P}\) is a partition of \(I\) and \(f|_J\) is continuous on \(J\) for all \(J \in \mathbf{P}\).
  Since \(f\) is bounded, we know that \(f|_J\) is bounded for all \(J \in \mathbf{P}\).
  Thus by \cref{i:11.5.3} \(f|_J\) is Riemann integrable on \(J\) for all \(J \in \mathbf{P}\).
  For each \(J \in \mathbf{P}\), we define \(F_J : I \to \R\) to be the function
  \[
    F_J(x) = \begin{dcases}
      f|_J(x) & \text{if } x \in J    \\
      0       & \text{if } x \notin J
    \end{dcases}
  \]
  Then by \cref{i:11.4.1}(g) \(F|_J\) is Riemann integrable for all \(J \in \mathbf{P}\) and
  \begin{align*}
    \sum_{J \in \mathbf{P}} \int_I F_J & = \sum_{J \in \mathbf{P}} \int_J f|_J &  & \by{i:11.4.1}[g] \\
                                       & = \int_I f.                           &  & \by{i:ex:11.4.3}
  \end{align*}
  Thus \(f\) is Riemann integrable on \(I\).
\end{proof}

\exercisesection

\begin{ex}\label{i:ex:11.5.1}
  Prove \cref{i:11.5.6}.
\end{ex}

\begin{proof}
  See \cref{i:11.5.6}.
\end{proof}
