\section{The Riemann-Stieltjes integral}\label{sec 11.8}

\begin{note}
    Let \(I\) be a bounded interval, let \(\alpha : I \to \mathbf{R}\) be a monotone increasing function, and let \(f : I \to \mathbf{R}\) be a function.
    Then there is a generalization of the Riemann integral, known as the \emph{Riemann-Stieltjes integral}.
    This integral is defined just like the Riemann integral, but with one twist:
    instead of taking the length \(\abs*{J}\) of intervals \(J\), we take the \(\alpha\)-length \(\alpha[J]\), defined as follows.
    If \(J\) is a point or the empty set, then \(\alpha[J] \coloneqq 0\).
    If \(J\) is an interval of the form \([a, b]\), \((a, b)\), \((a, b]\), or \([a, b)\), then \(\alpha[J] \coloneqq \alpha(b) - \alpha(a)\).
    Note that in the special case where \(\alpha\) is the identity function \(\alpha(x) \coloneqq x\), then \(\alpha[J]\) is just the same as \(\abs*{J}\).
    However, for more general monotone functions \(\alpha\), the \(\alpha\)-length \(\alpha[J]\) is a different quantity from \(\abs*{J}\).
    Nevertheless, it turns out one can still do much of the above theory, but replacing \(\abs*{J}\) by \(\alpha[J]\) throughout.
\end{note}

\begin{definition}[\(\alpha\)-length]\label{11.8.1}
    Let \(I\) be a bounded interval, and let \(\alpha : X \to \mathbf{R}\) be a monotone increasing function defined on some interval \(X\) which contains \(I\).
    Then we define the \emph{\(\alpha\)-length} \(\alpha[I]\) of \(I\) as follows.
    If \(I\) is the empty set, we set \(\alpha[I] = 0\).
    If \(I\) is a point of the form \(\{a\}\) a point for some real number \(a\), we set
    \[
        \alpha[\{\alpha\}] \coloneqq \lim_{x \to a^+ ; x \in X} \alpha(x) - \lim_{x \to a^- ; x \in X} \alpha(x),
    \]
    with the convention that \(\lim_{x \to a^+ ; x \in X} \alpha(x)\) (resp. \(\lim_{x \to a^- ; x \in X} \alpha(x)\)) is \(\alpha(a)\) when \(a\) is the right (resp. left) endpoint of \(X\).
    If \(I\) is an interval of the form \((a, b)\) for some real numbers \(b > a\), set
    \[
        \alpha[(a, b)] \coloneqq \lim_{x \to b^- ; x \in X} \alpha(x) - \lim_{x \to a^+ ; x \in X} \alpha(x).
    \]
    If \(I\) is an interval of the form \([a, b)\), \((a, b]\), or \([a, b]\) for some real numbers \(b > a\), then we set
    \[
        \alpha[I] = \begin{cases}
            \alpha(\{a\}) + \alpha((a, b)) & \text{if } I = [a, b) \\
            \alpha((a, b)) + \alpha(\{b\}) & \text{if } I = (a, b] \\
            \alpha(\{a\}) + \alpha((a, b)) + \alpha(\{b\}) & \text{if } I = [a, b]
        \end{cases}
    \]
\end{definition}

\begin{note}
    In the special case when \(\alpha\) is continuous, the definition of \(\alpha[I]\) where \(I\) is of the form \((a, b)\), \([a, b)\), \((a, b]\), or \([a, b]\) simplifies to \(\alpha[I] = \alpha(b) - \alpha(a)\).
\end{note}

\begin{note}
    We sometimes write \(\alpha\big|_a^b\) or \(\alpha(x)\big|_{x = a}^{x = b}\) instead of \(\alpha[[a, b]]\).
\end{note}

\setcounter{theorem}{3}
\begin{lemma}\label{11.8.4}
    Let \(I\) be a bounded interval, let \(\alpha : X \to \mathbf{R}\) be a function defined on some domain \(X\) which contains \(I\), and let \(\mathbf{P}\) be a partition of \(I\).
    Then we have
    \[
        \alpha[I] = \sum_{J \in \mathbf{P}} \alpha[J].
    \]
\end{lemma}

\begin{proof}
    We prove this by induction on \(n\).
    More precisely, we let \(P(n)\) be the property that whenever \(I\) is a bounded interval, and whenever \(\mathbf{P}\) is a partition of \(I\) with cardinality \(n\), that \(\alpha[I] = \sum_{J \in \mathbf{P}} \alpha[J]\).

    The base case \(P(0)\) is trivial;
    the only way that \(I\) can be partitioned into an empty partition is if \(I\) is itself empty, at which point the claim is easy.
    The case \(P(1)\) is also very easy;
    the only way that \(I\) can be partitioned into a singleton set \(\{J\}\) is if \(J = I\), at which point the claim is again very easy.

    Now suppose inductively that \(P(n)\) is true for some \(n \geq 1\), and now we prove \(P(n + 1)\).
    Let \(I\) be a bounded interval, and let \(\mathbf{P}\) be a partition of \(I\) of cardinality \(n + 1\).

    If \(I\) is the empty set or a point, then all the intervals in \(\mathbf{P}\) must also be either the empty set or a point, and so every interval has length zero and the claim is trivial.
    Thus we will assume that \(I\) is an interval of the form \((a, b)\), \((a, b]\), \([a, b)\), or \([a, b]\).

            Let us first suppose that \(b \in I\), i.e., \(I\) is either \((a, b]\) or \([a, b]\).
    Since \(b \in I\), we know that one of the intervals \(K\) in \(\mathbf{P}\) contains \(b\).
    Since \(K\) is contained in \(I\), it must therefore be of the form \((c, b]\), \([c, b]\), or \(\{b\}\) for some real number \(c\), with \(a \leq c \leq b\) (in the latter case of \(K = \{b\}\), we set \(c \coloneqq b\)).
    In particular, this means that the set \(I \setminus K\) is also an interval of the form \([a, c]\), \((a, c)\), \((a, c]\), \([a, c)\) when \(c > a\), or a point or empty set when \(a = c\).
    Either way, by Definition \ref{11.8.1} we easily see that
    \[
        \alpha[I] = \alpha(b) - \alpha(a) = \alpha(b) - \alpha(c) + \alpha(c) - \alpha(a) = \alpha[K] + \alpha[I \setminus K].
    \]
    On the other hand, since \(\mathbf{P}\) forms a partition of \(I\), we see that \(\mathbf{P} \setminus \{K\}\) forms a partition of \(I \setminus K\).
    By the induction hypothesis, we thus have
    \[
        \alpha[I \setminus K] = \sum_{J \in \mathbf{P} \setminus \{K\}} \alpha[J].
    \]
    Combining these two identities (and using the laws of addition for finite sets, see Proposition \ref{7.1.11}(e)) we obtain
    \[
        \alpha[I] = \sum_{J \in \mathbf{P}} \alpha[J]
    \]
    as desired.

    Now suppose that \(b \notin I\), i.e., \(I\) is either \((a, b)\) or \([a, b)\).
    Then one of the intervals \(K\) also is of the form \((c, b)\) or \([c, b)\) (see Exercise \ref{ex 11.1.3}).
            In particular, this means that the set \(I \setminus K\) is also an interval of the form \([a, c]\), \((a, c)\), \((a, c]\), \([a, c)\) when \(c > a\), or a point or empty set when \(a = c\).
    The rest of the argument then proceeds as above.
\end{proof}

\begin{definition}[piecewise constant Riemann-Stieltjes integral]\label{11.8.5}
    Let \(I\) be a bounded interval, and let \(\mathbf{P}\) be a partition of \(I\).
    Let \(\alpha : X \to \mathbf{R}\) be a function defined on some domain \(X\) which contains \(I\), and let \(f : I \to \mathbf{R}\) be a function which is piecewise constant with respect to \(\mathbf{P}\).
    Then we define
    \[
        p.c. \int_{[\mathbf{P}]} f \, d\alpha \coloneqq \sum_{J \in \mathbf{P}} c_J \alpha[J]
    \]
    where \(c_J\) is the constant value of \(f\) on \(J\).
\end{definition}