\section{The two fundamental theorems of calculus}\label{sec 11.9}

\begin{theorem}[First Fundamental Theorem of Calculus]\label{11.9.1}
    Let \(a < b\) be real numbers, and let \(f : [a, b] \to \mathbf{R}\) be a Riemann integrable function.
    Let \(F : [a, b] \to \mathbf{R}\) be the function
    \[
        F(x) \coloneqq \int_{[a, x]} f.
    \]
    Then \(F\) is continuous.
    Furthermore, if \(x_0 \in [a, b]\) and \(f\) is continuous at \(x_0\), then \(F\) is differentiable at \(x_0\), and \(F'(x_0) = f(x_0)\).
\end{theorem}

\begin{proof}
    Since \(f\) is Riemann integrable, it is bounded (by Definition \ref{11.3.4}).
    Thus we have some real number \(M\) such that \(-M \leq f(x) \leq M\) for all \(x \in [a, b]\).

    Now let \(x < y\) be two elements of \([a, b]\).
    Then notice that
    \[
        F(y) - F(x) = \int_{[a, y]} f - \int_{[a, x]} f = \int_{[x, y]} f
    \]
    by Theorem \ref{11.4.1}(h).
    By Theorem \ref{11.4.1}(e) we thus have
    \[
        \int_{[x, y]} f \leq \int_{[x, y]} M = p.c. \int_{[x, y]} M = M(y - x)
    \]
    and
    \[
        \int_{[x, y]} f \geq \int_{[x, y]} -M = p.c. \int_{[x, y]} -M = -M(y - x)
    \]
    and thus
    \[
        \abs*{F(y) - F(x)} \leq M(y - x).
    \]
    This is for \(y > x\).
    By interchanging \(x\) and \(y\) we thus see that
    \[
        \abs*{F(y) - F(x)} \leq M(x - y)
    \]
    when \(x > y\).
    Also, we have \(F(y) - F(x) = 0\) when \(x = y\).
    Thus in all
    three cases we have
    \[
        \abs*{F(y) - F(x)} \leq M \abs*{x - y}.
    \]
    Now let \(z \in [a, b]\), and let \((z_n)_{n = 0}^\infty\) be any sequence in \([a, b]\) converging to \(z\).
    Then we have
    \[
        -M \abs*{z_n - z} \leq F(z_n) - F(z) \leq M \abs*{z_n - z}
    \]
    for each \(n\).
    But \(-M \abs*{z_n - z}\) and \(M \abs*{z_n - z}\) both converge to \(0\) as \(n \to \infty\), so by the squeeze test \(F(z_n) - F(z)\) converges to \(0\) as \(n \to \infty\), and thus \(\lim_{n \to \infty} F(z_n) = F(z)\).
    Since this is true for all sequences \(z_n \in [a, b]\) converging to \(z\), we thus see that \(F\) is continuous at \(z\) (by Proposition \ref{9.4.7}).
    Since \(z\) was an arbitrary element of \([a, b]\), we thus see that \(F\) is continuous
    (The above proof also show that when \(F\) is Lipschitz continuous, \(F\) is also continuous, see Exercise \ref{ex 10.2.6}).

    Now suppose that \(x_0 \in [a, b]\), and \(f\) is continuous at \(x_0\).
    Choose any \(\varepsilon > 0\).
    Then by continuity, we can find a \(\delta > 0\) such that \(\abs*{f(x) - f(x_0)} \leq \varepsilon\) for all \(x\) in the interval \(I \coloneqq [x_0 - \delta, x_0 + \delta] \cap [a, b]\), or in other words
    \[
        f(x_0) - \varepsilon \leq f(x) \leq f(x_0) + \varepsilon \text{ for all } x \in I.
    \]
    We now show that
    \[
        \abs*{F(y) - F(x_0) - f(x_0)(y - x_0)} \leq \varepsilon \abs*{y - x_0}
    \]
    for all \(y \in I\), since Proposition \ref{10.1.7} will then imply that \(F\) is differentiable at \(x_0\) with derivative \(F'(x_0) = f(x_0)\) as desired.

    Now fix \(y \in I\).
    There are three cases.
    If \(y = x_0\), then \(F(y) - F(x_0) - f(x_0)(y - x_0) = 0\) and so the claim is obvious.
    If \(y > x_0\), then
    \[
        F(y) - F(x_0) = \int_{[x_0, y]} f.
    \]
    Since \(x_0\), \(y \in I\), and \(I\) is a connected set (by Corollary \ref{11.1.6}), then \([x_0, y]\) is a subset of \(I\), and thus we have
    \[
        f(x_0) - \varepsilon \leq f(x) \leq f(x_0) + \varepsilon \text{ for all } x \in [x_0, y],
    \]
    and thus by Theorem \ref{11.4.1}(e)
    \[
        (f(x_0) - \varepsilon) (y - x_0) \leq \int_{[x_0, y]} f \leq (f(x_0) + \varepsilon) (y - x_0)
    \]
    and so in particular
    \[
        \abs*{F(y) - F(x_0) - f(x_0)(y - x_0)} \leq \varepsilon \abs*{y - x_0}
    \]
    as desired.
    If \(y < x_0\), then
    \[
        F(y) - F(x_0) = - (F(x_0) - F(y)) = -\int_{[y, x_0]} f.
    \]
    Since \(x_0\), \(y \in I\), and \(I\) is a connected set (by Corollary \ref{11.1.6}), then \([y, x_0]\) is a subset of \(I\), and thus we have
    \[
        f(x_0) - \varepsilon \leq f(x) \leq f(x_0) + \varepsilon \text{ for all } x \in [y, x_0],
    \]
    and thus by Theorem \ref{11.4.1}(e)
    \begin{align*}
                 & (f(x_0) - \varepsilon) (x_0 - y) \leq \int_{[y, x_0]} f \leq (f(x_0) + \varepsilon) (x_0 - y)   \\
        \implies & -(f(x_0) - \varepsilon) (y - x_0) \leq \int_{[y, x_0]} f \leq -(f(x_0) + \varepsilon) (y - x_0) \\
        \implies & -\varepsilon (y - x_0) \leq \int_{[y, x_0]} f + f(x_0)(y - x_0) \leq \varepsilon (y - x_0)      \\
        \implies & -\varepsilon (y - x_0) \leq F(x_0) - F(y) + f(x_0)(y - x_0) \leq \varepsilon (y - x_0)          \\
        \implies & -\varepsilon (y - x_0) \leq F(y) - F(x_0) - f(x_0)(y - x_0) \leq \varepsilon (y - x_0)          \\
        \implies & \abs*{F(y) - F(x_0) - f(x_0)(y - x_0)} \leq \varepsilon (y - x_0)
    \end{align*}
    as desired.
\end{proof}

\begin{note}
    Informally, the first fundamental theorem of calculus asserts that
    \[
        \bigg(\int_{[a, x]} f\bigg)'(x) = f(x)
    \]
    given a certain number of assumptions on \(f\).
    Roughly, this means that the derivative of an integral recovers the original function.
\end{note}

\setcounter{theorem}{2}
\begin{definition}[Antiderivatives]\label{11.9.3}
    Let \(I\) be a bounded interval, and let \(f : I \to \mathbf{R}\) be a function.
    We say that a function \(F : I \to \mathbf{R}\) is an \emph{antiderivative} of \(f\) if \(F\) is differentiable on \(I\) and \(F'(x) = f(x)\) for all limit points \(x\) of \(I\).
\end{definition}

\begin{theorem}[Second Fundamental Theorem of Calculus]\label{11.9.4}
    Let \(a < b\) be real numbers, and let \(f : [a, b] \to \mathbf{R}\) be a Riemann integrable function.
    If \(F : [a, b] \to \mathbf{R}\) is an antiderivative of \(f\), then
    \[
        \int_{[a, b]} f = F(b) - F(a).
    \]
\end{theorem}

\begin{proof}
    The claim is trivial when \(b = a\), so assume \(b > a\), so in particular all points of \([a, b]\) are limit points.
    We will use Riemann sums.
    The idea is to show that
    \[
        U(f, \mathbf{P}) \geq F(b) - F(a) \geq L(f, \mathbf{P})
    \]
    for every partition \(\mathbf{P}\) of \([a, b]\).
    The left inequality asserts that \(F(b) - F(a)\) is a lower bound for \(\{U(f, \mathbf{P}) : \mathbf{P} \text{ is a partition of } [a, b]\}\), while the right inequality asserts that \(F(b) - F(a)\) is an upper bound for \(\{L(f, \mathbf{P}) : \mathbf{P} \text{ is a partition of } [a, b]\}\).
    But by Proposition \ref{11.3.12}, this means that
    \[
        \overline{\int}_{[a, b]} f \geq F(b) - F(a) \geq \underline{\int}_{[a, b]} f,
    \]
    but since \(f\) is assumed to be Riemann integrable, both the upper and lower Riemann integral equal \(\int_{[a, b]} f\).
    The claim follows.

    We have to show the bound \(U(f, \mathbf{P}) \geq F(b) - F(a) \geq L(f, \mathbf{P})\).
    We shall just show the first inequality \(U(f, \mathbf{P}) \geq F(b) - F(a)\);
    the other inequality is similar.

    Let \(\mathbf{P}\) be a partition of \([a, b]\).
    From Lemma \ref{11.8.4} we have
    \[
        F(b) - F(a) = \sum_{J \in \mathbf{P}} F[J] = \sum_{J \in \mathbf{P} : J \neq \emptyset} F[J],
    \]
    while from definition we have
    \[
        U(f, \mathbf{P}) = \sum_{J \in \mathbf{P} : J \neq \emptyset} \sup_{x \in J} f(x) \abs*{J}.
    \]
    Thus it will suffice to show that
    \[
        F[J] \leq \sup_{x \in J} f(x) \abs*{J}
    \]
    for all \(J \in \mathbf{P}\)
    (other than the empty set).

    When \(J\) is a point then the claim is clear, since both sides are zero.
    Now suppose that \(J = [c, d], (c, d], [c, d)\), or \((c, d)\) for some \(c < d\).
    Then the left-hand side is \(F[J] = F(d) - F(c)\).
    Note that \(F\), being differentiable, is continuous, so we may use the simplified formula for the \(F\)-length as opposed to the more complicated one in Definition \ref{11.8.1}.
    By the mean-value theorem (Corollary \ref{10.2.9}), this is equal to \((d - c) F'(e)\) for some \(e \in J\).
    But since \(F'(e) = f(e)\), we thus have
    \[
        F[J] = (d - c) f(e) = f(e) \abs*{J} \leq \sup_{x \in J} f(x) \abs*{J}
    \]
    as desired.
\end{proof}

\begin{note}
    One can use the second fundamental theorem of calculus to compute integrals relatively easily provided that you can find an anti-derivative of the integrand \(f\).
    the first fundamental theorem of calculus ensures that every \emph{continuous} Riemann integrable function has an anti-derivative.
    For discontinuous functions, the situation is more complicated.
    Also, not every function with an anti-derivative is Riemann integrable.
\end{note}

\begin{lemma}\label{11.9.5}
    Let \(I\) be a bounded interval, and let \(f : I \to \mathbf{R}\) be a function.
    Let \(F : I \to \mathbf{R}\) and \(G : I \to \mathbf{R}\) be two antiderivatives of \(f\).
    Then there exists a real number \(C\) such that \(F(x) = G(x) + C\) for all \(x \in I\).
\end{lemma}

\begin{proof}
    Since \(I\) is a bounded interval, \(\forall\ x \in \overline{I}\), \(x\) is a limit point, thus \(\forall\ x \in I\), \(x\) is also a limit point.
    Since \(F, G\) are antiderivatives of \(f\), by Definition \ref{11.9.3} we know that \(F, G\) are differentiable on \(I\).
    By Theorem \ref{10.1.13}(f) we know that \(F - G\) is differentiable on \(I\), and thus by Corollary \ref{10.1.12} we know that \(F - G\) are continuous on \(I\).
    Let \(x, y \in I\) and \(x < y\).
    Since \(I\) is a bounded interval, by Lemma \ref{11.1.10} we know that \(I\) is connected, and thus by Definition \ref{11.1.1} \([x, y] \subseteq I\).
    By the mean-value theorem (Corollary \ref{10.2.9}) we know that
    \[
        \exists\ c \in I : \frac{(F - G)(x) - (F - G)(y)}{x - y} = (F - G)'(c).
    \]
    Thus we have
    \begin{align*}
                 & \frac{(F - G)(x) - (F - G)(y)}{x - y} = (F - G)'(c)                                          \\
        \implies & \frac{(F - G)(x) - (F - G)(y)}{x - y} = F'(c) - G'(c) & \text{(by Theorem \ref{10.1.13}(f))} \\
        \implies & \frac{(F - G)(x) - (F - G)(y)}{x - y} = 0             & \text{(by hypothesis)}               \\
        \implies & (F - G)(x) = (F - G)(y)                                                                      \\
        \implies & F(x) - G(x) = F(y) - G(y)                             & \text{(by Definition \ref{9.2.1})}   \\
        \implies & F(x) = G(x) + F(y) - G(y).
    \end{align*}
    By setting \(C = F(y) - G(y)\) we are done.
\end{proof}

\begin{proposition}\label{11.9.6}
    Let \(a < b\) be real numbers, let \(x \in [a, b]\) and let \(f : [a, b] \to \mathbf{R}\) be a monotone function.
    If \(f\) is monotone increasing, then we have
    \[
        f(x_0+) = \inf_{x \in [a, b] \cap (x_0, \infty)} f(x)
    \]
    and
    \[
        f(x_0-) = \sup_{x \in [a, b] \cap (-\infty, x_0)} f(x).
    \]
    If \(f\) is monotone decreasing, then we have
    \[
        f(x_0+) = \sup_{x \in [a, b] \cap (x_0, \infty)} f(x)
    \]
    and
    \[
        f(x_0-) = \inf_{x \in [a, b] \cap (-\infty, x_0)} f(x).
    \]
\end{proposition}

\begin{proof}
    First suppose that \(f\) is monotone increasing.
    Since \(f\) is monotone increasing, we have
    \[
        \sup_{x \in [a, b] \cap (x_0, \infty)} f(x) \leq \sup_{x \in [a, b]} f(x)
    \]
    and
    \[
        \inf_{x \in [a, b] \cap (x_0, \infty)} f(x) \geq \inf_{x \in [a, b]} f(x)
    \]
    and thus \(f(x_0+)\) and \(f(x_0-)\) are finite.
    Let \(U = \inf_{x \in [a, b] \cap (x_0, \infty)} f(x)\).
    By the definition of \(U\), \(\forall\ \varepsilon \in \mathbf{R}^+\), we can always find an \(x \in [a, b] \cap (x_0, \infty)\) such that \(0 \leq f(x) - U \leq \varepsilon\).
    We also know that for all \(y \in [a, b] \cap (x_0, \infty)\) and \(y \leq x\), we have \(0 \leq f(y) - U \leq f(x) - U \leq \varepsilon\).
    Thus by setting \(\delta = x - x_0\) we have
    \[
        \forall\ y \in [a, b] \cap (x_0, \infty), \abs*{y - x_0} < \delta \implies \abs*{f(y) - U} \leq \varepsilon
    \]
    and by Definition \ref{9.3.6} and Definition \ref{9.5.1} we have \(f(x_0+) = U\).
    Similarly let \(L = \sup_{x \in [a, b] \cap (x_0, \infty)} f(x)\).
    By the definition of \(L\), \(\forall\ \varepsilon \in \mathbf{R}^+\), we can always find an \(x \in [a, b] \cap (-\infty, x_0)\) such that \(0 \leq L - f(x) \leq \varepsilon\).
    We also know that for all \(y \in [a, b] \cap (-\infty, x_0)\) and \(y \geq x\), we have \(0 \leq L - f(y) \leq L - f(x) \leq \varepsilon\).
    Thus by setting \(\delta = x_0 - x\) we have
    \[
        \forall\ y \in [a, b] \cap (-\infty, x_0), \abs*{y - x_0} < \delta \implies \abs*{f(y) - L} \leq \varepsilon
    \]
    and by Definition \ref{9.3.6} and Definition \ref{9.5.1} we have \(f(x_0-) = L\).

    Now suppose that \(f\) is monotone decreasing.
    Since \(f\) is monotone decreasing, we have
    \[
        \sup_{x \in [a, b] \cap (x_0, \infty)} f(x) \leq \sup_{x \in [a, b]} f(x)
    \]
    and
    \[
        \inf_{x \in [a, b] \cap (x_0, \infty)} f(x) \geq \inf_{x \in [a, b]} f(x)
    \]
    and thus \(f(x_0+)\) and \(f(x_0-)\) are finite.
    Let \(U = \sup_{x \in [a, b] \cap (x_0, \infty)} f(x)\).
    By the definition of \(U\), \(\forall\ \varepsilon \in \mathbf{R}^+\), we can always find an \(x \in [a, b] \cap (x_0, \infty)\) such that \(0 \leq U - f(x) \leq \varepsilon\).
    We also know that for all \(y \in [a, b] \cap (x_0, \infty)\) and \(y \leq x\), we have \(0 \leq U - f(y) \leq U - f(x) \leq \varepsilon\).
    Thus by setting \(\delta = x - x_0\) we have
    \[
        \forall\ y \in [a, b] \cap (x_0, \infty), \abs*{y - x_0} < \delta \implies \abs*{f(y) - U} \leq \varepsilon
    \]
    and by Definition \ref{9.3.6} and Definition \ref{9.5.1} we have \(f(x_0+) = U\).
    Similarly let \(L = \inf_{x \in [a, b] \cap (x_0, \infty)} f(x)\).
    By the definition of \(L\), \(\forall\ \varepsilon \in \mathbf{R}^+\), we can always find an \(x \in [a, b] \cap (-\infty, x_0)\) such that \(0 \leq f(x) - L \leq \varepsilon\).
    We also know that for all \(y \in [a, b] \cap (-\infty, x_0)\) and \(y \geq x\), we have \(0 \leq f(y) - L \leq f(x) - L \leq \varepsilon\).
    Thus by setting \(\delta = x_0 - x\) we have
    \[
        \forall\ y \in [a, b] \cap (-\infty, x_0), \abs*{y - x_0} < \delta \implies \abs*{f(y) - L} \leq \varepsilon
    \]
    and by Definition \ref{9.3.6} and Definition \ref{9.5.1} we have \(f(x_0-) = L\).
\end{proof}

\exercisesection

\begin{exercise}\label{ex 11.9.1}
    Let \(f : [0, 1] \to \mathbf{R}\) be the function in Exercise \ref{ex 9.8.5}.
    Show that for every rational number \(q \in \mathbf{Q} \cap (0, 1)\), the function \(F : [0, 1] \to \mathbf{R}\) defined by the formula \(F(x) \coloneqq \int_{[0, x]} f\) is not differentiable at \(q\).
\end{exercise}

\begin{proof}
    By Exercise \ref{ex 9.8.5} we know that \(f\) is strictly monotone increasing, thus by Proposition \ref{11.6.1} we know that \(f\) is Riemann integrable and \(F\) is well-defined.
    By Exercise \ref{ex 9.8.5} \(f\) is not continuous at \(q\), thus by Exercise \ref{ex 11.9.3} we know that \(F\) is not differentiable at \(q\).
\end{proof}

\begin{exercise}\label{ex 11.9.2}
    Prove Lemma \ref{11.9.5}.
\end{exercise}

\begin{proof}
    See Lemma \ref{11.9.5}.
\end{proof}

\begin{exercise}\label{ex 11.9.3}
    Let \(a < b\) be real numbers, and let \(f : [a, b] \to \mathbf{R}\) be a monotone increasing function.
    Let \(F : [a, b] \to \mathbf{R}\) be the function \(F(x) \coloneqq \int_{[a, x]} f\).
    Let \(x_0\) be an element of \((a, b)\).
    Show that \(F\) is differentiable at \(x_0\) if and only if \(f\) is continuous at \(x_0\).
\end{exercise}

\begin{proof}
    Since \(f\) is monotone increasing, by Proposition \ref{11.6.1} we know that \(f\) is Riemann integrable and \(F\) is well-defined.
    If \(f\) is continuous at \(x_0\) then by Theorem \ref{11.9.1} we know that \(F\) is differentiable at \(x_0\).
    So we only need to show that if \(F\) is differentiable at \(x_0\) then \(f\) is continuous at \(x_0\).

    Suppose that \(F\) is differentiable at \(x_0\).
    Suppose for sake of contradiction that \(f\) is not continuous at \(x_0\).
    Since \(f\) is monotone increasing, by Proposition \ref{11.9.6} we know that both \(f(x_0+)\) and \(f(x_0-)\) exist.
    Since \(f\) is not continuous at \(x_0\), we have \(f(x_0) \neq f(x_0-)\) or \(f(x_0) \neq f(x_0+)\).

    First suppose that \(f(x_0) \neq f(x_0-)\).
    Then by Proposition \ref{11.9.6} we have \(f(x_0-) < f(x_0)\).
    If \(x \in [a, b] \cap (-\infty, x_0)\), then we have
    \[
        \frac{F(x) - F(x_0)}{x - x_0} = \frac{-\int_{[x, x_0]} f}{x - x_0} = \frac{\int_{[x, x_0]} f}{x_0 - x} \leq \frac{\int_{[x, x_0]} f(x_0-)}{x_0 - x} = f(x_0-).
    \]
    If \(x \in [a, b] \cap (x_0, \infty)\), then we have
    \[
        \frac{F(x) - F(x_0)}{x - x_0} = \frac{\int_{[x_0, x]} f}{x - x_0} \geq \frac{\int_{[x_0, x]} f(x_0)}{x - x_0} = f(x_0).
    \]
    Thus we have
    \[
        F'(x_0-) \leq f(x_0-) < f(x_0) \leq F'(x_0+)
    \]
    and \(F'(x_0)\) does not exists, a contradiction.

    Now suppose that \(f(x_0) \neq f(x_0+)\).
    Then by Proposition \ref{11.9.6} we have \(f(x_0) < f(x_0+)\).
    If \(x \in [a, b] \cap (-\infty, x_0)\), then we have
    \[
        \frac{F(x) - F(x_0)}{x - x_0} = \frac{-\int_{[x, x_0]} f}{x - x_0} = \frac{\int_{[x, x_0]} f}{x_0 - x} \leq \frac{\int_{[x, x_0]} f(x_0)}{x_0 - x} = f(x_0).
    \]
    If \(x \in [a, b] \cap (x_0, \infty)\), then we have
    \[
        \frac{F(x) - F(x_0)}{x - x_0} = \frac{\int_{[x_0, x]} f}{x - x_0} \geq \frac{\int_{[x_0, x]} f(x_0+)}{x - x_0} = f(x_0+).
    \]
    Thus we have
    \[
        F'(x_0-) \leq f(x_0) < f(x_0+) \leq F'(x_0+)
    \]
    and \(F'(x_0)\) does not exists, a contradiction.
    So we conclude that \(f\) is continuous at \(x_0\).
\end{proof}