\section{Addition}\label{i:sec:2.2}

\begin{defn}[Addition of natural numbers]\label{i:2.2.1}
  Let \(m\) be a natural numbers.
  To add zero to \(m\), we define \(0 + m \coloneqq m\).
  Now suppose inductively that we have defined how to add \(n\) to \(m\).
  Then we can add \(n\pp\) to \(m\) by defining \((n\pp) + m \coloneqq (n + m)\pp\).
\end{defn}

\begin{note}
  From a logical point of view, there is no difference between a lemma, proposition, theorem, or corollary
  - they are all claims waiting to be proved.
  However, we use these terms to suggest different levels of importance and difficulty.
  A lemma is an easily proved claim which is helpful for proving other propositions and theorems, but is usually not particularly interesting in its own right.
  A proposition is a statement which is interesting in its own right, while a theorem is a more important statement than a proposition which says something definitive on the subject, and often takes more effort to prove than a proposition or lemma.
  A corollary is a quick consequence of a proposition or theorem that was proven recently.
\end{note}

\begin{ac}\label{i:ac:2.2.1}
  The sum \(n + m\) of two natural numbers \(n, m\) is again a natural number.
\end{ac}

\begin{proof}[\pf{i:ac:2.2.1}]
  Let \(m, n\) be a natural number.
  We use induction on \(n\).
  For \(n = 0\), by \cref{i:2.2.1} we have \(0 + m = m\), which is a natural number by definition.
  So the base case holds.
  Suppose inductively that for some natural number \(n\) we know that \(n + m\) is a natural number.
  We want to show that \((n\pp) + m\) is also a natural number.
  By \cref{i:2.2.1} we have \((n\pp) + m = (n + m)\pp\).
  By induction hypothesis we know that \(n + m\) is a natural number.
  Thus by \cref{i:2.2} we know that \((n + m)\pp\) is again a natural number.
  This closes the induction.
\end{proof}

\begin{lem}\label{i:2.2.2}
  For any natural number \(n\), \(n + 0 = n\).
\end{lem}

\begin{proof}[\pf{i:2.2.2}]
  We use induction on \(n\).
  The base case \(0 + 0 = 0\) follows since we know that \(0 + m = m\) for every natural number \(m\) (\cref{i:2.2.1}), and \(0\) is a natural number (\cref{i:2.1}).
  Now suppose inductively that \(n + 0 = n\).
  We wish to show that \((n\pp) + 0 = n\pp\).
  But by \cref{i:2.2.1}, \((n\pp) + 0\) is equal to \((n + 0)\pp\), which is equal to \(n\pp\) since \(n + 0 = n\).
  This closes the induction.
\end{proof}

\begin{lem}\label{i:2.2.3}
  For any natural numbers \(n\) and \(m\), \(n + (m\pp) = (n + m)\pp\).
\end{lem}

\begin{proof}[\pf{i:2.2.3}]
  We induct on \(n\) (keeping \(m\) fixed).
  We first consider the base case \(n = 0\).
  In this case we have to prove \(0 + (m\pp) = (0 + m)\pp\).
  But by \cref{i:2.2.1}, \(0 + (m\pp) = m\pp\) and \(0 + m = m\), so both sides are equal to \(m\pp\) and are thus equal to each other.
  Now we assume inductively that \(n + (m\pp) = (n + m)\pp\);
  we now have to show that \((n\pp) + (m\pp) = ((n\pp) + m)\pp\).
  The left-hand side is \((n + (m\pp))\pp\) by \cref{i:2.2.1}, which is equal to \(((n+m)\pp)\pp\) by the inductive hypothesis.
  Similarly, we have \((n\pp) + m = (n + m)\pp\) by \cref{i:2.2.1}, and so the right-hand side is also equal to \(((n + m)\pp)\pp\).
  Thus both sides are equal to each other, and we have closed the induction.
\end{proof}

\begin{ac}\label{i:ac:2.2.2}
  For any natural number \(n\), we have \(n\pp = n + 1\).
\end{ac}

\begin{proof}[\pf{i:ac:2.2.2}]
  Since \(n\) is a natural number, by \cref{i:2.2} we know that \(n\pp\) is also a natural number.
  Thus we can apply \cref{i:2.2.2} to derive the following fact:
  \begin{align*}
    n\pp & = (n\pp) + 0 &  & \by{i:2.2.2} \\
         & = (n + 0)\pp &  & \by{i:2.2.1} \\
         & = n + (0\pp) &  & \by{i:2.2.3} \\
         & = n + 1.
  \end{align*}
\end{proof}

\begin{prop}[Addition is commutative]\label{i:2.2.4}
  For any natural numbers \(n\) and \(m\), \(n + m = m + n\).
\end{prop}

\begin{proof}[\pf{i:2.2.4}]
  We shall use induction on \(n\) (keeping \(m\) fixed).
  First we do the base case \(n = 0\), i.e., we show \(0 + m = m + 0\).
  By \cref{i:2.2.1}, \(0 + m = m\), while by \cref{i:2.2.2}, \(m + 0 = m\).
  Thus the base case is done.
  Now suppose inductively that \(n + m = m + n\), now we have to prove that \((n\pp) + m = m + (n\pp)\) to close the induction.
  By \cref{i:2.2.1}, \((n\pp) + m = (n + m)\pp\).
  By \cref{i:2.2.3}, \(m + (n\pp) = (m + n)\pp\), but this is equal to \((n + m)\pp\) by the inductive hypothesis \(n+m=m+n\).
  Thus \((n\pp) + m = m + (n\pp)\) and we have closed the induction.
\end{proof}

\begin{prop}[Addition is associative]\label{i:2.2.5}
  For any natural numbers \(a, b, c\), we have \((a + b) + c = a + (b + c)\).
\end{prop}

\begin{proof}[\pf{i:2.2.5}]
  We use induction on \(c\) and keep both \(a\) and \(b\) fixed.
  For \(c = 0\), we have
  \begin{align*}
    (a + b) + 0 & = a + b        &  & \by{i:2.2.2} \\
                & = a + (b + 0). &  & \by{i:2.2.2}
  \end{align*}
  Thus the base case holds.
  Suppose inductively that \((a + b) + c = a + (b + c)\) for some natural number \(c\).
  We want to show that \((a + b) + (c\pp) = a + (b + (c\pp))\).
  But this is true since
  \begin{align*}
    (a + b) + (c\pp) & = ((a + b) + c)\pp  &  & \by{i:2.2.3} \\
                     & = (a + (b + c))\pp  &  & \byIH        \\
                     & = a + (b + c)\pp    &  & \by{i:2.2.3} \\
                     & = a + (b + (c\pp)). &  & \by{i:2.2.3}
  \end{align*}
  This closes the induction.
\end{proof}

\begin{note}
  Because of associativity showed in \cref{i:2.2.5} we can write sums such as \(a + b + c\) without having to worry about which order the numbers are being added together.
\end{note}

\begin{prop}[Cancellation law]\label{i:2.2.6}
  Let \(a, b, c\) be natural numbers such that \(a + b = a + c\).
  Then we have \(b = c\).
\end{prop}

\begin{proof}[\pf{i:2.2.6}]
  We prove this by induction on \(a\).
  First consider the base case \(a = 0\).
  Then we have \(0 + b = 0 + c\), which by \cref{i:2.2.1} implies that \(b = c\) as desired.
  Now suppose inductively that we have the cancellation law for \(a\) (so that \(a + b = a + c\) implies \(b = c\));
  we now have to prove the cancellation law for \(a\pp\).
  In other words, we assume that \((a\pp) + b = (a\pp) + c\) and need to show that \(b = c\).
  By \cref{i:2.2.1}, \((a\pp) + b = (a + b)\pp\) and \((a\pp) + c = (a + c)\pp\) and so we have \((a + b)\pp = (a + c)\pp\).
  By \cref{i:2.4}, we have \(a + b = a + c\).
  Since we already have the cancellation law for \(a\), we thus have \(b = c\) as desired.
  This closes the induction.
\end{proof}

\begin{defn}[Positive natural numbers]\label{i:2.2.7}
  A natural number \(n\) is said to be \emph{positive} iff it is not equal to \(0\).
\end{defn}

\begin{prop}\label{i:2.2.8}
  If \(a\) is a positive natural number and \(b\) is a natural number, then \(a + b\) is positive (and hence \(b + a\) is also, by \cref{i:2.2.4}).
\end{prop}

\begin{proof}[\pf{i:2.2.8}]
  We use induction on \(b\).
  If \(b = 0\), then \(a + b = a + 0 = a\), which is positive by \cref{i:2.2.7}, so this proves the base case.
  Now suppose inductively that \(a + b\) is positive.
  Then \(a + (b\pp) = (a + b)\pp\), which cannot be zero by \cref{i:2.3}, and is hence positive.
  This closes the induction.
\end{proof}

\begin{cor}\label{i:2.2.9}
  If \(a\) and \(b\) are natural numbers such that \(a + b = 0\), then \(a = 0\) and \(b = 0\).
\end{cor}

\begin{proof}[\pf{i:2.2.9}]
  Suppose for sake of contradiction that \(a \neq 0\) or \(b \neq 0\).
  If \(a \neq 0\) then \(a\) is positive, and hence \(a + b = 0\) is positive by \cref{i:2.2.8}, a contradiction.
  Similarly if \(b \neq 0\) then \(b\) is positive, and again \(a + b = 0\) is positive by \cref{i:2.2.8}, a contradiction.
  Thus \(a\) and \(b\) must both be zero.
\end{proof}

\begin{lem}\label{i:2.2.10}
  Let \(a\) be a positive natural number.
  Then there exists exactly one natural number \(b\) such that \(b\pp = a\).
\end{lem}

\begin{proof}[\pf{i:2.2.10}]
  Let \(P(n)\) be the statement ``either \(n = 0\) or there exists a natural number \(m\) such that \(m\pp = n\)''.
  We use induction to show that \(P(n)\) is true for all natural number \(n\).
  Clearly \(P(0)\) is true.
  So suppose that \(P(n)\) is true for some natural number \(n\).
  We want to show that \(P(n\pp)\) is true.
  By \cref{i:2.3} we know that \(n\pp \neq 0\).
  So we have to show that there exists a natural number \(m\) such that \(m\pp = n\pp\).
  By \cref{i:2.4} we see that \(m = n\).
  Thus \(P(n\pp)\) is true and this closes the induction.

  Now we prove the existence of \(b\).
  From first part of the proof we know that \(P(a)\) is true.
  Since \(a\) is a positive natural number, by \cref{i:2.2.7} we know that \(a \neq 0\).
  Thus there must exist a natural number \(b\) such that \(b\pp = a\).

  Finally we prove the uniqueness of \(b\).
  Suppose that there exists another natural number \(c\) such that \(c\pp = a\).
  But this means \(b\pp = c\pp\).
  Thus by \cref{i:2.4} we have \(b = c\).
\end{proof}

\begin{defn}[Ordering of the natural numbers]\label{i:2.2.11}
  Let \(n\) and \(m\) be natural numbers.
  We say that \(n\) is \emph{greater than or equal to} \(m\), and write \(n \geq m\) or \(m \leq n\), iff we have \(n = m + a\) for some natural number \(a\).
  We say that \(n\) is \emph{strictly greater than} \(m\), and write \(n > m\) or \(m < n\), iff \(n \geq m\) and \(n \neq m\).
\end{defn}

\begin{ac}\label{i:ac:2.2.3}
  We have \(n\pp > n\) for any natural number \(n\).
  Therefore there is no largest natural number \(n\), because the next number \(n\pp\) is always larger.
\end{ac}

\begin{proof}[\pf{i:ac:2.2.3}]
  Let \(n\) be a natural number.
  By \cref{i:ac:2.2.2} we have \(n\pp = n + 1\).
  Since \(1\) is a natural number, by \cref{i:2.2.11} we have \(n\pp \geq n\).
  To show that \(n\pp > n\), by \cref{i:2.2.11} we only need to show that \(n\pp \neq n\).

  We use induction to show that \(n\pp \neq n\).
  For \(n = 0\), we have \(0\pp \neq 0\) by \cref{i:2.3}.
  Thus the base case holds.
  Suppose inductively that \(n\pp \neq n\) for some natural number \(n\).
  Then we have
  \begin{align*}
             & n\pp \neq n          &  & \byIH      \\
    \implies & (n\pp)\pp \neq n\pp. &  & \by{i:2.4}
  \end{align*}
  This closes the induction and we conclude that \(n\pp > n\) for all natural number \(n\).
\end{proof}

\begin{ac}\label{i:ac:2.2.4}
  We have \(n \geq 0\) for every natural number \(n\).
  If \(n\) is a positive natural number, then \(n > 0\).
\end{ac}

\begin{proof}[\pf{i:ac:2.2.4}]
  Let \(n\) be a natural number.
  By \cref{i:2.2.1} we have \(n = 0 + n\).
  Thus by \cref{i:2.2.11} we have \(n \geq 0\).

  Now suppose that \(n\) is a positive natural number.
  By \cref{i:2.2.7} this means \(n \neq 0\).
  From first paragraph we see that \(n \geq 0\).
  Thus by \cref{i:2.2.11} we have \(n > 0\).
\end{proof}

\begin{prop}[Basic properties of order for natural numbers]\label{i:2.2.12}
  Let \(a, b, c\) be natural numbers.
  Then
  \begin{enumerate}
    \item (Order is reflexive) \(a \geq a\).
    \item (Order is transitive) If \(a \geq b\) and \(b \geq c\), then \(a \geq c\).
    \item (Order is anti-symmetric) If \(a \geq b\) and \(b \geq a\), then \(a = b\).
    \item (Addition preserves order) \(a \geq b\) iff \(a + c \geq b + c\).
    \item \(a < b\) iff \(a\pp \leq b\).
    \item \(a < b\) iff \(b = a + d\) for some \emph{positive} number \(d\).
  \end{enumerate}
\end{prop}

\begin{proof}[\pf{i:2.2.12}(a)]
  Let \(a\) be a natural number.
  Then
  \begin{align*}
             & \begin{dcases}
                 0 \text{ is a natural number} \\
                 a = a + 0
               \end{dcases} &  & \by{i:2.1,i:2.2.2}                \\
    \implies & a \geq a.                        &  & \by{i:2.2.11}
  \end{align*}
\end{proof}

\begin{proof}[\pf{i:2.2.12}(b)]
  Let \(a, b, c\) be natural numbers and suppose that \(a \geq b\) and \(b \geq c\).
  By \cref{i:2.2.11} there exist some natural numbers \(d\) and \(e\) such that \(a = b + d\) and \(b = c + e\).
  Then we have
  \begin{align*}
             & \begin{dcases}
                 a = b + d = (c + e) + d = c + (e + d) \\
                 e + d \text{ is a natural number}
               \end{dcases} &  & \by{i:2.2.5,i:ac:2.2.1}                   \\
    \implies & a \geq c.                                &  & \by{i:2.2.11}
  \end{align*}
\end{proof}

\begin{proof}[\pf{i:2.2.12}(c)]
  Let \(a, b\) be natural numbers and suppose that \(a \geq b\) and \(b \geq a\).
  By \cref{i:2.2.11} there exist some natural numbers \(c\) and \(d\) such that \(a = b + c\) and \(b = a + d\).
  Then we have
  \begin{align*}
             & \begin{dcases}
                 a = a + 0 \\
                 a = b + c = (a + d) + c = a + (d + c)
               \end{dcases} &  & \by{i:2.2.2,i:2.2.5}                  \\
    \implies & 0 = d + c                             &  & \by{i:2.2.6} \\
    \implies & d = c = 0                             &  & \by{i:2.2.9} \\
    \implies & a = b + 0 = b.                        &  & \by{i:2.2.2}
  \end{align*}
\end{proof}

\begin{proof}[\pf{i:2.2.12}(d)]
  Let \(a, b, c\) be natural numbers.
  Then we have
  \begin{align*}
             & a \geq b                                                               \\
    \implies & a = b + d \text{ for some natural number } d &  & \by{i:2.2.11}        \\
    \implies & a + c = c + a = c + (b + d)                  &  & \by{i:2.2.4}         \\
             & = (c + b) + d = (b + c) + d                  &  & \by{i:2.2.4,i:2.2.5} \\
    \implies & a + c \geq b + c                             &  & \by{i:2.2.11}
  \end{align*}
  and
  \begin{align*}
             & a + c \geq b + c                                                                 \\
    \implies & a + c = (b + c) + d \text{ for some natural number } d &  & \by{i:2.2.11}        \\
    \implies & c + a = (c + b) + d = c + (b + d)                      &  & \by{i:2.2.4,i:2.2.5} \\
    \implies & a = b + d                                              &  & \by{i:2.2.6}         \\
    \implies & a \geq b.                                              &  & \by{i:2.2.11}
  \end{align*}
  Thus we conclude that \(a \geq b \iff a + c \geq b + c\).
\end{proof}

\begin{proof}[\pf{i:2.2.12}(e)]
  Let \(a, b\) be natural numbers.
  First suppose that \(a < b\).
  Then we have
  \begin{align*}
             & a < b                                                                                         \\
    \implies & \begin{dcases}
                 a \leq b \\
                 a \neq b
               \end{dcases}                                                      &  & \by{i:2.2.11}          \\
    \implies & \begin{dcases}
                 b = a + c \text{ for some natural number } c \\
                 a \neq b
               \end{dcases}                     &  & \by{i:2.2.11}                                           \\
    \implies & c \neq 0                                                            &  & \by{i:2.2.2}         \\
    \implies & c \text{ is a positive natural number}                              &  & \by{i:2.2.8}         \\
    \implies & \text{there exists a natural number } d \text{ such that } d\pp = c &  & \by{i:2.2.10}        \\
    \implies & b = a + (d\pp) = (a + d)\pp = (a\pp) + d                            &  & \by{i:2.2.1,i:2.2.3} \\
    \implies & a\pp \leq b.                                                        &  & \by{i:2.2.11}
  \end{align*}

  Now suppose that \(a\pp \leq b\).
  Then we have
  \begin{align*}
             & a\pp \leq b                                                                 \\
    \implies & b = (a\pp) + c \text{ for some natural number } c &  & \by{i:2.2.11}        \\
    \implies & b = (a\pp) + c = (a + c)\pp = a + (c\pp)          &  & \by{i:2.2.1,i:2.2.3} \\
    \implies & a \leq b.                                         &  & \by{i:2.2.11}
  \end{align*}
  By \cref{i:2.3} we know that \(c\pp \neq 0\).
  Thus by \cref{i:2.2.6} we must have \(b \neq a\).
  (If \(b = a\), then \(b = a + c\pp = a + 0\) implies \(c\pp = 0\), a contradiction.)
  By \cref{i:2.2.11} this means \(a < b\).
  From all proofs above we conclude that \(a < b \iff a\pp \leq b\).
\end{proof}

\begin{proof}[\pf{i:2.2.12}(f)]
  Let \(a, b\) be natural numbers.
  Then we have
  \begin{align*}
         & a < b                                                                            \\
    \iff & a\pp \leq b                                            &  & \by{i:2.2.12}[e]     \\
    \iff & b = (a\pp) + c \text{ for some natural number } c      &  & \by{i:2.2.11}        \\
    \iff & b = (a\pp) + c = (a + c)\pp = a + (c\pp)               &  & \by{i:2.2.1,i:2.2.3} \\
    \iff & b = a + d \text{ for some positive natural number } d. &  & \by{i:2.3,i:2.2.10}
  \end{align*}
\end{proof}

\begin{prop}[Trichotomy of order for natural numbers]\label{i:2.2.13}
  Let \(a\) and \(b\) be natural numbers.
  Then exactly one of the following statements is true: \(a < b\), \(a = b\), or \(a > b\).
\end{prop}

\begin{proof}[\pf{i:2.2.13}]
  First we show that we cannot have more than one of the statements \(a < b\), \(a = b\), \(a > b\) holding at the same time.
  If \(a < b\) then \(a \neq b\) by \cref{i:2.2.11}, and if \(a > b\) then \(a \neq b\) by \cref{i:2.2.11}.
  If \(a > b\) and \(a < b\) then by \cref{i:2.2.12}(c) we have \(a = b\), a contradiction.
  Thus no more than one of the statements is true.

  Now we show that at least one of the statements is true.
  We keep \(b\) fixed and induct on \(a\).
  When \(a = 0\), by \cref{i:2.2.1} we have \(b = 0 + b\) and thus by \cref{i:2.2.11} we have \(0 \leq b\) for any natural number \(b\).
  So we have either \(0 = b\) or \(0 < b\), which proves the base case.
  Now suppose we have proven the proposition for \(a\), and now we prove the proposition for \(a\pp\).
  From the trichotomy for \(a\), there are three cases: \(a < b\), \(a = b\), and \(a > b\).
  \begin{itemize}
    \item If \(a > b\), then by \cref{i:2.2.12}(d) we have \(a\pp \geq b\pp\).
          Since \(b\pp > b\), by \cref{i:2.2.12}(b) we have \(a\pp \geq b\).
          Then we must have \(a\pp > b\), otherwise by \cref{i:2.2.11} we would have \(a\pp = b\) and by \cref{i:2.2.12}(e) this implies \(a < b\) and contradicts to \(a > b\)
          (the contradiction is proved by the first paragraph).
    \item If \(a = b\), then by \cref{i:2.4} we have \(a\pp = b\pp\).
          Since \(b\pp > b\), by \cref{i:2.2.12}(b) we have \(a\pp \geq b\).
          Then we must have \(a\pp > b\), otherwise by \cref{i:2.2.11} we would have \(a\pp = b\) and by \cref{i:2.2.12}(e) this implies \(a < b\) and contradicts to \(a = b\)
          (the contradiction is proved by the first paragraph).
    \item If \(a < b\), then by \cref{i:2.2.12}(e) we have \(a\pp \leq b\).
          Thus either \(a\pp = b\) or \(a\pp < b\), and in either case we are done.
  \end{itemize}
  This closes the induction.
\end{proof}

\begin{prop}[Strong principle of induction]\label{i:2.2.14}
  Let \(m_0\) be a natural number, and let \(P(m)\) be a property pertaining to an arbitrary natural number \(m\).
  Suppose that for each \(m \geq m_0\), we have the following implication: if \(P(m')\) is true for all natural numbers \(m_0 \leq m' < m\), then \(P(m)\) is also true.
  (In particular, this means that \(P(m_0)\) is true, since in this case the hypothesis is vacuous.)
  Then we can conclude that \(P(m)\) is true for all natural numbers \(m \geq m_0\).
\end{prop}

\begin{proof}[\pf{i:2.2.14}]
  Let \(n\) be a natural number and let \(Q(n)\) be the statement ``\(P(m)\) is true for all natural number \(m\) satisfying \(m_0 \leq m < n\)''.
  We use induction to show that \(Q(n)\) is true for all natural number \(n\).

  For \(n = 0\), we want to show that \(Q(0)\) is true.
  However, we know that \(0 \leq m_0\) for all natural number \(m_0\).
  Thus, either \(0 = m_0\) or \(0 < m_0\) and so we split into cases.
  \begin{itemize}
    \item If \(0 < m_0\), the statement ``\(P(m)\) is true for all natural number \(m\) satisfying \(m_0 \leq m < n\)'' is vacuously true since there does not exist a natural number \(m\) satisfying \(0 < m_0 \leq m < n = 0\).
          Thus \(Q(0)\) is true in this case.
    \item If \(0 = m_0\), then the statement ``\(P(m)\) is true for all natural number \(m\) satisfying \(m_0 \leq m < n\)'' is vacuously true since there does not exist a natural number \(m\) satisfying \(0 = m_0 \leq m < n = 0\).
          Hence, \(Q(0)\) is true in this case.
  \end{itemize}
  From all cases above we see that \(Q(0)\) is true.
  Thus the base case holds.

  Suppose inductively that \(Q(n)\) is true for some natural number \(n\).
  We need to show that \(Q(n\pp)\) is true.
  Using induction hypothesis (\(Q(n)\) is true) and the hypothesis of \(P\) we see that \(P(n)\) is true.
  Since \(n < n\pp\), we know that \(P(m)\) is true for all natural number \(m\) satisfying \(m_0 \leq m \leq n < n\pp\).
  So \(P(m)\) is true for all natural number \(m\) satisfying \(m_0 \leq m < n\pp\) which in turn implies that \(Q(n\pp)\) is true.
  This closes the induction and hence we can conclude that \(Q(n)\) is true for any natural number \(n\).

  Since \(Q(n)\) is true for all natural number \(n\), by the hypothesis of \(P\) we know that \(P(n)\) is true for all natural number \(n\).
  In particular, we see that \(P(n)\) is true for all natural number \(n\) satisfying \(n \geq m_0\).
\end{proof}

\begin{rmk}\label{i:2.2.15}
  In applications we usually use \cref{i:2.2.14} with \(m_0 = 0\) or \(m_0 = 1\).
\end{rmk}

\exercisesection

\begin{ex}\label{i:ex:2.2.1}
  Prove \cref{i:2.2.5}.
\end{ex}

\begin{proof}[\pf{i:ex:2.2.1}]
  See \cref{i:2.2.5}.
\end{proof}

\begin{ex}\label{i:ex:2.2.2}
  Prove \cref{i:2.2.10}.
\end{ex}

\begin{proof}[\pf{i:ex:2.2.2}]
  See \cref{i:2.2.10}.
\end{proof}

\begin{ex}\label{i:ex:2.2.3}
  Prove \cref{i:2.2.12}.
\end{ex}

\begin{proof}[\pf{i:ex:2.2.3}]
  See \cref{i:2.2.12}.
\end{proof}

\begin{ex}\label{i:ex:2.2.4}
  Justify the three statements marked in the proof of \cref{i:2.2.13}.
\end{ex}

\begin{proof}[\pf{i:ex:2.2.4}]
  See \cref{i:2.2.13}.
\end{proof}

\begin{ex}\label{i:ex:2.2.5}
  Prove \cref{i:2.2.14}.
\end{ex}

\begin{proof}[\pf{i:ex:2.2.5}]
  See \cref{i:2.2.14}.
\end{proof}

\begin{ex}[Principle of backwards induction]\label{i:ex:2.2.6}
  Let \(n\) be a natural number, and let \(P(m)\) be a property pertaining to the natural numbers such that whenever \(P(m\pp)\) is true, then \(P(m)\) is true.
  Suppose that \(P(n)\) is also true.
  Prove that \(P(m)\) is true for all natural numbers \(m \leq n\);
\end{ex}

\begin{proof}[\pf{i:ex:2.2.6}]
  We use induction on \(n\).
  For \(n = 0\), the only natural number \(m\) satisfying \(m \leq n = 0\) is \(0\).
  By hypothesis \(P(0)\) is true.
  Therefore, the base case holds trivially.

  Suppose inductively that for some natural number \(n\) we have the implication ``\(P(n)\) is true implies \(P(m)\) is true for all natural number \(m\) satisfying \(m \leq n\)''.
  We want to show the implication ``\(P(n\pp)\) is true implies \(P(m)\) is true for all natural number \(m\) satisfying \(m \leq n\pp\)''.
  But when \(P(n\pp)\) is true, by the hypothesis of \(P\) we know that \(P(n)\) is true.
  Thus we can apply induction hypothesis to derive ``\(P(m)\) is true for all natural number \(m\) satisfying \(m \leq n\)''.
  Combining the statement ``\(P(n\pp)\) is true'' we see that the statement ``\(P(m)\) is true for all natural number \(m\) satisfying \(m \leq n\pp\)'' is true.
  This closes the induction.
\end{proof}

\begin{ex}\label{i:ex:2.2.7}
  Let \(n\) be a natural number, and let \(P(m)\) be a property pertaining to the natural numbers such that whenever \(P(m)\) is true, \(P(m\pp)\) is true.
  Show that if \(P(n)\) is true, then \(P(m)\) is true for any natural number \(m\) satisfying \(m \geq n\).
  This principle is sometimes referred to as \emph{the principle of induction starting from the base case \(n\)}.
\end{ex}

\begin{proof}[\pf{i:ex:2.2.7}]
  Suppose that \(P(n)\) is true.
  Let \(Q(k) = P(n + k)\) for every natural number \(k\).
  We use induction to show that \(Q(k)\) is true for every natural number \(k\).

  For \(k = 0\), by \cref{i:2.2.2} we have \(Q(0) = P(n + 0) = P(n)\).
  Since \(P(n)\) is true by hypothesis, we know that \(Q(0)\) is true.
  Thus the base case holds.

  Suppose inductively that \(Q(k)\) is true for some natural number \(k\).
  We want to show that \(Q(k\pp)\) is true.
  By \cref{i:2.2.3} we have \(Q(k\pp) = P(n + (k\pp)) = P((n + k)\pp)\).
  By induction hypothesis we have \(Q(k) = P(n + k)\) is true.
  Thus we can use the hypothesis of \(P\) to show that \(P((n + k)\pp)\) is also true.
  This closes the induction.
  Thus we conclude that \(P(n + k)\) is true for every natural number \(k\).

  By \cref{i:2.2.11} we know that for every natural number \(m\), we have \(m \geq n \iff m = n + k\) for some natural number \(k\).
  Thus we see that \(P(m)\) is true for every natural number \(m\) satisfying \(m \geq n\).
\end{proof}
