\section{The Peano axioms}

\begin{note}
We now present one standard way to define the natural numbers, in terms of the \emph{Peano axioms}, which were first laid out by Giuseppe Peano (1858–1932).
This is not the only way to define the natural numbers.
For instance, another approach is to talk about the cardinality of finite sets, for instance one could take a set of five elements and define \(5\) to be the number of elements in that set.
\end{note}

\begin{note}
In some texts the natural numbers start at \(1\) instead of \(0\), but this is a matter of notational convention more than anything else.
In this text we shall refer to the set \(\{1, 2, 3,...\}\) as the positive integers \(\mathds{Z}^+\) rather than the natural numbers.
Natural numbers are sometimes also known as \emph{whole numbers}.
\end{note}

\begin{note}
In mathematics we try not to define a variable more than once in any given setting, as it can often lead to confusion;
many of the statements which were true for the old value of the variable can now become false, and vice versa.
\end{note}

\begin{axiom}\label{2.1}
\(0\) is a natural number.
\end{axiom}

\begin{axiom}\label{2.2}
If \(n\) is a natural number, then \(n++\) is also a natural number.
\end{axiom}

\begin{axiom}\label{2.3}
\(0\) is not the successor of any natural number;
i.e., we have \(n++ \neq 0\) for every natural number \(n\).
\end{axiom}

\begin{axiom}\label{2.4}
Different natural numbers must have different successors;
i.e., if \(n\), \(m\) are natural numbers and \(n \neq m\), then \(n++ \neq m++\).
Equivalently, if \(n++ = m++\), then we must have \(n = m\).
\end{axiom}

\begin{axiom}[Principle of mathematical induction]\label{2.5}
Let \(P(n)\) be any property pertaining to a natural number \(n\).
Suppose that \(P(0)\) is true, and suppose that whenever \(P(n)\) is true, \(P(n++)\) is also true.
Then \(P(n)\) is true for every natural number \(n\).
\end{axiom}

\begin{note}
Axioms \ref{2.1}-\ref{2.5} are known as the Peano axioms for the natural numbers.
\end{note}

\section{Addition}

\begin{definition}[Addition of natural numbers]\label{2.2.1}
Let \(m\) be a natural numbers.
To add zero to \(m\), we define \(0+m \coloneqq m\).
Now suppose inductively that we have defined how to add \(n\) to \(m\).
Then we can add \(n++\) to \(m\) by defining \((n++) + m \coloneqq (n + m)++\).
\end{definition}

\begin{additional corollary}\label{ac 2.2.1}
The sum of two natural numbers is again a natural number.
\end{additional corollary}

\begin{proof}
We use induction.
Let \(m\) be a natural number.
\(0 + m = m\) is a natural number according to the given condition.
Now suppose inductively that \(n\) is a natural number such that \(n + m\) is a natural number.
We wish to show that \((n++) + m\) is also a natural number.
But by Definition \ref{2.2.1} and Axiom \ref{2.2},
\((n++) + m = (n + m)++\) is a natural number.
This close the induction.
\end{proof}

\begin{lemma}\label{2.2.2}
For any natural number \(n\), \(n + 0 = n\).
\end{lemma}

\begin{proof}
We use induction.
The base case \(0 + 0 = 0\) follows since we know that \(0 + m = m\) for every natural number \(m\), and \(0\) is a natural number.
Now suppose inductively that \(n + 0 = n\).
We wish to show that \((n++) + 0 = n++\).
But by Definition \ref{2.2.1}, \((n++) + 0\) is equal to \((n + 0)++\), which is equal to \(n++\) since \(n + 0 = n\).
This closes the induction.
\end{proof}

\begin{lemma}\label{2.2.3}
For any natural numbers \(n\) and \(m\), \(n + (m++) = (n + m)++\).
\end{lemma}

\begin{proof}
We induct on \(n\) (keeping \(m\) fixed).
We first consider the base case \(n = 0\).
In this case we have to prove \(0 + (m++) = (0 + m)++\).
But by Definition \ref{2.2.1}, \(0 + (m++) = m++\) and \(0 + m = m\), so both sides are equal to \(m++\) and are thus equal to each other.
Now we assume inductively that \(n + (m++) = (n + m)++\);
we now have to show that \((n++) + (m++) = ((n++) + m)++\).
The left-hand side is \((n + (m++))++\) by Definition \ref{2.2.1}, which is equal to \(((n+m)++)++\) by the inductive hypothesis.
Similarly, we have \((n++) + m = (n + m)++\) by the Definition \ref{2.2.1}, and so the right-hand side is also equal to \(((n + m)++)++\).
Thus both sides are equal to each other, and we have closed the induction.
\end{proof}

\begin{additional corollary}\label{ac 2.2.2}
\(n++ = n + 1\).
\end{additional corollary}

\begin{proof}
\begin{align*}
n++ &= (n++) + 0 & \text{(by Lemma \ref{2.2.1})} \\
&= (n + 0)++ & \text{(by Definition \ref{2.2.1})} \\
&= n + (0++) & \text{(by Lemma \ref{2.2.3})} \\
&= n + 1.
\end{align*}
\end{proof}

\begin{proposition}[Addition is commutative]\label{2.2.4}
For any natural numbers \(n\) and \(m\), \(n + m = m + n\).
\end{proposition}

\begin{proof}
We shall use induction on \(n\) (keeping \(m\) fixed).
First we do the base case \(n = 0\), i.e., we show \(0 + m = m + 0\).
By the Definition \ref{2.2.1}, \(0 + m = m\), while by Lemma \ref{2.2.1}, \(m + 0 = m\).
Thus the base case is done.
Now suppose inductively that \(n + m = m + n\), now we have to prove that \((n++) + m = m + (n++)\) to close the induction.
By the Definition \ref{2.2.1}, \((n++) + m = (n + m)++\).
By Lemma \ref{2.2.3}, \(m + (n++) = (m + n)++\), but this is equal to \((n + m)++\) by the inductive hypothesis \(n+m=m+n\).
Thus \((n++) + m = m + (n++)\) and we have closed the induction.
\end{proof}

\begin{proposition}[Addition is associative]\label{2.2.5}
For any natural numbers \(a\), \(b\), \(c\), we have \((a + b) + c = a + (b + c)\).
\end{proposition}

\begin{proof}
We shall use induction on \(c\) and keeping both \(a\) and \(b\) fixed.
First we do the base case \(c = 0\), i.e., we show \((a + b) + 0 = a + (b + 0)\).
By Lemma \ref{2.2.1}, \((a + b) + 0 = a + b\) and \(a + (b + 0) = a + b\).
Thus the base case is done.
Now suppose inductively that \((a + b) + c = a + (b + c)\), we have to prove that \((a + b) + (c++) = a + (b + (c++))\) to close the induction.
By Lemma \ref{2.2.3}, \((a + b) + (c++) = ((a + b) + c)++\).
Also By Lemma \ref{2.2.3}, \(a + (b + (c++)) = a + ((b + c)++) = (a + (b + c))++\), but this is equal to \(((a + b) + c)++\) by the inductive hypothesis \((a + b) + c = a + (b + c)\).
Thus \((a + b) + (c++) = a + (b + (c++))\) and we have closed the induction.
\end{proof}

\begin{note}
Because of this associativity we can write sums such as \(a + b + c\) without having to worry about which order the numbers are being added together.
\end{note}

\begin{proposition}[Cancellation law]\label{2.2.6}
Let \(a\), \(b\), \(c\) be natural numbers such that \(a + b = a + c\).
Then we have \(b = c\).
\end{proposition}

\begin{proof}
We prove this by induction on \(a\).
First consider the base case \(a = 0\).
Then we have \(0 + b = 0 + c\), which by Definition \ref{2.2.1} implies that \(b = c\) as desired.
Now suppose inductively that we have the cancellation law for \(a\) (so that \(a + b = a + c\) implies \(b = c\));
we now have to prove the cancellation law for \(a++\).
In other words, we assume that \((a++) + b = (a++) + c\) and need to show that \(b = c\).
By the Definition \ref{2.2.1}, \((a++) + b = (a + b)++\) and \((a++) + c = (a + c)++\) and so we have \((a + b)++ = (a + c)++\).
By Axiom \ref{2.4}, we have \(a + b = a + c\).
Since we already have the cancellation law for \(a\), we thus have \(b = c\) as desired.
This closes the induction.
\end{proof}

\begin{definition}[Positive natural numbers]\label{2.2.7}
A natural number \(n\) is said to be \emph{positive} iff it is not equal to \(0\).
\end{definition}

\begin{proposition}\label{2.2.8}
If \(a\) is positive and \(b\) is a natural number, then \(a + b\) is positive (and hence \(b + a\) is also, by Proposition \ref{2.2.4}).
\end{proposition}

\begin{proof}
We use induction on \(b\).
If \(b = 0\), then \(a + b = a + 0 = a\), which is positive, so this proves the base case.
Now suppose inductively that \(a + b\) is positive.
Then \(a + (b++) = (a + b)++\), which cannot be zero by Axiom \ref{2.3}, and is hence positive.
This closes the induction.
\end{proof}

\begin{corollary}\label{2.2.9}
If \(a\) and \(b\) are natural numbers such that \(a + b = 0\), then \(a = 0\) and \(b = 0\).
\end{corollary}

\begin{proof}
Suppose for sake of contradiction that \(a \neq 0\) or \(b \neq 0\).
If \(a \neq 0\) then \(a\) is positive, and hence \(a + b = 0\) is positive by Proposition \ref{2.2.8}, a contradiction.
Similarly if \(b \neq 0\) then \(b\) is positive, and again \(a + b = 0\) is positive by Proposition \ref{2.2.8}, a contradiction.
Thus \(a\) and \(b\) must both be zero.
\end{proof}

\begin{lemma}\label{2.2.10}
Let \(a\) be a positive number.
Then there exists exactly one natural number \(b\) such that \(b++ = a\).
\end{lemma}

\begin{proof}
We use induction.
If \(a = 1\) and \(b++ = a\), then \(b = 0\).
We want to show that \(0\) is unique, so assume that \(0'\) is a natural number and \(0'++ = 1\).
By Axiom \ref{2.4}, if \(0++ = 0'++\), then \(0 = 0'\), thus \(0\) is unique, and this proves the base case.
Suppose inductively that \(a\) is positive, there exist exactly one natural number \(b\) such that \(b++ = a\).
Then \(a++ = (b++)++\) by induction hypothesis, so there exist a natural number \(b++\) such that \((b++)++ = a++\).
Now we need to show that \(b++\) is unique.
Assume that there exist another natural number \(c\) such that \(a++ = c++\).
Then by Axiom \ref{2.4}, \(a = c\), and by induction hypothesis, \(c = b++\), so \(c++ = (b++)++\), which means \(b++\) is unique.
This close the induction.
\end{proof}

\begin{definition}[Ordering of the natural numbers]\label{2.2.11}
Let \(n\) and \(m\) be natural numbers.
We say that \(n\) is greater than or equal to \(m\), and write \(n \geq m\) or \(m \leq n\), iff we have \(n = m + a\) for some natural number \(a\).
We say that \(n\) is strictly greater than \(m\), and write \(n > m\) or \(m < n\), iff \(n \geq m\) and \(n \neq m\).
\end{definition}

\begin{note}
There is no largest natural number \(n\), because the next number \(n++\) is always larger.
\end{note}

\begin{proposition}[Basic properties of order for natural numbers]\label{2.2.12}
Let \(a\), \(b\), \(c\) be natural numbers.
Then
\begin{enumerate}
\item (Order is reflexive) \(a \geq a\).
\item (Order is transitive) If \(a \geq b\) and \(b \geq c\), then \(a \geq c\).
\item (Order is anti-symmetric) If \(a \geq b\) and \(b \geq a\), then \(a = b\).
\item (Addition preserves order) \(a \geq b\) iff \(a + c \geq b + c\).
\item \(a < b\) iff \(a++ \leq b\).
\item \(a < b\) iff \(b = a + d\) for some \emph{positive} number \(d\).
\end{enumerate}
\end{proposition}

\begin{proof}{(a)}
\(\because a = a + 0\), \(\therefore a \geq a\) by Definition \ref{2.2.11}.
\end{proof}

\begin{proof}{(b)}
Let \(a = b + d\) and \(b = c + e\), where \(d\), \(e\) are natural numbers.
Substitute \(b\) with equation, we can derive \(a = b + d = (c + e) + d = c + (e + d)\).
By Additional Corollary \ref{ac 2.2.1}, \(e + d\) is a natural number.
Therefore by Definition \ref{2.2.11}, \(a \geq c\).
\end{proof}

\begin{proof}{(c)}
By the given conditions \(a \geq b\) and \(b \geq a\), let \(a = b + c\) and \(b = a + d\), where \(c\) and \(d\) are natural numbers.
Substitute \(b\) with equation, we can derive that \(a = b + c = (a + d) + c = a + (d + c)\).
By Proposition \ref{2.2.6}, \(a = a + (d + c) \implies 0 = d + c\), and by Corollary \ref{2.2.9}, \(d = c = 0\).
Therefore \(a = b + 0 = b\) and \(b = a + 0 = a\).
\end{proof}

\begin{proof}{(d)}
First we prove the necessary condition of (d).
Let \(a = b + d\), where \(d\) is a natural number.
Then \(a = b + d \implies (a + c) = (b + c) + d\), which means \(a + c \geq b + c\) by Definition \ref{2.2.11}.

Now we prove the sufficient condition of (d).
Let \(a + c = b + c + d\), where \(d\) is a natural number.
Then by Proposition \ref{2.2.6}, \(a + c = b + c + d \implies a = b + d\), which means \(a \geq b\) by Definition \ref{2.2.11}.
\end{proof}

\begin{proof}{(e)}
First we prove the necessary condition of (e).
Let \(b = a + c\), where \(c\) is a natural number.
Because \(a \neq b\), so \(c \neq 0\).
By Lemma \ref{2.2.10}, there exists a natural number \(d\) such that \(d++ = c\).
So \(b = a + c = a + (d++) = (a + d)++ = (a++) + d\) by Lemma \ref{2.2.3} and Definition \ref{2.2.1}.
Thus \(a++ \leq b\) by Definition \ref{2.2.11}.

Now we prove the sufficient condition of (e).
Let \(b = (a++) + c\), where \(c\) is a natural number.
Then \(b = a + (c++)\) and \(c++\) must not be zero, otherwise contradict to Axiom \ref{2.3}.
So \(b \neq a\), thus \(a < b\).
\end{proof}

\begin{proof}{(f)}
First we prove the necessary condition of (f).
Let \(b = a + d\), where \(d\) is a natural number.
Because \(a \neq b\), so \(d \neq 0\), which means \(d\) is positive by Definition \ref{2.2.7}.

Now we prove the sufficient condition of (f).
Let \(b = a + d\), where \(d\) is a positive natural number.
Because \(d \neq 0\), \(b \neq a\), thus \(a < b\).
\end{proof}

\begin{proposition}[Trichotomy of order for natural numbers]\label{2.2.13}
Let \(a\) and \(b\) be natural numbers.
Then exactly one of the following statements is true: \(a < b\), \(a = b\), or \(a > b\).
\end{proposition}

\begin{proof}
First we show that we cannot have more than one of the statements \(a < b\), \(a = b\), \(a > b\) holding at the same time.
If \(a < b\) then \(a \neq b\) by Definition \ref{2.2.11}, and if \(a > b\) then \(a \neq b\) by Definition \ref{2.2.11}.
If \(a > b\) and \(a < b\) then by Proposition \ref{2.2.12} we have \(a = b\), a contradiction.
Thus no more than one of the statements is true.
Now we show that at least one of the statements is true.
We keep \(b\) fixed and induct on \(a\).
When \(a = 0\) we have \(0 \leq b\) for all \(b\) because \(b = b + 0\), so we have either \(0 = b\) or \(0 < b\), which proves the base case.
Now suppose we have proven the proposition for \(a\), and now we prove the proposition for \(a++\).
From the trichotomy for \(a\), there are three cases: \(a < b\), \(a = b\), and \(a > b\).
If \(a > b\), then \(a++ > b\) because \(a++ > a > b\) and Proposition \ref{2.2.12}.
If \(a = b\), then \(a++ > b\) because \(a++ = b++ > b\).
Now suppose that \(a < b\).
Then by Proposition \ref{2.2.12}, we have \(a++ \leq b\).
Thus either \(a++ = b\) or \(a++ < b\), and in either case we are done.
This closes the induction.
\end{proof}

\begin{proposition}[Strong principle of induction]\label{2.2.14}
Let \(m_0\) be a natural number, and let \(P(m)\) be a property pertaining to an arbitrary natural number \(m\).
Suppose that for each \(m \geq m_0\), we have the following implication: if \(P(m')\) is true for all natural numbers \(m_0 \leq m' < m\), then \(P(m)\) is also true.
(In particular, this means that \(P(m_0)\) is true, since in this case the hypothesis is vacuous.)
Then we can conclude that \(P(m)\) is true for all natural numbers \(m \geq m_0\).
\end{proposition}

\begin{proof}
Let \(n\) be a natural number and let \(Q(n)\) be the property that \(P(m)\) is true for all \(m_0 \leq m < n\) for \(n \geq m_0\).
Using induction on \(n\), for the base case \(n = 0\), we want to show that \(Q(0)\) is true.
However, we know that \(0 \leq m_0\) for all natural number \(m_0\).
Thus, either \(0 = m_0\) or \(0 < m_0\) and so we split into cases.
If \(n = 0 < m_0\), the statement \(P(m) \ \forall\ m_0 \leq m < n\) is vacuously true (since the hypothesis applies for \(n \geq m_0\)) and thus \(Q(0)\) is true in this case.
For the second case, if \(n = 0 = m_0\), then the statement \(P(m) \ \forall\ m_0 \leq m < n\) is also vacuously true since there is no natural number \(m'\) such that \(0 \leq m' < 0\). Hence, \(Q(0)\) is true for this case and that completes the base case of the induction.

Now suppose inductively that for some \(n \geq m_0\), \(Q(n)\) is true, i.e \(P(m) \ \forall\ m_0 \leq m < n\) is true.
We need to show that \(Q(n++)\) is true.
By the definition of \(P\) in the hypothesis, \(P(n)\) is also true (because \(Q(n)\) is true).
Since \(n < n++\), then \(P(m) \ \forall\ m_0 \leq m \leq n < n++\) is true so \(P(m) \ \forall\ m_0 \leq m < n++\) is true which in turn implies that \(Q(n++)\) is true.
Which closes the induction and hence we can conclude that \(Q(n) \ \forall\ n\) is true.
However, \(Q(n)\) true implies \(P(m) \ \forall\ m_0 \leq m < n\) is true for all \(n \geq m_0\) and by the definition of \(P\), \(P(n)\) is also true for all \(n \geq m_0\) which concludes the proof.
\end{proof}

\begin{remark}\label{2.2.15}
In applications we usually use Proposition \ref{2.2.14} with \(m_0 = 0\) or \(m_0 = 1\).
\end{remark}

\exercisesection

\begin{exercise}\label{ex 2.2.1}
Prove Proposition \ref{2.2.5}.
\end{exercise}

\begin{proof}
See Proposition \ref{2.2.5}.
\end{proof}

\begin{exercise}\label{ex 2.2.2}
Prove Lemma \ref{2.2.10}.
\end{exercise}

\begin{proof}
See Lemma \ref{2.2.10}.
\end{proof}

\begin{exercise}\label{ex 2.2.3}
Prove Proposition \ref{2.2.12}.
\end{exercise}

\begin{proof}
See Proposition \ref{2.2.12}.
\end{proof}

\begin{exercise}\label{ex 2.2.4}
Justify the three statements marked in the proof of Proposition \ref{2.2.13}.
\end{exercise}

\begin{proof}
See Proposition \ref{2.2.13}.
\end{proof}

\begin{exercise}\label{ex 2.2.5}
Prove Proposition \ref{2.2.14}.
\end{exercise}

\begin{proof}
See Proposition \ref{2.2.14}.
\end{proof}

\begin{exercise}[Principle of backwards induction]\label{ex 2.2.6}
Let \(n\) be a natural number, and let \(P(m)\) be a property pertaining to the natural numbers such that whenever \(P(m++)\) is true, then \(P(m)\) is true.
Suppose that \(P(n)\) is also true.
Prove that \(P(m)\) is true for all natural numbers \(m \leq n\);
\end{exercise}

\begin{proof}
We use induction.
The base case \(n = 0\) is trivially true because \(\forall\ m \leq 0\) means \(m = 0\), thus \(P(m) = P(n) = P(0)\) is true.
Suppose inductively that \(P(n)\) is true, and \(P(m)\) is true \(\forall\ m \leq n\).
Then when \(P(n++)\) is true, by the given condition \(P(n)\) is true.
And by induction hypothesis \(P(m)\) is true \(\forall\ m \leq n \leq n++\), so \(P(n++)\) is true implies \(P(m)\) is true \(\forall\ m \leq n++\).
This close the induction.
\end{proof}

\section{Multiplication}
In the previous section we have proven all the basic facts that we know to be true about addition and order.
To save space and to avoid belaboring the obvious, we will now allow ourselves to use all the rules of algebra concerning addition and order that we are familiar with, without further comment.

\begin{definition}[Multiplication of natural numbers]\label{2.3.1}
Let \(m\) be a natural number.
To multiply zero to \(m\), we define \(0 \times m \coloneqq 0\).
Now suppose inductively that we have defined how to multiply \(n\) to \(m\).
Then we can multiply \(n++\) to \(m\) by defining \((n++) \times m \coloneqq (n \times m) + m\).
\end{definition}

\begin{additional corollary}\label{ac 2.3.1}
The product of two natural numbers is a natural number.
\end{additional corollary}

\begin{proof}
Let \(n\), \(m\) be two natural numbers.
We use induction on \(n\).
For \(n = 0\), \(n \times m = 0 \times m = 0\) is a natural number.
Now suppose inductively that \(n\) is a natural number such that \(n \times m\) is a natural number.
We wish to show that \((n++) \times m\) is also a natural number.
By Definition \ref{2.3.1}, \((n++) \times m = (n \times m) + m\).
By induction hypothesis and by Additional Corollary \ref{ac 2.2.1}, \((n \times m)\) is a natural number and sum of two natural number is again a natural number.
This close the induction.
\end{proof}

\begin{additional corollary}\label{ac 2.3.2}
Let \(n\) be a natural number.
Then \(n \times 0 = 0\).
\end{additional corollary}

\begin{proof}
We use induction.
For base case \(n = 0\), \(0 \times 0 = 0\) by Definition \ref{2.3.1}.
Suppose inductively that \(n \times 0 = 0\).
Then \((n++) \times 0 = (n \times 0) + 0 = 0 + 0 = 0\) by Definition \ref{2.3.1} and induction hypothesis.
This close the induction.
\end{proof}

\begin{additional corollary}\label{ac 2.3.3}
Let \(n\), \(m\) be natural numbers.
Then \(n \times (m++) = (n \times m) + n\).
\end{additional corollary}

\begin{proof}
We use induction on \(n\) (fixed \(m\).
For base case \(n = 0\), \(0 \times (m++) = 0\) by Definition \ref{2.3.1}.
Suppose inductively that \(n \times (m++) = (n \times m) + n\).
Then
    \begin{align*}
        (n++) \times (m++)
        &= (n \times (m++)) + (m++) & \text{(by Definition \ref{2.3.1})} \\
        &= (n \times m) + n + (m++) & \text{(by induction hypothesis)} \\
        &= (n \times m) + (n++) + m \\
        &= (n \times m) + m + (n++) \\
        &= ((n++) \times m) + (n++). & \text{(by Definition \ref{2.3.1})}
    \end{align*}
This close the induction.
\end{proof}

\begin{lemma}[Multiplication is commutative]\label{2.3.2}
Let \(n\), \(m\) be natural numbers.
Then \(n \times m = m \times n\)
\end{lemma}

\begin{proof}
We use induction on \(n\) (fixed \(m\)).
For base case \(n = 0\), we want to prove that \(0 \times m = m \times 0\).
By Definition \ref{2.3.1}, \(0 \times m = 0\), and by Lemma \ref{ac 2.3.2}, \(m \times 0 = 0\), thus base case is true.
Suppose inductively that \(n \times m = m \times n\) is true.
Then we want to show that \((n++) \times m = m \times (n++)\) is also true.
By Definition \ref{2.3.1}, \((n++) \times m = (n \times m) + m\).
By Additional Corollary \ref{ac 2.3.3} and induction hypothesis, \(m \times (n++) = (m \times n) + m = (n \times m) + m\).
Thus \((n++) \times m = m \times (n++)\), and this close the induction.
\end{proof}

\begin{note}
We will now abbreviate \(n \times m\) as \(nm\), and use the usual convention that multiplication takes precedence over addition, thus for instance \(ab + c\) means \((a \times b) + c\), not \(a \times (b + c)\).
\end{note}

\begin{lemma}[Positive natural numbers have no zero divisors]\label{2.3.3}
Let \(n\), \(m\) be natural numbers.
Then \(n \times m = 0\) if and only if at least one of \(n\), \(m\) is equal to zero.
In particular, if \(n\) and \(m\) are both positive, then \(nm\) is also positive.
\end{lemma}

\begin{proof}
We first prove the necessary condition of Lemma \ref{2.3.3}.
Suppose for sake of contradiction that \(n\) and \(m\) are positive, and \(n = a++\), \(m = b++\) for some natural number \(a\), \(b\)
Then \(n \times m = (a++) \times (b++) = (a \times (b++)) + (b++) = (a \times b) + a + (b++)\) by Definition \ref{2.3.1} and Lemma \ref{ac 2.3.3}.
But \(b++\) is positive, so \(n \times m\) is positive, a contradiction with \(n \times m = 0\).
Thus at least one of \(n\), \(m\) is \(0\).

Now we prove the sufficient condition of Lemma \ref{2.3.3}.
By Definition \ref{2.3.1}, \(n \times m = 0\) if \(n = 0\), and by Additional Corollary \ref{ac 2.3.2}, \(n \times m = 0\) if \(m = 0\).
\end{proof}

\begin{proposition}[Distributive law]\label{2.3.4}
For any natural numbers \(a\), \(b\), \(c\), we have \(a(b + c) = ab + ac\) and \((b + c)a = ba + ca\).
\end{proposition}

\begin{proof}
Since multiplication is commutative we only need to show the first identity \(a(b + c) = ab + ac\).
We keep \(a\) and \(b\) fixed, and use induction on \(c\).
Let’s prove the base case \(c = 0\), i.e., \(a(b + 0) = ab + a0\).
The left-hand side is \(ab\), while the right-hand side is \(ab + 0 = ab\), so we are done with the base case.
Now let us suppose inductively that \(a(b + c) = ab + ac\), and let us prove that \(a(b + (c++)) = ab + a(c++)\).
The left-hand side is \(a((b + c)++) = a(b + c) + a\), while the right-hand side is \(ab + ac + a = a(b + c) + a\) by the induction hypothesis, and so we can close the induction.
\end{proof}

\begin{proposition}[Multiplication is associative]\label{2.3.5}
For any natural numbers \(a\), \(b\), \(c\), we have \((a \times b) \times c = a \times (b \times c)\).
\end{proposition}

\begin{proof}
We keep \(a\) and \(b\) fixed, and use induction on \(c\).
The base case \(c = 0\), i.e., \((a \times b) \times 0 = a \times (b \times 0)\).
The left-hand side is \(0\), while the right-hand side is \(a \times 0 = 0\), so we are done with the base case.
Now let us suppose inductively that \((a \times b) \times c = a \times (b \times c)\), and let us prove that \((a \times b) \times (c++) = a \times (b \times (c++))\).
The left-hand side is \((a \times b) \times c + (a \times b)\), while the right-hand side is \(a \times (b \times c + b) = a \times (b \times c) + a \times b = (a \times b) \times c + a \times b\) by the Proposition \ref{2.3.4} and induction hypothesis, and so we can close the induction.
\end{proof}

\begin{proposition}[Multiplication preserves order]\label{2.3.6}
If \(a\), \(b\) are natural numbers such that \(a < b\), and \(c\) is positive, then \(ac < bc\).
\end{proposition}

\begin{proof}
Since \(a < b\), we have \(b = a + d\) for some positive \(d\).
Multiplying by \(c\) and using the distributive law we obtain \(bc = ac + dc\).
Since \(d\) is positive, and \(c\) is positive, \(dc\) is positive, and hence \(ac < bc\) as desired.
\end{proof}

\begin{corollary}[Cancellation law]\label{2.3.7}
Let \(a\), \(b\), \(c\) be natural numbers such that \(ac = bc\) and \(c\) is non-zero.
Then \(a = b\).
\end{corollary}

\begin{proof}
By the trichotomy of order, we have three cases: \(a < b\), \(a = b\), \(a > b\).
Suppose first that \(a < b\), then by Proposition \ref{2.3.6} we have \(ac < bc\), a contradiction.
We can obtain a similar contradiction when \(a > b\).
Thus the only possibility is that \(a = b\), as desired.
\end{proof}

\begin{remark}\label{2.3.8}
Just as Proposition \ref{2.2.6} will allow for a ``virtual subtraction'' which will eventually let us define genuine subtraction, Corollary \ref{2.3.7} provides a ``virtual division'' which will be needed to define genuine division later on.
\end{remark}

\begin{proposition}[Euclidean algorithm]\label{2.3.9}
Let \(n\) be a natural number, and let \(q\) be a positive number.
Then there exist natural numbers \(m\), \(r\) such that \(0 \leq r<q\) and \(n = mq + r\).
\end{proposition}

\begin{proof}
We use induction on \(n\) (fixed \(q\)).
For the base case \(n = 0\), let \(r = m = 0\), then \(0 = 0q + 0\), \(0 \leq 0 < q\), so we are done with base case.
Suppose inductively that \(n = mq + r\), \(0 \leq r < q\).
Then \(n++ = (mq + r)++ = mq + (r++)\).
By induction hypothesis, because \(r < q\), so \(r++ \leq q\) by Proposition \ref{2.2.12}.
By Proposition \ref{2.2.13}, we have two cases: \(r++ < q\), \(r++ = q\).
If \(r++ < q\), then \(n++ = mq + (r++)\) satisfied the condition.
If \(r++ = q\), then \(n++ = mq + q = (m++) \times q\), also satisfied the condition.
This close the induction.
\end{proof}

\begin{remark}\label{2.3.10}
In other words, we can divide a natural number \(n\) by a positive number \(q\) to obtain a quotient \(m\) (which is another natural number) and a remainder \(r\) (which is less than \(q\)).
This algorithm marks the beginning of \emph{number theory}, which is a beautiful and important subject but one which is beyond the scope of this text.
\end{remark}

\begin{definition}[Exponentiation for natural numbers]\label{2.3.11}
Let \(m\) be a natural number.
To raise \(m\) to the power \(0\), we define \(m^0 \coloneqq 1\); in particular, we define \(0^0 \coloneqq 1\).
Now suppose recursively that \(m^n\) has been defined for some natural number \(n\), then we define \(m^{n++} \coloneqq m^n \times m\).
\end{definition}

\exercisesection

\begin{exercise}\label{ex 2.3.1}
Prove Lemma \ref{2.3.2}.
\end{exercise}

\begin{proof}
See Lemma \ref{2.3.2}
\end{proof}

\begin{exercise}\label{ex 2.3.2}
Prove Lemma \ref{2.3.3}
\end{exercise}

\begin{proof}
See Lemma \ref{2.3.3}
\end{proof}

\begin{exercise}\label{ex 2.3.3}
Prove Proposition \ref{2.3.5}
\end{exercise}

\begin{proof}
See Proposition \ref{2.3.5}
\end{proof}

\begin{exercise}\label{ex 2.3.4}
Prove the identity \((a + b)^2 = a^2 + 2ab + b^2\) for all natural numbers \(a\), \(b\).
\end{exercise}

\begin{proof}
\begin{align*}
    (a + b)^2 &= (a + b)^1 \times (a + b) & \text{(By Definition \ref{2.3.11})} \\
    &= (a + b)^0 \times (a + b) \times (a + b) & \text{(By Definition \ref{2.3.11})} \\
    &= 1 \times (a + b) \times (a + b) & \text{(By Definition \ref{2.3.11})} \\
    &= (1 \times (a + b)) \times (a + b) & \text{(By Proposition \ref{2.3.5})} \\
    &= (1 \times a + 1 \times b)) \times (a + b) & \text{(By Proposition \ref{2.3.4})} \\
    &= ((0 \times a + a) + (0 \times b + b)) \times (a + b) & \text{(By Definition \ref{2.3.1})} \\
    &= (a + b) \times (a + b) & \text{(By Definition \ref{2.3.1})} \\
    &= a(a + b) + b(a + b) & \text{(By Proposition \ref{2.3.5})} \\
    &= aa + ab + ba + bb & \text{(By Proposition \ref{2.3.5})} \\
    &= a^2 + ab + ba + b^2 & \text{(By Definition \ref{2.3.11})} \\
    &= a^2 + ab + ab + b^2 & \text{(By Lemma \ref{2.3.2})} \\
    &= a^2 + (0 + ab) + ab + b^2 & \text{(By Definition \ref{2.2.1})} \\
    &= a^2 + 0ab + ab + ab + b^2 & \text{(By Definition \ref{2.3.1})} \\
    &= a^2 + 1ab + ab + b^2 & \text{(By Lemma \ref{2.3.2})} \\
    &= a^2 + 2ab + b^2 & \text{(By Definition \ref{2.3.1})} \\
\end{align*}
\end{proof}

\begin{exercise}\label{ex 2.3.5}
Prove Proposition \ref{2.3.9}
\end{exercise}

\begin{proof}
See Proposition \ref{2.3.9}
\end{proof}