\section{Absolute value and exponentiation}

\begin{definition}[Absolute value]\label{4.3.1}
If \(x\) is a rational number, the \emph{absolute value} \(|x|\) of \(x\) is defined as follows.
If \(x\) is positive, then \(|x| \coloneqq x\).
If \(x\) is negative, then \(|x| \coloneqq -x\).
If \(x\) is zero, then \(|x| \coloneqq 0\).
\end{definition}

\begin{definition}[Distance]\label{4.3.2}
Let \(x\) and \(y\) be rational numbers.
The quantity \(|x - y|\) is called the \emph{distance between \(x\) and \(y\)} and is sometimes denoted \(d(x, y)\), thus \(d(x, y) \coloneqq |x - y|\).
\end{definition}

\begin{proposition}[Basic properties of absolute value and distance]\label{4.3.3}
Let \(x\), \(y\), \(z\) be rational numbers.
\begin{enumerate}
    \item (Non-degeneracy of absolute value)
    We have \(|x| \geq 0\).
    Also, \(|x| = 0\) if and only if \(x\) is \(0\).
    \item (Triangle inequality for absolute value)
    We have \(|x + y| \leq |x| + |y|\).
    \item We have the inequalities \(-y \leq x \leq y\) if and only if \(y \geq |x|\).
    In particular, we have \(-|x| \leq x \leq |x|\).
    \item (Multiplicativity of absolute value)
    We have \(|xy| = |x| |y|\).
    In particular, \(|-x| = |x|\).
    \item (Non-degeneracy of distance)
    We have \(d(x, y) \geq 0\).
    Also, \(d(x, y) = 0\) if and only if \(x = y\).
    \item (Symmetry of distance)
    \(d(x, y) = d(y, x)\).
    \item (Triangle inequality for distance)
    \(d(x, z) \leq d(x, y) + d(y, z)\).
\end{enumerate}
\end{proposition}

\begin{proof}{(a)}
By Lemma \ref{4.2.7}, exactly one of the three statements is true:
\begin{enumerate}[label=(\roman*)]
    \item \(x = 0\).
    Then by Definition \ref{4.3.1}, \(|x| = 0\).
    \item \(x\) is a positive rational number.
    Then by Definition \ref{4.3.1}, \(|x| = x\), which is a positive rational number.
    \item \(x\) is a negative rational number.
    Then by Definition \ref{4.3.1}, \(|x| = -x\).
    By Additional Corollary \ref{ac 4.2.5}, \(-x = (-1)x\) is a positive rational number.
\end{enumerate}
So \(|x|\) is either \(0\) or a positive rational number, which by Definition \ref{4.2.8}, \(|x| \geq 0\).

Now we proof that \(|x| = 0 \iff x = 0\).
By Definition \ref{4.3.1}, \(x = 0 \implies |x| = 0\), so we only need to show that \(|x| = 0 \implies x = 0\).
By Lemma \ref{4.2.7}, exactly one of the following is true:
\(x = 0\), \(x\) is positive rational number or \(x\) is negative rational number.
If \(x\) is positive rational number, then \(|x| = x \neq 0\).
If \(x\) is negative rational number, then \(|x| = -x \neq 0\).
So \(x\) can only be \(0\), which means \(|x| = 0 \implies x = 0\).
\end{proof}

\begin{proof}{(b)}
By Lemma \ref{4.2.7}, \(x\) and \(y\) can both have three different cases.
\begin{enumerate}[label=(\Roman*)]
    \item For \(x = 0\) and
    \begin{enumerate}[label=(\roman*)]
        \item \(y = 0\).
        By Proposition \ref{4.2.4} and Definition \ref{4.3.1}, \(|0 + 0| = |0| = 0\), and \(|0| + |0| = 0 + 0 = 0\).
        Thus \(|x + y| = 0 = |x| + |y|\).
        \item \(y\) is positive.
        By Proposition \ref{4.2.4} and Definition \ref{4.3.1}, \(|0 + y| = |y| = y\), and \(|0| + |y| = 0 + y = y\).
        Thus \(|x + y| = y = |x| + |y|\).
        \item \(y\) is negative.
        By Proposition \ref{4.2.4} and Definition \ref{4.3.1}, \(|0 + y| = |y| = -y\), and \(|0| + |y| = 0 + (-y) = -y\).
        Thus \(|x + y| = -y = |x| + |y|\).
    \end{enumerate}
    \item For \(x\) is positive and
    \begin{enumerate}[label=(\roman*)]
        \item \(y = 0\).
        Which is just equivalent to the case \(x = 0\) and \(y\) is positive.
        \item \(y\) is positive.
        By Additional Corollary \ref{ac 4.2.4} and Definition \ref{4.3.1}, \(|x + y| = x + y\), and \(|x| + |y| = x + y\).
        Thus \(|x + y| = x + y = |x| + |y|\).
        \item \(y\) is negative.
        Let \(y = -a\), where \(a\) is a positive rational number.
        By Proposition \ref{4.2.9}, exactly one of the following three statements is true:
        \begin{enumerate}[label=(\arabic*)]
            \item \(x = a\).
            By Proposition \ref{4.2.4} and Definition \ref{4.3.1}, \((|x| + |-a|) - |x - a| = (|x| + |-a|) - |0| = (x + a) - 0 = x + a\).
            By Additional Corollary \ref{ac 4.2.4}, \(x + a\) is a positive rational number.
            Thus by Definition \ref{4.2.8}, \(|x + y| = 0 < x + a = |x| + |y|\).
            \item \(x > a\).
            By Definition \ref{4.2.8}, \(x - a\) is a positive rational number, so by Definition \ref{4.3.1}, \(|x - a| = x - a\).
            By Definition \ref{4.3.1}, \(|x| + |-a| = x + a\).
            By Proposition \ref{4.2.4} and Additional Corollary \ref{ac 4.2.3}, \((|x| + |y|) - |x + y| = (x + a) - (x - a) = 2a\).
            By Additional Corollary \ref{ac 4.2.5}, \(2a\) is a positive rational number.
            Thus by Definition \ref{4.2.8}, \(|x + y| = x - a < x + a = |x| + |y|\).
            \item \(x < a\).
            By Definition \ref{4.2.8}, \(x - a\) is a negative rational number, so by Definition \ref{4.3.1}, \(|x - a| = -(x - a) = a - x\).
            By Definition \ref{4.3.1}, \(|x| + |-a| = x + a\).
            By Proposition \ref{4.2.4} and Additional Corollary \ref{ac 4.2.3}, \((|x| + |y|) - |x + y| = (x + a) - (a - x) = 2x\).
            By Additional Corollary \ref{ac 4.2.5}, \(2x\) is a positive rational number.
            Thus by Definition \ref{4.2.8}, \(|x + y| = a - x < x + a = |x| + |y|\).
        \end{enumerate}
    \end{enumerate}
    \item For \(x\) is negative and
    \begin{enumerate}[label=(\roman*)]
        \item \(y = 0\).
        Which is just equivalent to the case \(x = 0\) and \(y\) is negative.
        \item \(y\) is positive.
        Which is just equivalent to the case \(x\) is positive and \(y\) is negative.
        \item \(y\) is negative.
        By Additional Corollary \ref{ac 4.2.4} and Definition \ref{4.3.1}, \(|x + y| = -(x + y)\).
        By Definition \ref{4.3.1}, Additional Corollary \ref{ac 4.2.3} and Proposition \ref{4.2.4}, \(|x| + |y| = (-x) + (-y) = (-1)x + (-1)y = (-1)(x + y) = -(x + y)\).
        Thus \(|x + y| = -(x + y) = |x| + |y|\).
    \end{enumerate}
\end{enumerate}
For all cases above, we get either \(|x + y| = |x| + |y|\) or \(|x + y| < |x| + |y|\).
So by Definition \ref{4.2.8}, \(|x + y| \leq |x| + |y|\).
\end{proof}

\begin{proof}{(c)}
We first prove that \(-y \leq x \leq y\) implies \(y \geq |x|\).
\begin{align*}
& -y \leq x \leq y \\
\iff & (-y \leq x) \land (x \leq y) & \text{(by Definition \ref{4.2.8})} \\
\iff & ((-y < x) \lor (x = -y)) \land ((x < y) \lor (x = y)) & \text{(by Definition \ref{4.2.8})} \\
\iff & ((-x < y) \lor (x = -y)) \land ((x < y) \lor (x = y)) & \text{(by Exercise \ref{ex 4.2.6})} \\
\iff & ((-x < y) \lor (-x = y)) \land ((x < y) \lor (x = y)). & \text{(by Additional Corollary \ref{ac 4.2.3})}
\end{align*}
By Proposition \ref{4.2.9}, exactly one of the two statements \(-x < y\) and \(-x = y\) is true.
Similarly, exactly one of the two statements \(x < y\) and \(x = y\) is true.
So we can divide into four cases:
\begin{enumerate}[label=(\Roman*)]
    \item \((-x < y) \land (x < y)\).
    % By Definition \ref{4.3.1}, if \(|x| = -x\), then \(|x| = -x < y\);
    % if \(|x| = x\), then \(|x| = x < y\);
    % if \(\).
    % In either case we get \(|x| < y\).
    By Definition \ref{4.3.1}, we get three cases:
    \begin{enumerate}[label=(\roman*)]
        \item \(|x| = 0\).
        Then \(-x = -0 = 0 < y\) and \(0 < y\).
        So \(|x| = 0 < y\).
        \item \(|x| = x\).
        So \(|x| = x < y\).
        \item \(|x| = -x\).
        So \(|x| = -x < y\).
    \end{enumerate}
    \item \((-x < y) \land (x = y)\).
    By Definition \ref{4.3.1}, we get three cases:
    \begin{enumerate}[label=(\roman*)]
        \item \(|x| = 0\).
        This case does not exist since \(-0 = 0 < y\) and \(0 = y\) can not be true at the same time by Proposition \ref{4.2.9}.
        \item \(|x| = x\).
        So \(|x| = x = y\).
        \item \(|x| = -x\).
        So \(|x| = -x < y\).
    \end{enumerate}
    \item \((-x = y) \land (x < y)\).
    By Definition \ref{4.3.1}, we get three cases:
    \begin{enumerate}[label=(\roman*)]
        \item \(|x| = 0\).
        This case does not exist since \(-0 = 0 = y\) and \(0 < y\) can not be true at the same time by Proposition \ref{4.2.9}.
        \item \(|x| = x\).
        So \(|x| = x < y\).
        \item \(|x| = -x\).
        So \(|x| = -x = y\).
    \end{enumerate}
    \item \((-x = y) \land (x = y)\).
    Which means \(-x = x \implies 0 = 2x \implies x = 0\) by Proposition \ref{4.2.4}.
    So \(|x| = 0 = y\).
\end{enumerate}
From all cases above, we can see that either \(|x| = y\) or \(|x| < y\) is true.
Thus by Definition \ref{4.2.8}, \(|x| \leq y\).

Now we prove that \(y \geq |x|\) implies \(-y \leq x \leq y\).
By Definition \ref{4.2.8}, \(y \geq |x| \iff (y > |x|) \lor (y = |x|)\).
By Lemma \ref{4.2.7} and Definition \ref{4.3.1}, \(x\) has three different cases:
\begin{enumerate}[label=(\roman*)]
    \item \(x = 0\) and \(|x| = 0\).
    So \(y > |x| = 0 = x\) or \(y = |x| = 0 = x\) is true, which means \(y \geq x\) by Definition \ref{4.2.8}.
    \item \(x\) is a positive rational number and \(|x| = x\).
    So \(y > |x| = x\) or \(y = |x| = x\) is true, which means \(y \geq x\) by Definition \ref{4.2.8}.
    \item \(x\) is a negative rational number and \(|x| = -x\).
    By Exercise \ref{ex 4.2.6}, \(y > -x \iff -y < x\).
    And by Additional Corollary \ref{4.2.3}, \(y = -x \iff -y = x\).
    So \(-y < x\) or \(-y = x\) is true, which means \(-y \leq x\) by Definition \ref{4.2.8}.
\end{enumerate}
Combine all cases above we get \(-y \leq x \leq y\).

Finally we show that \(-|x| \leq x \leq |x|\).
Because \(|x| = |x|\), so \(|x| \geq |x|\) by Definition \ref{4.2.8}.
Then from previous proof we get \(|x| \geq |x| \iff -|x| \leq x \leq |x|\).
\end{proof}

\begin{proof}{(d)}
By Lemma \ref{4.2.7}, exactly one of the following three statements is true:
\begin{enumerate}[label=(\Roman*)]
    \item \(x = 0\).
    By Proposition \ref{4.2.4} and Definition \ref{4.3.1}, \(|xy| = |0y| = |0| = 0\) and \(|x||y| = |0||y| = 0|y| = 0\).
    So \(|xy| = 0 = |x||y|\).
    \item \(x\) is a positive rational number.
    By Lemma \ref{4.2.7}, exactly one of the following three statements is true:
    \begin{enumerate}[label=(\roman*)]
        \item \(y = 0\).
        Which is just the same case as \(x = 0\).
        \item \(y\) is a positive rational number.
        By Additional Corollary \ref{ac 4.2.5}, \(xy\) is a positive rational number.
        Then by Definition \ref{4.3.1}, \(|xy| = xy\) and \(|x||y| = xy\).
        So \(|xy| = xy = |x||y|\).
        \item \(y\) is a negative rational number.
        By Additional Corollary \ref{ac 4.2.6}, \(xy\) is a negative rational number.
        Then by Definition \ref{4.3.1}, \(|xy| = -(xy)\).
        By Definition \ref{4.3.1}, Additional Corollary \ref{ac 4.2.3} and Proposition \ref{4.2.4}, \(|x||y| = x(-y) = x((-1)y) = (x(-1))y = ((-1)x)y = (-1)(xy) = -(xy)\).
        So \(|xy| = -(xy) = |x||y|\).
    \end{enumerate}
    \item \(x\) is a negative rational number.
    By Lemma \ref{4.2.7}, exactly one of the following three statements is true:
    \begin{enumerate}[label=(\roman*)]
        \item \(y = 0\).
        Which is just the same case as \(x = 0\).
        \item \(y\) is a positive rational number.
        Which is just the same case as \(x\) is a positive rational number and \(y\) is a negative rational number.
        \item \(y\) is a negative rational number.
        By Additional Corollary \ref{ac 4.2.5}, \(xy\) is a positive rational number.
        Then by Definition \ref{4.3.1}, \(|xy| = xy\).
        By Definition \ref{4.3.1}, Additional Corollary \ref{ac 4.2.3} and Proposition \ref{4.2.4}, \(|x||y| = (-x)(-y) = ((-1)x)((-1)y) = (x(-1))((-1)y) = (x((-1)(-1)))y = (x1)y = xy\).
        So \(|xy| = xy = |x||y|\).
    \end{enumerate}
\end{enumerate}
From all cases above, we can see that \(|xy| = |x||y|\).

Now we show that \(|-x| = |x|\).
Using previous proof, let \(y = -1\), we get \(|x(-1)| = |x||-1|\).
By Proposition \ref{4.2.4}, Definition \ref{4.3.1} and Additional Corollary \ref{4.2.3}, \(|x(-1)| = |(-1)x| = |-x|\).
Again by Proposition \ref{4.2.4}, Definition \ref{4.3.1} and Additional Corollary \ref{4.2.3}, \(|x||-1| = |x|(-(-1)) = |x|((-1)(-1)) = |x|1 = |x|\).
Thus \(|-x| = |x|\).
\end{proof}

\begin{proof}{(e)}
We first show that \(d(x, y) \geq 0\).
By Definition \ref{4.3.2}, \(d(x, y) = |x - y|\).
By Proposition \ref{4.3.3}(a), \(|x - y| \geq 0\).
So \(d(x, y) = |x - y| \geq 0\).

Now we show that \(d(x, y) = 0 \iff x = y\).
\begin{align*}
& d(x, y) = 0 & \text{(by the given condition)} \\
\iff & |x - y| = 0 & \text{(by Definition \ref{4.3.2})} \\
\iff & x - y = 0 & \text{(by Proposition \ref{4.3.3}(a))} \\
\iff & (x - y) + y = 0 + y & \text{(by Proposition \ref{4.2.4})} \\
\iff & (x - y) + y = y & \text{(by Proposition \ref{4.2.4})} \\
\iff & (x + (-y)) + y = y \\
\iff & x + ((-y) + y) = y & \text{(by Proposition \ref{4.2.4})} \\
\iff & x + 0 = y & \text{(by Proposition \ref{4.2.4})} \\
\iff & x = y. & \text{(by Proposition \ref{4.2.4})}
\end{align*}
\end{proof}

\begin{proof}{(f)}
\begin{align*}
d(x, y) &= |x - y| & \text{(by Definition \ref{4.3.2})} \\
&= |x + (-y)| \\
&= |-(x + (-y))| & \text{(by Proposition \ref{4.3.3}(d))} \\
&= |(-1)(x + (-1)y)| & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
&= |(-1)x + (-1)((-1)y)| & \text{(by Proposition \ref{4.2.4})} \\
&= |(-1)x + ((-1)(-1))y| & \text{(by Proposition \ref{4.2.4})} \\
&= |(-1)x + 1y| & \text{(by Proposition \ref{4.2.4})} \\
&= |(-1)x + y| & \text{(by Proposition \ref{4.2.4})} \\
&= |y + (-1)x| & \text{(by Proposition \ref{4.2.4})} \\
&= |y + -x| & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
&= |y - x| \\
&= d(y, x). & \text{(by Definition \ref{4.3.2})}
\end{align*}
\end{proof}

\begin{proof}{(g)}
\begin{align*}
d(x, z) &= |x - z| & \text{(by Definition \ref{4.3.2})} \\
&= |x + (-z)| \\
&= |(x + (-z)) + 0| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (-z)) + ((-y) + y)| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (((-z) + (-y)) + y)| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (((-y) + (-z)) + y)| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (-y)) + ((-z) + y)| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (-y)) + (y + (-z))| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x - y) + (y - z)| \\
&\leq |x - y| + |y - z| & \text{(by Proposition \ref{4.3.3}(b))} \\
&= d(x, y) + d(y, z). & \text{(by Definition \ref{4.3.2})} \\
\end{align*}
\end{proof}

\begin{definition}[\(\varepsilon\)-closeness]\label{4.3.4}
Let \(\varepsilon > 0\) be a rational number, and let \(x\), \(y\) be rational numbers.
We say that \(y\) is \emph{\(\varepsilon\)-close} to \(x\) iff we have \(d(y, x) \leq \varepsilon\).
\end{definition}

\begin{remark}\label{4.3.5}
This definition is not standard in mathematics textbooks;
we will use it as ``scaffolding'' to construct the more important notions of limits (and of Cauchy sequences) later on, and once we have those more advanced notions we will discard the notion of \(\varepsilon\)-close.
\end{remark}

\begin{note}
We do not bother defining a notion of \(\varepsilon\)-close when \(\varepsilon\) is zero or negative, because if \(\varepsilon\) is zero then \(x\) and \(y\) are only \(\varepsilon\)-close when they are equal, and when \(\varepsilon\) is negative then \(x\) and \(y\) are never \(\varepsilon\)-close.
\end{note}

\begin{note}
In any event it is a long-standing tradition in analysis that the Greek letters \(\varepsilon\), \(\delta\) should only denote small positive numbers.
\end{note}

\setcounter{theorem}{6}
\begin{proposition}\label{4.3.7}
Let \(x, y, z, w\) be rational numbers.
\begin{enumerate}
    \item If \(x = y\), then \(x\) is \(\varepsilon\)-close to \(y\) for every \(\varepsilon > 0\).
    Conversely, if \(x\) is \(\varepsilon\)-close to \(y\) for every \(\varepsilon > 0\), then we have \(x = y\).
    \item Let \(\varepsilon > 0\).
    If \(x\) is \(\varepsilon\)-close to \(y\), then \(y\) is \(\varepsilon\)-close to \(x\).
    \item Let \(\varepsilon, \delta > 0\).
    If \(x\) is \(\varepsilon\)-close to \(y\), and \(y\) is \(\delta\)-close to \(z\), then \(x\) and \(z\) are \((\varepsilon + \delta)\)-close.
    \item Let \(\varepsilon, \delta > 0\).
    If \(x\) and \(y\) are \(\varepsilon\)-close, and \(z\) and \(w\) are \(\delta\)-close, then \(x + z\) and \(y + w\) are \((\varepsilon + \delta)\)-close, and \(x - z\) and \(y - w\) are also \((\varepsilon + \delta)\)-close.
    \item Let \(\varepsilon > 0\).
    If \(x\) and \(y\) are \(\varepsilon\)-close, they are also \(\varepsilon'\)-close for every \(\varepsilon' > \varepsilon\).
    \item Let \(\varepsilon > 0\).
    If \(y\) and \(z\) are both \(\varepsilon\)-close to \(x\), and \(w\) is between \(y\) and \(z\) (i.e., \(y \leq w \leq z\) or \(z \leq w \leq y\)), then \(w\) is also \(\varepsilon\)-close to \(x\).
    \item Let \(\varepsilon > 0\).
    If \(x\) and \(y\) are \(\varepsilon\)-close, and \(z\) is non-zero, then \(xz\) and \(yz\) are \(\varepsilon|z|\)-close.
    \item Let \(\varepsilon, \delta > 0\).
    If \(x\) and \(y\) are \(\varepsilon\)-close, and \(z\) and \(w\) are \(\delta\)-close, then \(xz\) and \(yw\) are \((\varepsilon|z| + \delta|x| + \varepsilon\delta)\)-close.
\end{enumerate}
\end{proposition}

\begin{proof}{(a)}
We first prove that \(x = y\) implies \(x\) is \(\varepsilon\)-close to \(y\), \(\forall\ \varepsilon > 0\).
By Proposition \ref{4.3.3}, \(x = y \iff d(x, y) = 0\).
Then by the given condition, \(\forall\ \varepsilon > 0 = d(x, y)\), which means \(d(x, y) < \varepsilon\) by Proposition \ref{4.2.9}.
By Definition \ref{4.2.8}, \(d(x, y) < \varepsilon \implies d(x, y) \leq \varepsilon\).
Thus by Definition \ref{4.3.4}, \(x\) is \(\varepsilon\)-close to \(y\).

Now we prove that \(\forall\ \varepsilon > 0\), \(x\) is \(\varepsilon\)-close to \(y\) implies \(x = y\).
By the given condition, \(d(x, y) \leq \varepsilon\).
Suppose for sake of contradiction that \(x \neq y\), then \(d(x, y) > 0\) by Proposition \ref{4.3.3}.
Since \(d(x, y) > 0\), \(d(x, y) / 2 > 0\).
But by the given condition, \(\forall\ \varepsilon > 0\), \(d(x, y) \leq \varepsilon\).
Then we get \(d(x, y) \leq d(x, y) / 2\), and because \(d(x, y) \neq 0\), \(d(x, y) < d(x, y) / 2\), a contradiction.
So \(x = y\).
\end{proof}

\begin{proof}{(b)}
\begin{align*}
& x \text{ is \(\varepsilon\)-close to } y \\
\implies & d(x, y) \leq \varepsilon & \text{(by Definition \ref{4.3.4})} \\
\implies & d(y, x) \leq \varepsilon & \text{(by Proposition \ref{4.3.3}, \(d(x, y) = d(y, x)\))} \\
\implies & y \text{ is \(\varepsilon\)-close to } x. & \text{(by Definition \ref{4.3.4})}
\end{align*}
\end{proof}

\begin{proof}{(c)}
By Definition \ref{4.3.4}, \(x\) is \(\varepsilon\)-close to \(y\) implies \(d(x, y) \leq \varepsilon\), and \(y\) is \(\delta\)-close to \(z\) implies \(d(y, z) \leq \delta\).
So
\begin{align*}
& d(y, z) \leq \delta \\
\implies & \varepsilon + d(y, z) \leq \varepsilon + \delta. & \text{(by Proposition \ref{4.3.3})} \\
& d(x, y) \leq \varepsilon \\
\implies & d(x, y) + d(y, z) \leq \varepsilon + d(y, z) & \text{(by Proposition \ref{4.3.3})} \\
\implies & d(x, y) + d(y, z) \leq \varepsilon + \delta & \text{(by Proposition \ref{4.2.9})} \\
\implies & x \text{ is \((\varepsilon + \delta)\)-close to } z. & \text{(by Definition \ref{4.3.4})}
\end{align*}
\end{proof}

\begin{proof}{(d)}
We first prove that If \(x\) and \(y\) are \(\varepsilon\)-close, and \(z\) and \(w\) are \(\delta\)-close, then \(x + z\) and \(y + w\) are \((\varepsilon + \delta)\)-close.
By Definition \ref{4.3.4}, \(x\) is \(\varepsilon\)-close to \(y\) implies \(d(x, y) \leq \varepsilon\), and \(z\) is \(\delta\)-close to \(w\) implies \(d(z, w) \leq \delta\).
Because
\begin{align*}
& d(z, w) \leq \delta \\
\implies & \varepsilon + d(z, w) \leq \varepsilon + \delta. & \text{(by Proposition \ref{4.2.9})} \\
& d(x, y) \leq \varepsilon \\
\implies & d(x, y) + d(z, w) \leq \varepsilon + d(z, w) & \text{(by Proposition \ref{4.2.9})} \\
\implies & d(x, y) + d(z, w) \leq \varepsilon + \delta. & \text{(by Proposition \ref{4.2.9})}
\end{align*}
So
\begin{align*}
& d(x + z, y + w) \\
&= |(x + z) - (y + w)| & \text{(by Definition \ref{4.3.2})} \\
&= |(x + z) + (-(y + w))| \\
&= |(x + z) + (-1)(y + w)| & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
&= |(x + z) + ((-1)y + (-1)w)| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (z + (-1)y)) + (-1)w| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + ((-1)y + z)) + (-1)w| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (-1)y) + (z + (-1)w)| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (-y)) + (z + (-w))| & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
&= |(x - y) + (z - w)| \\
&\leq |x - y| + |z - w| & \text{(by Proposition \ref{4.3.3})} \\
&= d(x, y) + d(z, w) & \text{(by Definition \ref{4.3.2})} \\
&\leq \varepsilon + \delta. & \text{(by the given conditions)}
\end{align*}
Thus by Definition \ref{4.3.4}, \(x + z\) and \(y + w\) are \((\varepsilon + \delta)\)-close.

Now we prove that If \(x\) and \(y\) are \(\varepsilon\)-close, and \(z\) and \(w\) are \(\delta\)-close, then \(x - z\) and \(y - w\) are \((\varepsilon + \delta)\)-close.
By Definition \ref{4.3.4}, \(x\) is \(\varepsilon\)-close to \(y\) implies \(d(x, y) \leq \varepsilon\), and \(z\) is \(\delta\)-close to \(w\) implies \(d(z, w) \leq \delta\).
Again because
\begin{align*}
& d(z, w) \leq \delta \\
\implies & \varepsilon + d(z, w) \leq \varepsilon + \delta. & \text{(by Proposition \ref{4.2.9})} \\
& d(x, y) \leq \varepsilon \\
\implies & d(x, y) + d(z, w) \leq \varepsilon + d(z, w) & \text{(by Proposition \ref{4.2.9})} \\
\implies & d(x, y) + d(z, w) \leq \varepsilon + \delta. & \text{(by Proposition \ref{4.2.9})}
\end{align*}
So
\begin{align*}
& d(x - z, y - w) \\
&= |(x - z) - (y - w)| & \text{(by Definition \ref{4.3.2})} \\
&= |(x + (-z)) + (-(y + (-w)))| \\
&= |(x + (-1)z) + (-1)(y + (-1)w)| & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
&= |(x + (-1)z) + ((-1)y + (-1)(-1)w)| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (-1)z) + ((-1)y + 1w)| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (-1)z) + ((-1)y + w)| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (-z)) + ((-y) + w)| & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
&= |(x + ((-z) + (-y))) + w| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + ((-y) + (-z))) + w| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (-y)) + ((-z) + w)| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x + (-y)) + (w + (-z))| & \text{(by Proposition \ref{4.2.4})} \\
&= |(x - y) + (w - z)| \\
&\leq |x - y| + |w - z| & \text{(by Proposition \ref{4.3.3})} \\
&= |x - y| + |-(w - z)| & \text{(by Proposition \ref{4.3.3})} \\
&= |x - y| + |-(w + (-z))| \\
&= |x - y| + |(-1)(w + (-1)z)| & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
&= |x - y| + |(-1)w + (-1)((-1)z)| & \text{(by Proposition \ref{4.2.4})} \\
&= |x - y| + |(-1)w + ((-1)(-1))z| & \text{(by Proposition \ref{4.2.4})} \\
&= |x - y| + |(-1)w + 1z| & \text{(by Proposition \ref{4.2.4})} \\
&= |x - y| + |(-1)w + z| & \text{(by Proposition \ref{4.2.4})} \\
&= |x - y| + |z + (-1)w| & \text{(by Proposition \ref{4.2.4})} \\
&= |x - y| + |z + (-w)| & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
&= |x - y| + |z - w| \\
&= d(x, y) + d(z, w) & \text{(by Definition \ref{4.3.2})} \\
&\leq \varepsilon + \delta. & \text{(by the given conditions)}
\end{align*}
Thus by Definition \ref{4.3.4}, \(x - z\) and \(y - w\) are \((\varepsilon + \delta)\)-close.
\end{proof}

\begin{proof}{(e)}
By the given condition, \(\forall\ \varepsilon'\), \(\varepsilon < \varepsilon'\).
And by Definition \ref{4.3.4}, \(x\) is \(\varepsilon\)-close to \(y\) implies \(d(x, y) \leq \varepsilon\).
If \(d(x, y) = \varepsilon\), then \(d(x, y) < \varepsilon'\).
If \(d(x, y) < \varepsilon\), then by Proposition \ref{4.2.9}, \(d(x, y) < \varepsilon'\).
Thus \(d(x, y) < \varepsilon'\), by Definition \ref{4.2.8}, \(d(x, y) \leq \varepsilon'\), which means \(x\) is \(\varepsilon'\)-close to \(y\) by Definition \ref{4.3.4}.
\end{proof}

\begin{proof}{(f)}
\begin{align*}
& y \text{ is } \varepsilon\text{-close to } x \\
\implies & d(y, x) \leq \varepsilon & \text{(by Definition \ref{4.3.4})} \\
\implies & |y - x| \leq \varepsilon & \text{(by Definition \ref{4.3.2})} \\
\implies & (-1)|y - x| \geq (-1)\varepsilon & \text{(by Exercise \ref{ex 4.2.6})} \\
\implies & -|y - x| \geq -\varepsilon. \\
& z \text{ is } \varepsilon\text{-close to } x \\
\implies & d(z, x) \leq \varepsilon & \text{(by Definition \ref{4.3.4})} \\
\implies & |z - x| \leq \varepsilon & \text{(by Definition \ref{4.3.2})} \\
\implies & (-1)|z - x| \geq (-1)\varepsilon & \text{(by Exercise \ref{ex 4.2.6})} \\
\implies & -|z - x| \geq -\varepsilon. \\
& y \leq w \leq z \\
\implies & y + (-x) \leq w + (-x) \leq z + (-x) & \text{(by Lemma \ref{4.2.9})} \\
\implies & y - x \leq w - x \leq z - x \\
\implies & -|y - x| \leq y - x \leq w - x \leq z - x \leq |z - x| & \text{(by Proposition \ref{4.3.3})} \\
\implies & -|y - x| \leq w - x \leq |z - x| \\
\implies & -\varepsilon \leq -|y - x| \leq w - x \leq |z - x| \leq \varepsilon & \text{(by the given conditions)} \\
\implies & -\varepsilon \leq w - x \leq \varepsilon \\
\implies & \varepsilon \geq |w - x| & \text{(by Proposition \ref{4.3.3})} \\
\implies & |w - x| \leq \varepsilon & \text{(by Proposition \ref{4.2.9})} \\
\implies & d(w, x) \leq \varepsilon & \text{(by Definition \ref{4.3.2})} \\
\implies & w \text{ is } \varepsilon\text{-close to } x. \\
& z \leq w \leq y \\
\implies & z + (-x) \leq w + (-x) \leq y + (-x) & \text{(by Lemma \ref{4.2.9})} \\
\implies & z - x \leq w - x \leq y - x \\
\implies & -|z - x| \leq z - x \leq w - x \leq y - x \leq |y - x| & \text{(by Proposition \ref{4.3.3})} \\
\implies & -|z - x| \leq w - x \leq |y - x| \\
\implies & -\varepsilon \leq -|z - x| \leq w - x \leq |y - x| \leq \varepsilon & \text{(by the given conditions)} \\
\implies & -\varepsilon \leq w - x \leq \varepsilon \\
\implies & \varepsilon \geq |w - x| & \text{(by Proposition \ref{4.3.3})} \\
\implies & |w - x| \leq \varepsilon & \text{(by Proposition \ref{4.2.9})} \\
\implies & d(w, x) \leq \varepsilon & \text{(by Definition \ref{4.3.2})} \\
\implies & w \text{ is } \varepsilon\text{-close to } x.
\end{align*}
\end{proof}

\begin{proof}{(g)}
By Proposition \ref{4.3.3}, \(|z| \geq 0\).
So
\begin{align*}
& x \text{ is } \varepsilon\text{-close to } y \\
\implies & d(x, y) \leq \varepsilon & \text{(by Definition \ref{4.3.4})} \\
\implies & |x - y| \leq \varepsilon & \text{(by Definition \ref{4.3.2})} \\
\implies & |x - y||z| \leq \varepsilon|z| & \text{(by Proposition \ref{4.2.9})} \\
\implies & |(x - y)z| \leq \varepsilon|z| & \text{(by Proposition \ref{4.3.3})} \\
\implies & |(x + (-y))z| \leq \varepsilon|z| \\
\implies & |xz + (-y)z| \leq \varepsilon|z| & \text{(by Proposition \ref{4.2.4})} \\
\implies & |xz + ((-1)y)z| \leq \varepsilon|z| & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
\implies & |xz + (-1)(yz)| \leq \varepsilon|z| & \text{(by Proposition \ref{4.2.4})} \\
\implies & |xz + (-(yz))| \leq \varepsilon|z| & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
\implies & |xz - yz| \leq \varepsilon|z| \\
\implies & d(xz, yz) \leq \varepsilon|z| & \text{(by Definition \ref{4.3.2})} \\
\implies & xz \text{ is } \varepsilon|z|\text{-close to } yz. & \text{(by Definition \ref{4.3.4})} \\
\end{align*}
\end{proof}

\begin{proof}{(h)}
Let \(\varepsilon, \delta > 0\), and suppose that \(x\) and \(y\) are \(\varepsilon\)-close.
If we write \(a \coloneqq y - x\), then we have \(y = x + a\) and that \(|a| \leq \varepsilon\).
Similarly, if \(z\) and \(w\) are \(\delta\)-close, and we define \(b \coloneqq w - z\), then \(w = z + b\) and \(|b| \leq \delta\).

Since \(y = x + a\) and \(w = z + b\), we have
\[
    yw = (x + a)(z + b) = xz + az + xb + ab.
\]
Thus
\[
    |yw - xz| = |az + bx + ab| \leq |az| + |bx| + |ab| = |a||z| + |b||x| + |a||b|.
\]
Since \(|a| \leq \varepsilon\) and \(|b| \leq \delta\), we thus have
\[
    |yw - xz| \leq \varepsilon|z| + \delta|x| + \varepsilon\delta
\]
and thus that \(yw\) and \(xz\) are \((\varepsilon|z| + \delta|x| + \varepsilon\delta)\)-close.
\end{proof}

\begin{remark}\label{4.3.8}
One should compare statements (a)-(c) of Proposition \ref{4.3.7} with the reflexive, symmetric, and transitive axioms of equality.
It is often useful to think of the notion of ``\(\varepsilon\)-close'' as an approximate substitute for that of equality in analysis.
\end{remark}

\begin{definition}[Exponentiation to a natural number]\label{4.3.9}
Let \(x\) be a rational number.
To raise \(x\) to the power \(0\), we define \(x^0 \coloneqq 1\);
in particular we define \(0^0 \coloneqq 1\).
Now suppose inductively that \(x^n\) has been defined for some natural number \(n\), then we define \(x^{n+1} \coloneqq x^n \times x\).
\end{definition}

\begin{proposition}[Properties of exponentiation, I]\label{4.3.10}
Let \(x\), \(y\) be rational numbers, and let \(n\), \(m\) be natural numbers.
\begin{enumerate}
    \item We have \(x^n x^m = x^{n + m}\), \((x^n)^m = x^{nm}\), and \((xy)^n = x^n y^n\).
    \item Suppose \(n > 0\).
    Then we have \(x^n = 0\) if and only if \(x = 0\).
    \item If \(x \geq y \geq 0\), then \(x^n \geq y^n \geq 0\).
    If \(x > y \geq 0\) and \(n > 0\), then \(x^n > y^n \geq 0\).
    \item We have \(|x^n| = |x|^n\).
\end{enumerate}
\end{proposition}

\begin{proof}{(a)}
We first prove that \(x^n x^m = x^{n + m}\).
We use induction on \(n\).
For \(n = 0\), \(x^0 x^m = 1 x^m = x^m\) by Definition \ref{4.3.9} and Proposition \ref{4.2.4}.
And \(x^{0 + m} = x^m\) by Proposition \ref{4.2.4}.
So \(x^0 x^m = x^{0 + m}\), and the base case holds.
Suppose inductively that for some \(n\), \(x^n x^m = x^{n + m}\).
Then for \(n++\), \(x^{n++} x^m = (x^n x) x^m = x^n (x x^m) = (x^n x^m)x\) by Definition \ref{4.3.9} and Proposition \ref{4.2.4}.
And \(x^{(n++) + m} = x^{(n + m)++} = x^{n + m} x = (x^n x^m)x\) by Definition \ref{2.2.1} and induction hypothesis.
So \(x^{n++} x^m = x^{(n++) + m}\), and this close the induction.

Next we prove that \((x^n)^m = x^{nm}\).
We use induction on \(m\).
For \(m = 0\), \((x^n)^0 = 1\) by Definition \ref{4.3.9}.
And \(x^{n0} = x^0 = 1\) by Additional Corollary \ref{ac 2.3.2} and Definition \ref{4.3.9}.
So \((x^n)^0 = x^{n0}\), and the base case holds.
Suppose inductively that for some \(m\), \((x^n)^m = x^{nm}\).
Then for \(m++\), \((x^n)^{m++} = (x^n)^m x^n = x^{nm} x^n\) by Definition \ref{4.3.9} and induction hypothesis.
And \(x^{n(m++)} = x^{nm + n} = x^{nm} x^n\) by Additional Corollary \ref{ac 2.3.3} and previous prove.
So \((x^n)^{m++} = x^{n(m++)}\), and this close the induction.

Finally we prove that \((xy)^n = x^n y^n\).
We use induction on \(n\).
For \(n = 0\), \((xy)^0 = 1\) by Definition \ref{4.3.9}.
And \(x^0 y^0 = 1 \times 1 = 1\) by Definition \ref{4.3.9}.
So \((xy)^0 = x^0 y^0\), and the base case holds.
Suppose inductively that for some \(n\), \((xy)^n = x^n y^n\).
Then for \(n++\), \((xy)^{n++} = (xy)^n (xy) = (x^n y^n)(xy)\) by Definition \ref{4.3.9} and induction hypothesis.
And \(x^{n++} y^{n++} = (x^n x)(y^n y) = (x^n (x y^n))y = (x^n (y^n x))y = (x^n y^n)(xy)\) by Definition \ref{4.3.9} and Proposition \ref{4.2.4}.
So \((xy)^{n++} = x^{n++} y^{n++}\), and this close the induction.
\end{proof}

\begin{proof}{(b)}
We first prove that \(x^n = 0\) implies \(x = 0\).
We use induction on \(n\) and begin with \(n = 1\).
For \(n = 1\), \(x^1 = x^0 x = 1x = x = 0\) by Definition \ref{4.3.9} and Proposition \ref{4.2.4}.
So \(x^1 = 0 \implies x = 0\), and the base case holds.
Suppose inductively that for some \(n\), \(x^n = 0 \implies x = 0\).
Then for \(n++\), \(x^{n++} = x^n x = 0\) by Definition \ref{4.3.9}.
Suppose for sake of contradiction that \(x \neq 0\).
Then \(x^n x = 0 \implies x^n = 0 / x = 0\).
But by induction hypothesis, \(x^n = 0 \implies x = 0\), a contradiction.
Thus \(x = 0\), which means \(x^{n++} = 0 \implies x = 0\), and this close the induction.

Now we prove that \(x = 0\) implies \(x^n = 0\).
Since \(n > 0\), \(x^n = x^{(n - 1) + 1} = x^{n - 1} x = x^{n - 1} 0 = 0\) by Proposition \ref{4.3.10}(a).
Thus \(x = 0 \implies x^n = 0\).
\end{proof}

\begin{proof}{(c)}
We first prove that \(x \geq y \geq 0\) implies \(x^n \geq y^n \geq 0\).
We use induction on \(n\).
For \(n = 0\), \(x^0 = 1\) and \(y^0 = 1\) by Definition \ref{4.3.9}, and \(1 \geq 1 \geq 0\), so the base case holds.
Suppose inductively that for some \(n\), \(x^n \geq y^n \geq 0\).
Then for \(n++\), \(x^{n++} = x^n x \geq y^n x\) and \(y^n x \geq y^n y = y^{n++}\) by Definition \ref{4.3.9}, induction hypothesis, Proposition \ref{4.2.9} and the given conditions.
So \(x^{n++} \geq y^{n++}\) by Proposition \ref{4.2.4}.
And \(y^{n++} = y^n y \geq y^n 0 = 0\) by Definition \ref{4.3.9} and the given conditions.
So \(y^{n++} \geq 0\).
Thus \(x^{n++} \geq y^{n++} \geq 0\), and this close the induction.

Now we prove that \(x > y \geq 0\) and \(n > 0\), then \(x^n > y^n \geq 0\).
We use induction on \(n\) and begin with \(n = 1\).
For \(n = 1\), \(x^1 = x^0 x = 1x = x\) and \(y^1 = y^0 y = 1y = y\) by Definition \ref{4.3.9} and Proposition \ref{4.2.4}.
So \(x^1 > y^1 \geq 0\) by the given conditions, and the base case holds.
Suppose inductively that for some \(n\), \(x^n > y^n \geq 0\).
Then for \(n++\), \(x^{n++} = x^n x > y^n x\) and \(y^n x > y^n y = y^{n++}\) by Definition \ref{4.3.9}, induction hypothesis, Proposition \ref{4.2.9} and the given conditions.
So \(x^{n++} > y^{n++}\) by Proposition \ref{4.2.4}.
And \(y^{n++} = y^n y \geq y^n 0 = 0\) by Definition \ref{4.3.9} and the given conditions.
So \(y^{n++} \geq 0\).
Thus \(x^{n++} > y^{n++} \geq 0\), and this close the induction.
\end{proof}

\begin{proof}{(d)}
We use induction on \(n\).
For \(n = 0\), \(|x^0| = |1| = 1\) and \(|x|^0 = 1\) by Definition \ref{4.3.9} and Definition \ref{4.3.1}.
So \(|x^0| = |x|^0\), and the base case holds.
Suppose inductively that for some \(n\), \(|x^n| = |x|^n\).
Then for \(n++\), \(|x^{n++}| = |x^n x| = |x^n||x| = |x|^n |x|\) and \(|x|^{n++} = |x|^n |x|\) by Definition \ref{4.3.9}, Proposition \ref{4.3.3} and induction hypothesis.
So \(|x^{n++}| = |x|^{n++}\), and this close the induction.
\end{proof}

\begin{definition}[Exponentiation to a negative number]\label{4.3.11}
Let \(x\) be a non-zero rational number.
Then for any negative integer \(-n\), we define \(x^{-n} \coloneqq 1 / x^n\).
\end{definition}

\begin{proposition}[Properties of exponentiation, II]\label{4.3.12}
Let \(x\), \(y\) be nonzero rational numbers, and let \(n\), \(m\) be integers.
\begin{enumerate}
    \item We have \(x^n x^m = x^{n + m}\), \((x^n)^m = x^{nm}\), and \((xy)^n = x^n y^n\).
    \item If \(x \geq y > 0\), then \(x^n \geq y^n > 0\) if \(n\) is positive, and \(0 < x^n \leq y^n\) if \(n\) is negative.
    \item If \(x, y > 0\), \(n \neq 0\), and \(x^n = y^n\), then \(x = y\).
    \item We have \(|x^n| = |x|^n\).
\end{enumerate}
\end{proposition}

\begin{proof}{(a)}
We first prove that \(x^n x^m = x^{n + m}\).
By Lemma \ref{4.1.5}, exactly one of the following three statements is true:
\begin{enumerate}[label=(\Roman*)]
    \item \(n = 0\).
    Then \(x^0 x^m = 1x^m = x^m\) by Definition \ref{4.3.9} and Proposition \ref{4.2.4}.
    And \(x^{0 + m} = x^m\) by Proposition \ref{4.2.4}.
    So \(x^0 x^m = x^{0 + m}\).
    \item \(n\) is a positive integer.
    Again by Lemma \ref{4.1.5}, exactly one of the following three statements is true:
    \begin{enumerate}[label=(\roman*)]
        \item \(m = 0\).
        Then \(x^n x^0 = x^0 x^n\) and \(x^{n + 0} = x^{0 + n}\) by Proposition \ref{4.2.4} and Proposition \ref{4.1.6}, which is the same case as \(n = 0\).
        \item \(m\) is a positive integer.
        Then by Proposition \ref{4.3.10}, \(x^n x^m = x^{n + m}\).
        \item \(m\) is a negative integer.
        Again by Lemma \ref{4.1.5}, exactly one of the following three statements is true:
        \begin{enumerate}[label=(\arabic*)]
            \item \(n + m = 0\).
            Then \(n = -m\) and \(x^n x^m = x^{-m} x^m = 1\) by Proposition \ref{4.2.4}.
            And \(x^{n + m} = x^{(-m) + m} = x^0 = 1\) by Proposition \ref{4.2.4} and Definition \ref{4.3.9}.
            So \(x^{-m} x^m = x^{(-m) + m}\).
            \item \(n + m\) is a positive integer.
            By Proposition \ref{4.3.10}, \(x \neq 0 \implies x^{-m} \neq 0\).
            Then
            \begin{align*}
            & (x^n x^m) x^{-m} \\
            &= x^n (x^m x^{-m}) & \text{(by Proposition \ref{4.2.4})} \\
            &= x^n 1 & \text{(by Proposition \ref{4.2.4})} \\
            &= x^n & \text{(by Proposition \ref{4.2.4})} \\
            &= x^{n + 0} & \text{(by Proposition \ref{4.1.6})} \\
            &= x^{n + (m + (-m))} & \text{(by Proposition \ref{4.1.6})} \\
            &= x^{(n + m) + (-m)} & \text{(by Proposition \ref{4.1.6})} \\
            &= x^{n + m} x^{-m}. & \text{(by Proposition \ref{4.3.10})}
            \end{align*}
            So
            \begin{align*}
            & (x^n x^m) x^{-m} = x^{n + m} x^{-m} \\
            \implies & ((x^n x^m) x^{-m}) x^m = (x^{n + m} x^{-m}) x^m & \text{(by Lemma \ref{4.2.3})} \\
            \implies & (x^n x^m)(x^{-m} x^m) = x^{n + m} (x^{-m} x^m) & \text{(by Proposition \ref{4.2.4})} \\
            \implies & (x^n x^m)1 = x^{n + m} 1 & \text{(by Proposition \ref{4.2.4})} \\
            \implies & x^n x^m = x^{n + m}. & \text{(by Proposition \ref{4.2.4})} \\
            \end{align*}
            \item \(n + m\) is a negative integer.
            Then \(-(n + m)\) is a positive integer.
            By Proposition \ref{4.3.10}, \(x \neq 0 \implies x^n \neq 0\), \(x^{-m} \neq 0\) and \(x^{-(n + m)} \neq 0\).
            So
            \begin{align*}
            & x^n (x^{-n} x^{-m}) \\
            &= (x^n x^{-n})x^{-m} & \text{(by Proposition \ref{4.2.4})} \\
            &= 1x^{-m} & \text{(by Proposition \ref{4.2.4})} \\
            &= x^{-m} & \text{(by Proposition \ref{4.2.4})} \\
            &= x^{0 + (-m)} & \text{(by Proposition \ref{4.1.6})} \\
            &= x^{(n + (-n)) + (-m)} & \text{(by Proposition \ref{4.1.6})} \\
            &= x^{n + ((-n) + (-m))} & \text{(by Proposition \ref{4.1.6})} \\
            &= x^{n + ((-1)n + (-1)m)} & \text{(by Exercise \ref{ex 4.1.3})} \\
            &= x^{n + (-1)(n + m)} & \text{(by Proposition \ref{4.1.6})} \\
            &= x^{n + (-(n + m))} & \text{(by Exercise \ref{ex 4.1.3})} \\
            &= x^n x^{-(n + m)}. & \text{(by Proposition \ref{4.3.10})} \\
            \end{align*}
            And
            \begin{align*}
            & x^n (x^{-n} x^{-m}) = x^n x^{-(n + m)} \\
            \implies & x^{-n} (x^n (x^{-n} x^{-m})) = x^{-n} (x^n x^{-(n + m)}) & \text{(by Lemma \ref{4.2.3})} \\
            \implies & (x^{-n} x^n)(x^{-n} x^{-m}) = (x^{-n} x^n)x^{-(n + m)} & \text{(by Proposition \ref{4.2.4})} \\
            \implies & 1(x^{-n} x^{-m}) = 1x^{-(n + m)} & \text{(by Proposition \ref{4.2.4})} \\
            \implies & x^{-n} x^{-m} = x^{-(n + m)} & \text{(by Proposition \ref{4.2.4})} \\
            \implies & x^n (x^{-n} x^{-m}) = x^n x^{-(n + m)} & \text{(by Lemma \ref{4.2.3})} \\
            \implies & (x^n x^{-n}) x^{-m} = x^n x^{-(n + m)} & \text{(by Proposition \ref{4.2.4})} \\
            \implies & 1x^{-m} = x^n x^{-(n + m)} & \text{(by Proposition \ref{4.2.4})} \\
            \implies & x^{-m} = x^n x^{-(n + m)} & \text{(by Proposition \ref{4.2.4})} \\
            \implies & x^m x^{-m} = x^m (x^n x^{-(n + m)}) & \text{(by Lemma \ref{4.2.3})} \\
            \implies & 1 = x^m (x^n x^{-(n + m)}) & \text{(by Proposition \ref{4.2.4})} \\
            \implies & 1 = (x^m x^n) x^{-(n + m)} & \text{(by Proposition \ref{4.2.4})} \\
            \implies & 1x^{n + m} = ((x^m x^n) x^{-(n + m)}) x^{n + m} & \text{(by Lemma \ref{4.2.3})} \\
            \implies & 1x^{n + m} = (x^m x^n)(x^{-(n + m)} x^{n + m}) & \text{(by Proposition \ref{4.2.4})} \\
            \implies & 1x^{n + m} = (x^m x^n)1 & \text{(by Proposition \ref{4.2.4})} \\
            \implies & x^{n + m} = x^m x^n & \text{(by Proposition \ref{4.2.4})} \\
            \implies & x^{n + m} = x^n x^m. & \text{(by Proposition \ref{4.2.4})}
            \end{align*}
        \end{enumerate}
    \end{enumerate}
    \item \(n\) is a negative integer.
    Again by Lemma \ref{4.1.5}, exactly one of the following three statements is true:
    \begin{enumerate}[label=(\roman*)]
        \item \(m = 0\).
        Then \(x^n x^0 = x^0 x^n\) and \(x^{n + 0} = x^{0 + n}\) by Proposition \ref{4.2.4} and Proposition \ref{4.1.6}, which is the same case as \(n = 0\).
        \item \(m\) is a positive integer.
        Then \(x^n x^m = x^m x^n\) and \(x^{n + m} = x^{m + n}\) by Proposition \ref{4.2.4} and Proposition \ref{4.1.6}, which is the same case as \(n\) is positive and \(m\) is negative.
        \item \(m\) is a negative integer.
        Then \(-n, -m, -(n + m)\) are positive integers.
        By Proposition \ref{4.3.10}, \(x \neq 0 \implies x^{-n} \neq 0\), \(x^{-m} \neq 0\) and \(x^{-(n + m)} \neq 0\).
        So
        \begin{align*}
        & x^{-n} x^{-m} = x^{-(n + m)} & \text{(by Proposition \ref{4.3.10})} \\
        \implies & x^n (x^{-n} x^{-m}) = x^n x^{-(n + m)} & \text{(by Lemma \ref{4.2.3})} \\
        \implies & (x^n x^{-n}) x^{-m} = x^n x^{-(n + m)} & \text{(by Proposition \ref{4.2.4})} \\
        \implies & 1x^{-m} = x^n x^{-(n + m)} & \text{(by Proposition \ref{4.2.4})} \\
        \implies & x^{-m} = x^n x^{-(n + m)} & \text{(by Proposition \ref{4.2.4})} \\
        \implies & x^m x^{-m} = x^m (x^n x^{-(n + m)}) & \text{(by Lemma \ref{4.2.3})} \\
        \implies & 1 = x^m (x^n x^{-(n + m)}) & \text{(by Proposition \ref{4.2.4})} \\
        \implies & 1 = (x^m x^n) x^{-(n + m)} & \text{(by Proposition \ref{4.2.4})} \\
        \implies & 1x^{n + m} = ((x^m x^n) x^{-(n + m)})x^{-(n + m)} & \text{(by Lemma \ref{4.2.3})} \\
        \implies & 1x^{n + m} = (x^m x^n)(x^{-(n + m)} x^{-(n + m)}) & \text{(by Proposition \ref{4.2.4})} \\
        \implies & 1x^{n + m} = (x^m x^n)1 & \text{(by Proposition \ref{4.2.4})} \\
        \implies & x^{n + m} = x^m x^n & \text{(by Proposition \ref{4.2.4})} \\
        \implies & x^{n + m} = x^n x^m. & \text{(by Proposition \ref{4.2.4})}
        \end{align*}
    \end{enumerate}
\end{enumerate}
From all cases above, we can conclude that \(x^n x^m = x^{n + m}\).

Next we prove that \((x^n)^m = x^{nm}\).
By Lemma \ref{4.1.5}, exactly one of the following three statements is true:
\begin{enumerate}[label=(\Roman*)]
    \item \(n = 0\).
    Then by Definition \ref{4.3.9}, \((x^0)^m = 1^m\) and \(x^{0m} = x^0 = 1\).
    Again By Lemma \ref{4.1.5}, exactly one of the following three statements is true:
    \begin{enumerate}[label=(\roman*)]
        \item \(m = 0\).
        Then \(1^0 = 1\) by Definition \ref{4.3.9}, so \((x^0)^0 = x^{0 \times 0}\).
        \item \(m\) is a positive integer.
        We claim that \(1^m = 1\) by using induction on \(m\).
        For \(m = 0\), \(1^0 = 1\) by Definition \ref{4.3.9}, so the base case holds.
        Suppose inductively that for some \(m\), \(1^m = 1\).
        Then for \(m++\), \(1^{m++} = 1^m \times 1 = 1 \times 1 = 1\) by Definition \ref{4.3.9} and induction hypothesis, and this close the induction.
        So \((x^0)^m = x^{0m}\).
        \item \(m\) is a negative integer.
        Then \(-m\) is a positive integer, and \(1^m = 1 / 1^{-m}\) by Definition \ref{4.3.11}.
        From previous prove, we show that \(1^{-m} = 1\).
        So \((x^0)^m = x^{0m}\).
    \end{enumerate}
    \item \(n\) is a positive integer.
    Again by Lemma \ref{4.1.5}, exactly one of the following three statements is true:
    \begin{enumerate}[label=(\roman*)]
        \item \(m = 0\).
        Then by Definition \ref{4.3.9}, \((x^n)^0 = 1\) and \(x^{n0} = x^0 = 1\).
        So \((x^n)^0 = x^{n0}\).
        \item \(m\) is a positive integer.
        Then by Proposition \ref{4.3.10}, \((x^n)^m = x^{nm}\).
        \item \(m\) is a negative integer.
        Then \(-m\) is a positive integer.
        So
        \begin{align*}
        & (x^n)^{-m} = x^{n(-m)} & \text{(by Proposition \ref{4.3.10})} \\
        \implies & (x^n)^{-m} = x^{n((-1)m)} & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
        \implies & (x^n)^{-m} = x^{(n(-1))m} & \text{(by Proposition \ref{4.1.6})} \\
        \implies & (x^n)^{-m} = x^{((-1)n)m} & \text{(by Proposition \ref{4.1.6})} \\
        \implies & (x^n)^{-m} = x^{(-1)(nm)} & \text{(by Proposition \ref{4.1.6})} \\
        \implies & (x^n)^{-m} = x^{-(nm)} & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
        \implies & 1 / (x^n)^m = 1 / x^{nm} & \text{(by Definition \ref{4.3.11})} \\
        \implies & 1x^{nm} = 1(x^n)^m  & \text{(by Definition \ref{4.2.1})} \\
        \implies & x^{nm} = (x^n)^m.  & \text{(by Proposition \ref{4.2.4})}
        \end{align*}
    \end{enumerate}
    \item \(n\) is a negative integer.
    Then \(-n\) is a positive integer.
    Again by Lemma \ref{4.1.5}, exactly one of the following three statements is true:
    \begin{enumerate}[label=(\roman*)]
        \item \(m = 0\).
        Then by Definition \ref{4.3.9}, \((x^n)^0 = 1\) and \(x^{n0} = x^0 = 1\).
        So \((x^n)^0 = x^{n0}\).
        \item \(m\) is a positive integer.
        So
        \begin{align*}
        & (x^{-n})^m = x^{(-n)m} & \text{(by Proposition \ref{4.3.10})} \\
        \implies & (x^{-n})^m = x^{((-1)n)m} & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
        \implies & (x^{-n})^m = x^{(-1)(nm)} & \text{(by Proposition \ref{4.1.6})} \\
        \implies & (x^{-n})^m = x^{-(nm)} & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
        \implies & (1 / x^n)^m = 1 / x^{nm}. & \text{(by Definition \ref{4.3.11})}
        \end{align*}
        We claim that \((1 / x^n)^m = 1 / (x^n)^m\) by using induction on \(m\).
        For \(m = 0\), \((1 / x^n)^0 = 1\) and \(1 / (x^n)^0 = 1 / 1 = 1\) by Definition \ref{4.3.9}.
        So \((1 / x^n)^0 = 1 / (x^n)^0\), and the base case holds.
        Suppose inductively that for some \(m\), \((1 / x^n)^m = 1 / (x^n)^m\).
        Then for \(m++\), \((1 / x^n)^{m++} = (1 / x^n)^m \times (1 / x^n) = 1 / (x^n)^m \times (1 / x^n) = (1 \times 1) / ((x^n)^m \times x^n) = 1 / (x^n)^{m++}\) by Definition \ref{4.3.9}, induction hypothesis and Definition \ref{4.2.2}.
        This close the induction.
        So
        \begin{align*}
        & 1 / (x^n)^m = 1 / x^{nm} \\
        \implies & 1x^{nm} = 1(x^n)^m & \text{(by Definition \ref{4.2.1})} \\
        \implies & x^{nm} = (x^n)^m. & \text{(by Proposition \ref{4.2.4})}
        \end{align*}
        \item \(m\) is a negative integer.
        Then \(-m\) is a positive integer.
        So
        \begin{align*}
        & (x^n)^m \\
        &= 1 / (x^n)^{-m} & \text{(by Definition \ref{4.3.11})} \\
        &= 1 / x^{n(-m)} & \text{(from case above)} \\
        &= 1 / x^{n((-1)m)} & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
        &= 1 / x^{(n(-1))m} & \text{(by Proposition \ref{4.2.4})} \\
        &= 1 / x^{((-1)n)m} & \text{(by Proposition \ref{4.2.4})} \\
        &= 1 / x^{(-1)(nm)} & \text{(by Proposition \ref{4.2.4})} \\
        &= 1 / x^{-(nm)} & \text{(by Additional Corollary \ref{ac 4.2.3})} \\
        &= x^{nm}. & \text{(by Definition \ref{4.3.11})}
        \end{align*}
    \end{enumerate}
\end{enumerate}
From all cases above, we can conclude that \((x^n)^m = x^{nm}\).

Finally we prove that \((xy)^n = x^n y^n\).
By Lemma \ref{4.1.5}, exactly one of the following three statements is true:
\begin{enumerate}[label=(\roman*)]
    \item \(n = 0\).
    Then by Proposition \ref{4.3.10}, \((xy)^0 = x^0 y^0\).
    \item \(n\) is a positive integer.
    Then by Proposition \ref{4.3.10}, \((xy)^n = x^n y^n\).
    \item \(n\) is a negative integer.
    Then \(-n\) is a positive integer.
    So
    \begin{align*}
    & (xy)^{-n} = x^{-n} y^{-n} & \text{(by Proposition \ref{4.3.10})} \\
    \implies & 1 / (xy)^n = (1 / x^n)(1 / y^n) & \text{(by Definition \ref{4.3.11})} \\
    \implies & 1 / (xy)^n = (1 \times 1) / (x^n y^n) & \text{(by Definition \ref{4.2.2})} \\
    \implies & 1 / (xy)^n = 1 / (x^n y^n) \\
    \implies & 1(x^n y^n) = 1(xy)^n & \text{(by Definition \ref{4.2.1})} \\
    \implies & (x^n y^n) = (xy)^n. & \text{(by Proposition \ref{4.2.4})}
    \end{align*}
\end{enumerate}
From all cases above, we can conclude that \((xy)^n = x^n y^n\).
\end{proof}

\begin{proof}{(b)}
By Definition \ref{4.2.8}, \(x \geq y > 0 \implies x \geq y \geq 0\).
If \(n\) is a positive integer, then by Proposition \ref{4.3.10}, \(x \geq y \geq 0 \implies x^n \geq y^n \geq 0\).
By the given conditions and Proposition \ref{4.3.10}, \(y \neq 0 \implies y^n \neq 0\).
So \(x \geq y > 0 \implies x^n \geq y^n > 0\) when \(n\) is a positive integer.

If \(n\) is a negative integer, then let \(n = -a\), where \(a\) is a positive integer.
By the given conditions we have two cases:
\begin{enumerate}[label=(\roman*)]
    \item \(x > y\).
    By Definition \ref{4.2.6}, \((x > 0 \implies 1 / x > 0) \land (y > 0 \implies 1 / y > 0)\).
    By Additional Corollary \ref{ac 4.2.5} and Definition \ref{4.2.2}, \((1 / x) \times (1 / y) = 1 / xy > 0\).
    So \(x > y \implies x \times (1 / xy) > y \times (1 / xy) \implies 1 / y > 1 / x\) by Proposition \ref{4.2.9}.
    From previous prove, we get \(1 / y \geq 1 / x > 0 \implies (1 / y)^a \geq (1 / x)^a > 0\).
    But by Definition \ref{4.3.11}, Proposition \ref{4.3.12}(a) and Additional Corollary \ref{ac 4.2.3}, \((1 / y)^a = (y^{-1})^a = y^{(-1)a} = y^{-a} = y^n\) and \((1 / x)^a = (x^{-1})^a = x^{(-1)a} = x^{-a} = x^n\).
    So \(x \geq y > 0 \implies y^n \geq x^n > 0\) when \(n\) is a negative integer.
    \item \(x = y\).
    Then \(x^a = y^a > 0\) by Proposition \ref{4.3.10}.
    By Definition \ref{4.2.6}, \((x^a > 0 \implies 1 / x^a > 0) \land (y^a > 0 \implies 1 / y^a > 0)\).
    But by Definition \ref{4.3.11}, \((1 / x^a = x^{-a} = x^n) \land (1 / y^a = y^{-a} = y^n)\).
    And by Definition \ref{4.2.8}, \(x^n = y^n \implies y^n \geq x^n\).
    So \(x \geq y > 0 \implies y^n \geq x^n > 0\) when \(n\) is a negative integer.
\end{enumerate}
From all cases above, we conclude that \(x \geq y > 0 \implies y^n \geq x^n > 0\) when \(n\) is a negative integer.
\end{proof}

\begin{proof}{(c)}
Suppose for sake of contradiction that \(x \neq y\)
Then by Proposition \ref{4.2.9}, exactly one of the following two statements is true:
\begin{enumerate}[label=(\Roman*)]
    \item \(x > y\).
    Then by Lemma \ref{4.1.5} and the given conditions, exactly one of the following two statements is true:
    \begin{enumerate}[label=(\roman*)]
        \item \(n\) is a positive integer.
        But by Proposition \ref{4.3.10}, \(x^n > y^n\), a contradiction.
        \item \(n\) is a negative integer.
        Let \(n = -a\), where \(a\) is a negative integer.
        By Definition \ref{4.2.6}, \((x > 0 \implies 1 / x > 0) \land (y > 0 \implies 1 / y > 0)\).
        By Additional Corollary \ref{ac 4.2.5} and Definition \ref{4.2.2}, \((1 / x) \times (1 / y) = 1 / xy > 0\).
        So \(x > y \implies x \times (1 / xy) > y \times (1 / xy) \implies 1 / y > 1 / x\) by Proposition \ref{4.2.9}.
        By Proposition \ref{4.3.10}, \((1 / y)^a > (1 / x)^a\).
        But by Definition \ref{4.3.11}, Proposition \ref{4.3.12}(a) and Additional Corollary \ref{ac 4.2.3}, \((1 / y)^a = (y^{-1})^a = y^{(-1)a} = y^{-a} = y^n\) and \((1 / x)^a = (x^{-1})^a = x^{(-1)a} = x^{-a} = x^n\).
        So \(x > y \implies y^n > x^n\), a contradiction.
    \end{enumerate}
    \item \(x < y\).
    By Proposition \ref{4.2.9}, \(x < y \implies y > x\), which just the same as \(x > y\).
\end{enumerate}
From all cases above we get a contradiction, so \(x = y\).
\end{proof}

\begin{proof}{(d)}
By Lemma \ref{4.1.5}, exactly one of the following three statements is true:
\begin{enumerate}[label=(\Roman*)]
    \item \(n = 0\).
    Then by Proposition \ref{4.3.10}, \(|x^0| = |x|^0\).
    \item \(n\) is a positive integer.
    Then by Proposition \ref{4.3.10}, \(|x^n| = |x|^n\).
    \item \(n\) is a negative integer.
    Then by Definition \ref{4.3.11}, \(|x^n| = |1 / x^{-n}|\) and \(|x|^n = 1 / |x|^{-n}\).
    By Lemma \ref{4.2.7}, exactly one of the following three statements is true:
    \begin{enumerate}[label=(\roman*)]
        \item \(x^{-n} = 0\).
        This case does not exist because \(x \neq 0 \implies x^{-n} \neq 0\).
        \item \(x^{-n}\) is a positive rational number.
        Then by Definition \ref{4.2.6} and Definition \ref{4.3.1}, \(1 / x^{-n}\) is a positive rational number and \(|1 / x^{-n}| = 1 / x^{-n} = 1 / |x^{-n}|\).
        And by Proposition \ref{4.3.10}, \(1 / |x|^{-n} = 1 / |x^{-n}|\).
        So \(|x^n| = |x|^n\).
        \item \(x^{-n}\) is a negative rational number.
        By Additional Corollary \ref{ac 4.2.3}, \(1 / x^{-n}\) is a negative rational number.
        By Definition \ref{4.3.1}, \(|1 / x^{-n}| = -(1 / x^{-n})\).
        Again by Additional Corollary \ref{ac 4.2.3}, \(-(1 / x^{-n}) = 1 / -(x^{-n})\).
        Again by Definition \ref{4.3.1}, \(1 / -(x^{-n}) = 1 / |x^{-n}|\).
        And by Proposition \ref{4.3.10}, \(1 / |x|^{-n} = 1 / |x^{-n}|\).
        So \(|x^n| = |x|^n\).
    \end{enumerate}
\end{enumerate}
From all cases above, we conclude that \(|x^n| = |x|^n\).
\end{proof}

\exercisesection

\begin{exercise}\label{ex 4.3.1}
Prove Proposition \ref{4.3.3}.
\end{exercise}

\begin{proof}
See Proposition \ref{4.3.3}.
\end{proof}

\begin{exercise}\label{ex 4.3.2}
Prove the remaining claims in Proposition \ref{4.3.7}.
\end{exercise}

\begin{proof}
See Proposition \ref{4.3.7}.
\end{proof}

\begin{exercise}\label{ex 4.3.3}
Prove Proposition \ref{4.3.10}.
\end{exercise}

\begin{proof}
See Proposition \ref{4.3.10}.
\end{proof}

\begin{exercise}
Prove Proposition \ref{4.3.12}.
\end{exercise}

\begin{proof}
See Proposition \ref{4.3.12}.
\end{proof}

\begin{exercise}\label{ex 4.3.5}
Prove that \(2^N \geq N\) for all positive integers \(N\).
\end{exercise}

\begin{proof}
We use induction on \(N\) and begin with \(N = 1\).
For \(N = 1\), \(2^1 = 2^0 \times 2 = 1 \times 2 = 2 \geq 1\) by Definition \ref{4.3.9}, so the base case holds.
Suppose inductively that for some \(N\), \(2^N \geq N\).
Then for \(N++\),
\begin{align*}
& (2N - N = N) \land (N \text{ is a positive integer}) \\
\implies & 2N > N & \text{(by Definition \ref{4.2.8})} \\
\implies & N < 2N & \text{(by Proposition \ref{4.2.9})} \\
\implies & N++ \leq 2N. & \text{(by Proposition \ref{2.2.12})} \\
& 2^N \geq N & \text{(by induction hypothesis)} \\
\implies & N \leq 2^N & \text{(by Lemma \ref{4.2.3})} \\
\implies & 2N \leq 2 \times 2^N & \text{(by Lemma \ref{4.2.3})} \\
\implies & 2N \leq 2^N \times 2 & \text{(by Proposition \ref{4.2.4})} \\
\implies & 2N \leq 2^{N++} & \text{(by Definition \ref{4.3.9})} \\
\implies & N++ \leq 2^{N++} & \text{(by Proposition \ref{4.2.9})} \\
\implies & 2^{N++} \geq N++. & \text{(by Proposition \ref{4.2.9})}
\end{align*}
This close the induction.
\end{proof}