\section{Rearrangement of series}\label{sec:7.4}

\begin{note}
  One feature of finite sums is that no matter how one rearranges the terms in a sequence, the total sum is the same.
  A more rigorous statement of this, involving bijections, has already appeared earlier, see \cref{7.1.12}.
\end{note}

\begin{prop}\label{7.4.1}
  Let \(\sum_{n = 0}^\infty a_n\) be a convergent series of non-negative real numbers, and let \(f : \N \to \N\) be a bijection.
  Then \(\sum_{m = 0}^\infty a_{f(m)}\) is also convergent, and has the same sum:
  \[
    \sum_{n = 0}^\infty a_n = \sum_{m = 0}^\infty a_{f(m)}.
  \]
\end{prop}

\begin{proof}
  We introduce the partial sums \(S_N \coloneqq \sum_{n = 0}^N a_n\) and \(T_M \coloneqq \sum_{m = 0}^M a_{f(m)}\).
  We know that the sequences \((S_N)_{N = 0}^\infty\) and \((T_M)_{M = 0}^\infty\) are increasing.
  Write \(L \coloneqq \sup(S_N)_{N = 0}^\infty\) and \(L' \coloneqq \sup(T_M)_{M = 0}^\infty\).
  By \cref{6.3.8} we know that \(L\) is finite, and in fact \(L = \sum_{n = 0}^\infty a_n\);
  by \cref{6.3.8} again we see that we will thus be done as soon as we can show that \(L' = L\).

  Fix \(M\), and let \(Y\) be the set \(Y \coloneqq \set{m \in \N : m \leq M}\).
  Note that \(f\) is a bijection between \(Y\) and \(f(Y)\).
  By \cref{7.1.11}, we have
  \[
    T_M = \sum_{m = 0}^M a_{f(m)} = \sum_{m \in Y} a_{f(m)} = \sum_{n \in f(Y)} a_n.
  \]
  The sequence \((f(m))_{m = 0}^M\) is finite, hence bounded, i.e., there exists an \(N\) such that \(f(m) \leq N\) for all \(m \leq M\).
  In particular \(f(Y)\) is a subset of \(\set{n \in \N : n \leq N}\), and so by \cref{7.1.11} again (and the assumption that all the \(a_n\) are non-negative)
  \[
    T_M = \sum_{n \in f(Y)} a_n \leq \sum_{n \in \set{n \in \N : n \leq N}} a_n = \sum_{n = 0}^N a_n = S_N.
  \]
  But since \((S_N)_{N = 0}^\infty\) has a supremum of \(L\), we thus see that \(S_N \leq L\), and hence that \(T_M \leq L\) for all \(M\).
  Since \(L'\) is the least upper bound of \((T_M)_{M = 0}^\infty\), this implies that \(L' \leq L\).

  Now we fix \(N\), and let \(X\) be the set \(X \coloneqq \set{n \in \N : n \leq N}\).
  Note that \(f^{-1}\) is a bijection between \(X\) and \(f^{-1}(X)\).
  By \cref{7.1.11}, we have
  \[
    S_N = \sum_{n = 0}^N a_n = \sum_{n \in X} a_n = \sum_{m \in f^{-1}(X)} a_{f(m)}.
  \]
  The sequence \((f^{-1}(n))_{n = 0}^N\) is finite, hence bounded, i.e., there exists an \(M\) such that \(f^{-1}(n) \leq M\) for all \(n \leq N\).
  In particular \(f^{-1}(X)\) is a subset of \(\set{m \in \N : m \leq M}\), and so by \cref{7.1.11} again (and the assumption that all the \(a_n\) are non-negative)
  \[
    S_N = \sum_{m \in f^{-1}(X)} a_{f(m)} \leq \sum_{m \in \set{m \in \N : m \leq M}} a_{f(m)} = \sum_{m = 0}^M a_{f(m)} = T_M.
  \]
  But since \((T_M)_{M = 0}^\infty\) has a supremum of \(L'\), we thus see that \(T_M \leq L'\), and hence that \(S_N \leq L'\) for all \(N\).
  Since \(L\) is the least upper bound of \((S_N)_{N = 0}^\infty\), this implies that \(L \leq L'\).

  Combining these two inequalities we obtain \(L = L'\), as desired.
\end{proof}

\setcounter{thm}{2}
\begin{prop}[Rearrangement of series]\label{7.4.3}
  Let \(\sum_{n = 0}^\infty a_n\) be an absolutely convergent series of real numbers, and let \(f : \N \to \N\) be a bijection.
  Then \(\sum_{m = 0}^\infty a_{f(m)}\) is also absolutely convergent, and has the same sum:
  \[
    \sum_{n = 0}^\infty a_n = \sum_{m = 0}^\infty a_{f(m)}.
  \]
\end{prop}

\begin{proof}
  We apply \cref{7.4.1} to the infinite series \(\sum_{n = 0}^\infty \abs{a_n}\), which by hypothesis is a convergent series of non-negative numbers.
  If we write \(L \coloneqq \sum_{n = 0}^\infty \abs{a_n}\), then by \cref{7.4.1} we know that \(\sum_{m = 0}^\infty \abs{a_{f(m)}}\) also converges to \(L\).

  Now write \(L' \coloneqq \sum_{n = 0}^\infty a_n\).
  We have to show that \(\sum_{m = 0}^\infty a_{f(m)}\) also converges to \(L'\).
  In other words, given any \(\varepsilon > 0\), we have to find an \(M\) such that \(\sum_{m = 0}^{M'} a_{f(m)}\) is \(\varepsilon\)-close to \(L'\) for every \(M' \geq M\).

  Since \(\sum_{n = 0}^\infty \abs{a_n}\) is convergent, we can use \cref{7.2.5} and find an \(N_1\) such that \(\sum_{n = p + 1}^q \abs{a_n} \leq \varepsilon / 2\) for all \(p, q \geq N_1\).
  Since \(\sum_{n = 0}^\infty a_n\) converges to \(L'\), the partial sums \(\sum_{n = 0}^N a_n\) also converge to \(L'\), and so there exists \(N \geq N_1\) such that \(\sum_{n = 0}^N a_n\) is \(\varepsilon / 2\)-close to \(L'\).

  Now the sequence \((f^{-1}(n))_{n = 0}^N\) is finite, hence bounded, so there exists an \(M\) such that \(f^{-1}(n) \leq M\) for all \(0 \leq n \leq N\).
  In particular, for any \(M' \geq M\), the set \(\set{f(m) : m \in \N; m \leq M'}\) contains \(\set{n \in \N : n \leq N}\).
  So by \cref{7.1.11}, for any \(M' \geq M\),
  \[
    \sum_{m = 0}^{M'} a_{f(m)} = \sum_{n \in \set{f(m) : m \in \N; m \leq M'}} a_n = \sum_{n = 0}^N a_n + \sum_{n \in X} a_n
  \]
  where \(X\) is the set
  \[
    X = \set{f(m) : m \in \N; m \leq M'} \setminus \set{n \in \N : n \leq N}.
  \]
  The set \(X\) is finite, and is therefore bounded by some natural number \(q\);
  we must therefore have
  \[
    X \subseteq \set{n \in \N : N + 1 \leq n \leq q}.
  \]
  Thus
  \[
    \abs{\sum_{m = 0}^{M'} a_{f(m)} - \sum_{n = 0}^N a_n} = \abs{\sum_{n \in X} a_n} \leq \sum_{n \in X} \abs{a_n} \leq \sum_{n = N + 1}^q \abs{a_n} \leq \varepsilon / 2
  \]
  by our choice of \(N\).
  Thus \(\sum_{m = 0}^{M'} a_{f(m)}\) is \(\varepsilon / 2\)-close to \(\sum_{n = 0}^N a_n\), which as mentioned before is \(\varepsilon / 2\)-close to \(L'\).
  Thus \(\sum_{m = 0}^{M'} a_{f(m)}\) is \(\varepsilon\)-close to \(L'\) for all \(M' \geq M\), as desired.
\end{proof}

\begin{note}
  There is in fact a surprising result of Riemann, which shows that a series which is conditionally convergent but not absolutely convergent can in fact be rearranged to converge to \emph{any} value
  (or rearranged to diverge).
\end{note}

\begin{note}
  To summarize, rearranging series is safe when the series is absolutely convergent, but is somewhat dangerous otherwise.
  (This is not to say that rearranging a series that is not absolutely convergent necessarily gives you the wrong answer
  - for instance, in theoretical physics one often performs similar maneuvres, and one still (usually) obtains a correct answer at the end
  - but doing so is risky, unless it is backed by a rigorous result such as \cref{7.4.3}.)
\end{note}

\exercisesection

\begin{ex}\label{ex:7.4.1}
  Let \(\sum_{n = 0}^\infty a_n\) be an absolutely convergent series of real numbers.
  Let \(f : \N \to \N\) be an increasing function (i.e., \(f(n + 1) > f(n)\) for all \(n \in \N\)).
  Show that \(\sum_{n = 0}^\infty a_{f(n)}\) is also an absolutely convergent series.
  What happens if we assume \(f\) is merely one-to-one, rather than increasing?
\end{ex}

\begin{proof}
  Since \(f\) is bijective implies \(f\) is one-to-one, we only need to proof the case for \(f\) being one-to-one.

  Let \(S_N = \sum_{n = 0}^N \abs{a_n}\) and \(T_N = \sum_{n = 0}^N \abs{a_{f(n)}}\).
  Since \(\sum_{n = 0}^\infty a_n\) is absolutely convergent and \(S_N\) is an increasing sequence, by \cref{6.3.8} we have \(\lim_{N \to \infty} S_N = \sup(S_N)_{N = 0}^\infty\).
  Since \((f(n))_{n = 0}^N\) is a finite sequence with \(N + 1\) unique elements (\(f\) is assumed to be one-to-one), it is bounded by some \(M \in \N\), thus \(\set{f(n) : n \in \N \land n \leq N} \subseteq \set{n \in \N : n \leq M}\).
  Now we have
  \[
    T_N = \sum_{n = 0}^N \abs{a_{f(n)}} \leq \sum_{n \in \N : n \leq M} \abs{a_n} = S_M \leq \sup(S_M)_{M = 0}^\infty = \lim_{M \to \infty} S_M,
  \]
  which means \(T_N\) is bounded.
  Since \((T_N)_{N = 0}^\infty\) is an increasing sequence and is bounded, by \cref{6.3.8} \((T_N)_{N = 0}^\infty\) converges, and thus \(\sum_{n = 0}^\infty a_{f(n)}\) is absolutely convergent.
\end{proof}

\begin{ex}\label{ex:7.4.2}
  Obtain an alternate proof of \cref{7.4.3} using \cref{7.4.1}, \cref{7.2.14}, and expressing \(a_n\) as the difference of \(a_n + \abs{a_n}\) and \(\abs{a_n}\).
  (This proof is due to Will Ballard.)
\end{ex}

\begin{proof}
  From hypothesis we know that \(\sum_{n = 0}^\infty \abs{a_n}\) converges, thus by \cref{7.2.9} we know that \(\sum_{n = 0}^\infty a_n\) converges.
  Since \(\sum_{n = 0}^\infty \abs{a_n}\) converges, by \cref{7.2.14}(b) we know that \(\sum_{n = 0}^\infty 2 \abs{a_n}\) converges and
  \begin{align*}
             & \forall n \geq 0, -\abs{a_n} \leq a_n \leq \abs{a_n}                             \\
    \implies & 0 \leq a_n + \abs{a_n} \leq 2\abs{a_n}                                           \\
    \implies & \sum_{n = 0}^\infty \big(a_n + \abs{a_n}\big) \text{ converges}. &  & \by{7.3.2}
  \end{align*}
  Now we write \(a_n = a_n + \abs{a_n} - \abs{a_n}\).
  Since \(0 \leq a_n + \abs{a_n}\), we have
  \begin{align*}
             & \sum_{n = 0}^\infty (a_n + \abs{a_n}) \text{ and } \sum_{n = 0}^\infty \abs{a_n} \text{ converges}                                               \\
    \implies & \Bigg(\sum_{n = 0}^\infty (a_n + \abs{a_n}) = \sum_{m = 0}^\infty \big(a_{f(m)} + \abs{a_{f(m)}}\big)\Bigg)                                      \\
             & \land \Bigg(\sum_{n = 0}^\infty \abs{a_n} = \sum_{m = 0}^\infty \abs{a_{f(m)}}\Bigg)                        &  & \by{7.4.1}                      \\
    \implies & \sum_{n = 0}^\infty \big(a_n + \abs{a_n}\big) - \sum_{n = 0}^\infty \abs{a_n}                                                                    \\
             & = \sum_{m = 0}^\infty \big(a_{f(m)} + \abs{a_{f(m)}}\big) - \sum_{m = 0}^\infty \abs{a_{f(m)}}                                                   \\
    \implies & \sum_{n = 0}^\infty a_n = \sum_{m = 0}^\infty a_{f(m)}.                                                     &  & \text{(by \cref{7.2.14}(a)(b))}
  \end{align*}
\end{proof}