\section{Cauchy sequences and complete metric spaces}\label{sec 1.4}

\begin{definition}[Subsequences]\label{1.4.1}
    Suppose that \((x^{(n)})_{n = m}^\infty\) is a sequence of points in a metric space \((X, d)\).
    Suppose that \(n_1, n_2, n_3, \dots\) is an increasing sequence of integers which are at least as large as \(m\), thus
    \[
        m \leq n_1 < n_2 < n_3 < \dots.
    \]
    Then we call the sequence \((x^{(n_j)})_{j = 1}^\infty\) a \emph{subsequence} of the original sequence \((x^{(n)})_{n = m}^\infty\).
\end{definition}

\setcounter{theorem}{2}
\begin{lemma}\label{1.4.3}
    Let \((x^{(n)})_{n = m}^\infty\) be a sequence in \((X, d)\) which converges to some limit \(x_0\).
    Then every subsequence \((x^{(n_j)})_{j = 1}^\infty\) of that sequence also converges to \(x_0\).
\end{lemma}

\begin{proof}
    \begin{align*}
                 & \lim_{n \to \infty} d(x^{(n)}, x_0) = 0                                                                                                                                                     \\
        \implies & \forall\ \varepsilon \in \mathbf{R}^+, \exists\ N \in \mathbf{N} : \forall\ n \geq N, d(x^{(n)}, x_0) \leq \varepsilon                                & \text{(by Definition \ref{1.1.14})} \\
        \implies & \forall\ \varepsilon \in \mathbf{R}^+, \exists\ j \in \mathbf{N} : \forall\ j \geq N, (n_j \geq j) \land \big(d(x^{(n_j)}, x_0) \leq \varepsilon\big)                                       \\
        \implies & \lim_{j \to \infty} d(x^{(n_j)}, x_0) = 0.                                                                                                            & \text{(by Definition \ref{1.1.14})}
    \end{align*}
\end{proof}

\begin{definition}[Limit points]\label{1.4.4}
    Suppose that \((x^{(n)})_{n = m}^\infty\) is a sequence of points in a metric space \((X, d)\), and let \(L \in X\).
    We say that \(L\) is a \emph{limit point} of \((x^{(n)})_{n = m}^\infty\) iff for every \(N \geq m\) and \(\varepsilon > 0\) there exists an \(n \geq N\) such that \(d(x^{(n)}, L) \leq \varepsilon\).
\end{definition}

\begin{proposition}\label{1.4.5}
    Let \((x^{(n)})_{n = m}^\infty\) be a sequence of points in a metric space \((X, d)\), and let \(L \in X\).
    Then the following are equivalent:
    \begin{itemize}
        \item \(L\) is a limit point of \((x^{(n)})_{n = m}^\infty\).
        \item There exists a subsequence \((x^{(n_j)})_{j = 1}^\infty\) of the original sequence \((x^{(n)})_{n = m}^\infty\) which converges to \(L\).
    \end{itemize}
\end{proposition}

\begin{proof}
    We first show that if \(L\) is a limit point of \((x^{(n)})_{n = m}^\infty\) in \((X, d)\), then there exists a subsequence of \((x^{(n)})_{n = m}^\infty\) which converges to \(L\) with respect to \(d\).
    Let \(N \in \mathbf{N}\).
    Since \(L\) is a limit point of \((x^{(n)})_{n = m}^\infty\) in \((X, d)\), by Definition \ref{1.4.4} we know that
    \[
        \forall\ \varepsilon \in \mathbf{R}^+, \forall\ N \geq m, \exists\ n \geq N : d(x^{(n)}, L) \leq \varepsilon.
    \]
    In particular, for every \(j \in \mathbf{Z}^+\), we have
    \[
        \forall\ N \geq m, \exists\ n \geq N : d(x^{(n)}, L) \leq \frac{1}{j}.
    \]
    Now we define \(n_j\) as follow:
    \[
        n_j = \begin{cases}
            \min\big\{n \in \mathbf{Z}^+ : d(x^{(n)}, L) \leq 1\big\}                                           & \text{if } j = 1    \\
            \min\Big\{n \in \mathbf{Z}^+ : (n > n_{j - 1}) \land \big(d(x^{(n)}, L) \leq \frac{1}{j}\big)\Big\} & \text{if } j \neq 1
        \end{cases}
    \]
    By well-ordering theorem \(n_j\) is well-defined for every \(j \in \mathbf{Z}^+\).
    By applying squeeze test to the subsequence \((x^{(n_j)})_{j = 1}^\infty\) we have
    \begin{align*}
                 & 0 \leq d(x^{(n_j)}, L) \leq \frac{1}{j}                                                                     \\
        \implies & 0 = \lim_{j \to \infty} 0 \leq \lim_{j \to \infty} d(x^{(n_j)}, L) \leq \lim_{j \to \infty} \frac{1}{j} = 0 \\
        \implies & \lim_{j \to \infty} d(x^{(n_j)}, L) = 0
    \end{align*}
    and thus by Definition \ref{1.1.14} the sequence \((x^{(n_j)})_{j = 1}^\infty\) converges to \(L\) with respect to \(d\).

    Now we show that if a subsequence of \((x^{(n)})_{n = m}^\infty\) converges to \(L\) with respect to \(d\), then \(L\) is a limit point of \((x^{(n)})_{n = m}^\infty\) in \((X, d)\).
    Let \(N \in \mathbf{N}\) and let \((x^{(n_j)})_{j = 1}^\infty\) be a subsequence of \((x^{(n)})_{n = 0}^\infty\) where \(\lim_{j \to \infty} d(x^{(n_j)}, L) = 0\).
    Then we have
    \begin{align*}
                 & \lim_{n \to \infty} d(x^{(n_j)}, L) = 0                                                                                                                      \\
        \implies & \forall\ \varepsilon \in \mathbf{R}^+, \exists\ j \geq 1 : d(x^{(n_j)}, L) \leq \varepsilon                            & \text{(by Definition \ref{1.1.14})} \\
        \implies & \forall\ \varepsilon \in \mathbf{R}^+,  \forall\ N \geq m, \exists\ n \geq n_j \geq N : d(x^{(n)}, L) \leq \varepsilon                                       \\
        \implies & L \text{ is a limit point of } (x^{(n)})_{n = m}^\infty \text{ in } (X, d).                                            & \text{(by Definition \ref{1.4.4})}
    \end{align*}
\end{proof}

\begin{definition}[Cauchy sequences]\label{1.4.6}
    Let \((x^{(n)})_{n = m}^\infty\) be a sequence of points in a metric space \((X, d)\).
    We say that this sequence is a \emph{Cauchy sequence} iff for every \(\varepsilon > 0\), there exists an \(N \geq m\) such that \(d(x^{(j)}, x^{(k)}) \leq \varepsilon\) for all \(j, k \geq N\).
\end{definition}

\begin{note}
    Here the book use \(<\) instead of \(\leq\), but the two inequalities are more or less the same.
    I use \(\leq\) to ensure consistency with Definition 5.1.8 in Analysis I.
\end{note}

\begin{lemma}[Convergent sequences are Cauchy sequences]\label{1.4.7}
    Let \((x^{(n)})_{n = m}^\infty\) be a sequence in \((X, d)\) which converges to some limit \(x_0\).
    Then \((x^{(n)})_{n = m}^\infty\) is also a Cauchy sequence.
\end{lemma}

\begin{proof}
    Let \(N \in \mathbf{N}\).
    Since \(\lim_{n \to \infty} d(x^{(n)}, x_0) = 0\), by Definition \ref{1.1.14} we know that
    \[
        \forall\ \varepsilon \in \mathbf{R}^+, \exists\ N \geq m : \forall\ n \geq N, d(x^{(n)}, x_0) \leq \frac{\varepsilon}{2}.
    \]
    Let \(k \in \mathbf{N}\) and \(k \geq N\).
    Then we have
    \begin{align*}
        d(x^{(n)}, x^{(k)}) & \leq d(x^{(n)}, x_0) + d(x_0, x^{(k)})             & \text{(by Definition \ref{1.1.2}(d))} \\
                            & = d(x^{(n)}, x_0) + d(x^{(k)}, x_0)                & \text{(by Definition \ref{1.1.2}(c))} \\
                            & \leq \frac{\varepsilon}{2} + \frac{\varepsilon}{2}                                         \\
                            & = \varepsilon
    \end{align*}
    and by Definition \ref{1.4.6} \((x^{(n)})_{n = m}^\infty\) is a Cauchy sequence in \((X, d)\).
\end{proof}

\setcounter{theorem}{8}
\begin{lemma}\label{1.4.9}
    Let \((x^{(n)})_{n = m}^\infty\) be a Cauchy sequence in \((X, d)\).
    Suppose that there is some subsequence \((x^{(n_j)})_{j = 1}^\infty\) of this sequence which converges to a limit \(x_0\) in \(X\).
    Then the original sequence \((x^{(n)})_{n = m}^\infty\) also converges to \(x_0\).
\end{lemma}

\begin{proof}
    Let \(N_1, N_2, i, k \in \mathbf{N}\).
    Since \(\lim_{j \to \infty} d(x^{(n_j)}, x_0) = 0\), by Definition \ref{1.1.14} we know that
    \[
        \forall\ \varepsilon \in \mathbf{R}^+, \exists\ N_1 \geq 1 : \forall\ j \geq N_1, d(x^{(n_j)}, x_0) \leq \frac{\varepsilon}{2}.
    \]
    Now fix such \(\varepsilon\).
    Since \((x^{(n)})_{n = m}^\infty\) is a Cauchy sequence in \((X, d)\), by Definition \ref{1.4.6} we know that
    \[
        \exists\ N_2 \geq m : \forall\ i, k \geq N_2, d(x^{(i)}, x^{(k)}) \leq \frac{\varepsilon}{2}.
    \]
    Let \(N = \max(N_1, N_2)\).
    Then \(\forall\ i, j \geq N\), we have
    \begin{align*}
        d(x^{(i)}, x_0) & \leq d(x^{(i)}, x^{(n_j)}) + d(x^{(n_j)}, x_0)     & \text{(by Definition \ref{1.1.2}(d))} \\
                        & \leq \frac{\varepsilon}{2} + \frac{\varepsilon}{2} & (n_j \geq j \geq N)                   \\
                        & = \varepsilon.
    \end{align*}
    Since \(\varepsilon\) is arbitrary, by Definition \ref{1.1.14} \(\lim_{n \to \infty} d(x^{(n)}, x_0) = 0\).
\end{proof}

\begin{definition}[Complete metric spaces]\label{1.4.10}
    A metric space \((X, d)\) is said to be \emph{complete} iff every Cauchy sequence in \((X, d)\) is in fact convergent in \((X, d)\).
\end{definition}

\setcounter{theorem}{11}
\begin{proposition}\label{1.4.12}
    \quad
    \begin{enumerate}
        \item Let \((X, d)\) be a metric space, and let \((Y, d|_{Y \times Y})\) be a subspace of \((X, d)\).
              If \((Y, d|_{Y \times Y})\) is complete, then \(Y\) must be closed in \(X\).
        \item Conversely, suppose that \((X, d)\) is a complete metric space, and \(Y\) is a closed subset of \(X\).
              Then the subspace \((Y, d|_{Y \times Y})\) is also complete.
    \end{enumerate}
\end{proposition}

\begin{proof}
    We first show that the statement (a) is true.
    Suppose that \((X, d)\) is a metric space and \((Y, d|_{Y \times Y})\) is a complete subspace of \((X, d)\).
    Let \(x_0 \in \partial Y\).
    By Proposition \ref{1.2.10} \(\exists\ (x^{(n)})_{n = m}^\infty\) in \(Y\) such that \(\lim_{n \to \infty} d(x^{(n)}, x_0) = 0\).
    Since \((Y, d|_{Y \times Y})\) is complete, by Definition \ref{1.4.10} we know that \(x_0 \in Y\).
    Thus we have \(\partial Y \subseteq Y\) and by Definition \ref{1.2.12} \(Y\) is closed.

    Now we show that the statement (b) is true.
    Suppose that \((X, d)\) is a complete metric space and \(Y\) is a closed subset of \(X\).
    Let \((x^{(n)})_{n = m}^\infty\) be a Cauchy sequence in \(Y\).
    Since \(Y \subseteq X\), we know that \((x^{(n)})_{n = m}^\infty\) is also in \(X\).
    Since \((X, d)\) is complete, we know that \(\lim_{n \to \infty} d(x^{(n)}, x_0) = 0\) for some \(x_0 \in X\).
    Since \(\lim_{n \to \infty} d(x^{(n)}, x_0) = 0\), we know that \(x_0\) is an adherent point of \(Y\).
    But \(Y\) is closed, thus by Proposition \ref{1.2.15}(b) we know that \(x_0 \in Y\), and by Definition \ref{1.4.10} \((Y, d|_{Y \times Y})\) is complete.
\end{proof}

\exercisesection

\begin{exercise}\label{ex 1.4.1}
    Prove Lemma \ref{1.4.3}.
\end{exercise}

\begin{proof}
    See Lemma \ref{1.4.3}.
\end{proof}

\begin{exercise}\label{ex 1.4.2}
    Prove Proposition 1.4.5.
\end{exercise}

\begin{proof}
    See Proposition \ref{1.4.5}.
\end{proof}

\begin{exercise}\label{ex 1.4.3}
    Prove Lemma \ref{1.4.7}.
\end{exercise}

\begin{proof}
    See Lemma \ref{1.4.7}.
\end{proof}

\begin{exercise}\label{ex 1.4.4}
    Prove Lemma \ref{1.4.9}.
\end{exercise}

\begin{proof}
    See Lemma \ref{1.4.9}.
\end{proof}

\begin{exercise}\label{ex 1.4.5}
    Let \((x^{(n)})_{n = m}^\infty\) be a sequence of points in a metric space \((X, d)\), and let \(L \in X\).
    Show that if \(L\) is a limit point of the sequence \((x^{(n)})_{n = m}^\infty\), then \(L\) is an adherent point of the set \(\{x^{(n)} : n \geq m\}\).
    Is the converse true?
\end{exercise}

\begin{proof}
    We first show that if \(L\) is a limit point of \((x^{(n)})_{n = m}^\infty\), then \(L\) is an adherent point of \(\{x^{(n)} : n \geq m\}\).
    Suppose that \(L\) is a limit point of \((x^{(n)})_{n = m}^\infty\).
    Let \(E = \{x^{(n)} : n \geq m\}\).
    By Proposition \ref{1.4.5} we know that \(\exists\ (x^{(n_j)})_{j = 1}^\infty\) in \(E\) such that \(\lim_{j \to \infty} d(x^{(n_j)}, L) = 0\).
    Thus by Proposition \ref{1.2.10} \(L\) is an adherent point of \(E\).

    Now we show that if \(L\) is an adherent point of \(\{x^{(n)} : n \geq m\}\), then \(L\) may not be a limit point of \((x^{(n)})_{n = m}^\infty\).
    Let \((X, d) = (\mathbf{R}, d_{l^1})\), let \(x^{(n)} = 1 / n\) and let \(E = \{x^{(n)} : n \geq m\}\).
    Then by Definition \ref{1.2.9} we know that \(1\) is an adherent point of \(E\).
    But by Definition \ref{1.4.4} we know that \(1\) is not an limit point of \((x^{(n)})_{n = m}^\infty\).
    Thus if \(L\) is an adherent point of \(\{x^{(n)} : n \geq m\}\), then \(L\) may not be a limit point of \((x^{(n)})_{n = m}^\infty\).
\end{proof}

\begin{exercise}\label{ex 1.4.6}
    Show that every Cauchy sequence can have at most one limit point.
\end{exercise}

\begin{proof}
    Suppose for sake of contradiction that there exists a Cauchy sequence \((x^{(n)})_{n = m}^\infty\) in some metric space \((X, d)\) which has two limit points \(L\) and \(L'\).
    Then by Lemma \ref{1.4.5} \(\exists\ (x^{(n_i)})_{i = 1}^\infty, (x^{(n_j)})_{j = 1}^\infty,\) which converges to \(L\) and \(L'\) respectively.
    Since \((x^{(n)})_{n = m}^\infty\) is a Cauchy sequence, by Lemma \ref{1.4.9} we know that \((x^{(n)})_{n = m}^\infty\) converges to \(L\) and \(L'\), which contradict to Proposition \ref{1.1.20}.
    Thus every Cauchy sequence can have at most one limit point.
\end{proof}

\begin{exercise}\label{ex 1.4.7}
    Prove Proposition \ref{1.4.12}.
\end{exercise}

\begin{proof}
    See Proposition \ref{1.4.12}.
\end{proof}

\begin{exercise}\label{ex 1.4.8}
    The following construction generalizes the construction of the reals from the rationals in Chapter 5, allowing one to view any metric space as a subspace of a complete metric space.
    In what follows we let \((X, d)\) be a metric space.
    \begin{enumerate}
        \item Given any Cauchy sequence \((x^{(n)})_{n = m}^\infty\) in \(X\), we introduce the \emph{formal limit} \\
              \(\text{LIM}_{n \to \infty} x_n\).
              We say that two formal limits \(\text{LIM}_{n \to \infty} x_n\) and \(\text{LIM}_{n \to \infty} y_n\) are equal if \(\text{lim}_{n \to \infty} d(x_n, y_n)\) is equal to zero.
              Show that this equality relation obeys the reflexive, symmetry, and transitive axioms.
        \item Let \(\overline{X}\) be the space of all formal limits of Cauchy sequences in \(X\), with the above equality relation.
              Define a metric \(d_{\overline{X}} : \overline{X} \times \overline{X} \to [0, \infty)\) by setting
              \[
                  d_{\overline{X}}(\text{LIM}_{n \to \infty} x_n, \text{LIM}_{n \to \infty} y_n) \coloneqq \lim_{n \to \infty} d(x_n, y_n).
              \]
              Show that this function is well-defined (this means not only that the limit \\
              \(\lim_{n \to \infty} d(x_n, y_n)\) exists, but also that the axiom of substitution is obeyed;
              cf. Lemma 5.3.7), and gives \(\overline{X}\) the structure of a metric space.
        \item Show that the metric space \((\overline{X}, d_{\overline{X}})\) is complete.
        \item We identify an element \(x \in X\) with the corresponding formal limit \(\text{LIM}_{n \to \infty} x\) in \(X\);
              show that this is legitimate by verifying that \(x = y \iff \text{LIM}_{n \to \infty} x = \text{LIM}_{n \to \infty} y\).
              With this identification, show that \(d(x, y) = d_{\overline{X}}(x, y)\), and thus \((X, d)\) can now be thought of as a subspace of \((\overline{X}, d_{\overline{X}})\).
        \item Show that the closure of \(X\) in \(\overline{X}\) is \(\overline{X}\) (which explains the choice of notation \(\overline{X}\)).
        \item Show that the formal limit agrees with the actual limit, thus if \((x_n)_{n = 1}^\infty\) is any Cauchy sequence in \(X\), then we have \(\lim_{n \to \infty} x_n = \text{LIM}_{n \to \infty} x_n\) in \(\overline{X}\).
    \end{enumerate}
\end{exercise}

\begin{proof}{(a)}
    Let \((x^{(n)})_{n = m}^\infty\), \((y^{(n)})_{n = m}^\infty\), \((z^{(n)})_{n = m}^\infty\) be Cauchy sequences in \((X, d)\).

    First suppose that \(\text{LIM}_{n \to \infty} x^{(n)}\) is well-defined.
    Then we have
    \begin{align*}
                 & \lim_{n \to \infty} d(x^{(n)}, x^{(n)}) = \lim_{n \to \infty} 0 = 0   & \text{(by Definition \ref{1.1.2}(a))} \\
        \implies & \text{LIM}_{n \to \infty} x^{(n)} = \text{LIM}_{n \to \infty} x^{(n)} & \text{(by definition)}
    \end{align*}
    and thus the equality relation of Exercise \ref{ex 1.4.8}(a) is reflexive.

    Next suppose that \(\text{LIM}_{n \to \infty} x^{(n)}, \text{LIM}_{n \to \infty} y^{(n)}\) are well-defined and \(\text{LIM}_{n \to \infty} x^{(n)} = \text{LIM}_{n \to \infty} y^{(n)}\).
    Then we have
    \begin{align*}
             & \text{LIM}_{n \to \infty} x^{(n)} = \text{LIM}_{n \to \infty} y^{(n)}                                         \\
        \iff & \lim_{n \to \infty} d(x^{(n)}, y^{(n)}) = 0                           & \text{(by definition)}                \\
        \iff & \lim_{n \to \infty} d(y^{(n)}, x^{(n)}) = 0                           & \text{(by Definition \ref{1.1.2}(c))} \\
        \iff & \text{LIM}_{n \to \infty} y^{(n)} = \text{LIM}_{n \to \infty} x^{(n)} & \text{(by definition)}
    \end{align*}
    and thus the equality relation of Exercise \ref{ex 1.4.8}(a) is symmetry.

    Finally suppose that \(\text{LIM}_{n \to \infty} x^{(n)}, \text{LIM}_{n \to \infty} y^{(n)}, \text{LIM}_{n \to \infty} z^{(n)}\) are well-defined.
    Suppose also that \(\text{LIM}_{n \to \infty} x^{(n)} = \text{LIM}_{n \to \infty} y^{(n)}\) and \(\text{LIM}_{n \to \infty} y^{(n)} = \text{LIM}_{n \to \infty} z^{(n)}\).
    Then we have
    \begin{align*}
                 & \begin{cases}
            \text{LIM}_{n \to \infty} x^{(n)} = \text{LIM}_{n \to \infty} y^{(n)} \\
            \text{LIM}_{n \to \infty} y^{(n)} = \text{LIM}_{n \to \infty} z^{(n)}
        \end{cases}                                                                                                  \\
        \implies & \begin{cases}
            \lim_{n \to \infty} d(x^{(n)}, y^{(n)}) = 0 \\
            \lim_{n \to \infty} d(y^{(n)}, z^{(n)}) = 0
        \end{cases}                                                             & \text{(by definition)}             \\
        \implies & \lim_{n \to \infty} \big(d(x^{(n)}, y^{(n)}) + d(y^{(n)}, z^{(n)})\big) = 0                                                 \\
        \implies & 0 \leq \lim_{n \to \infty} d(x^{(n)}, z^{(n)})                                                                              \\
                 & \quad \leq \lim_{n \to \infty} \big(d(x^{(n)}, y^{(n)}) + d(y^{(n)}, z^{(n)})\big) = 0 & \text{(by Definition \ref{1.1.2})} \\
        \implies & \lim_{n \to \infty} d(x^{(n)}, z^{(n)}) = 0                                            & \text{(by squeeze test)}           \\
        \implies & \text{LIM}_{n \to \infty} x^{(n)} = \text{LIM}_{n \to \infty} z^{(n)}                  & \text{(by definition)}
    \end{align*}
    and thus the equality relation of Exercise \ref{ex 1.4.8}(a) is transitive.
\end{proof}

\begin{proof}{(b)}
    Let \((x^{(n)})_{n = m}^\infty\), \((y^{(n)})_{n = m}^\infty\), \((z^{(n)})_{n = m}^\infty\) be Cauchy sequences in \((X, d)\) with formal limits in \(\overline{X}\).
    We first show that the limit
    \[
        d_{\overline{X}}(\text{LIM}_{n \to \infty} x^{(n)}, \text{LIM}_{n \to \infty} y^{(n)}) \coloneqq \lim_{n \to \infty} d(x^{(n)}, y^{(n)})
    \]
    exists.
    Let \((a^{(n)})_{n = m}^\infty\) be the sequence \(a^{(n)} = d(x^{(n)}, y^{(n)})\).
    To show that the above limit exists, it will suffice to show that \((a^{(n)})_{n = m}^\infty\) converges in \(\mathbf{R}\).
    Let \(N_1, N_2, j, k \in \mathbf{N}\).
    Since \((x^{(n)})_{n = m}^\infty\) is a Cauchy sequence, by Definition \ref{1.4.6} we know that
    \[
        \forall\ \varepsilon \in \mathbf{R}^+, \exists\ N_1 \geq m : \forall\ j, k \geq N_1, d(x^{(j)}, x^{(k)}) \leq \frac{\varepsilon}{2}.
    \]
    Similarly,
    \[
        \forall\ \varepsilon \in \mathbf{R}^+, \exists\ N_2 \geq m : \forall\ j, k \geq N_2, d(y^{(j)}, y^{(k)}) \leq \frac{\varepsilon}{2}.
    \]
    Let \(N = \max(N_1, N_2)\).
    Then by Definition \ref{1.1.2} we have
    \begin{align*}
         & \forall\ j, k \geq N, \abs*{a_j - a_k}                                                            \\
         & = \abs*{d(x^{(j)}, y^{(j)}) - d(x^{(k)}, y^{(k)})}                                                \\
         & = \abs*{d(x^{(j)}, y^{(j)}) + d(y^{(j)}, x^{(k)}) - d(y^{(j)}, x^{(k)}) - d(x^{(k)}, y^{(k)})}    \\
         & \leq \abs*{d(x^{(j)}, y^{(j)}) + d(y^{(j)}, x^{(k)}) + d(y^{(j)}, x^{(k)}) + d(x^{(k)}, y^{(k)})} \\
         & = d(x^{(j)}, y^{(j)}) + d(y^{(j)}, x^{(k)}) + d(y^{(j)}, x^{(k)}) + d(x^{(k)}, y^{(k)})           \\
         & \leq d(x^{(j)}, x^{(k)}) + d(y^{(j)}, y^{(k)})                                                    \\
         & \leq \varepsilon / 2 + \varepsilon / 2                                                            \\
         & = \varepsilon
    \end{align*}
    and thus \((a^{(n)})_{n = m}^\infty\) is a Cauchy sequence in \(\mathbf{R}\).
    Since \(\mathbf{R}\) is complete, we know that \((a^{(n)})_{n = m}^\infty\) converges in \(\mathbf{R}\).

    Next we show that \(d_{\overline{X}}\) obeys the axiom of substitution.
    Suppose that \(\text{LIM}_{n \to \infty} x^{(n)} = \text{LIM}_{n \to \infty} z^{(n)}\) and \(\lim_{n \to \infty} d(x^{(n)}, y^{(n)})\) exists.
    Then we have
    \begin{align*}
                 & \begin{cases}
            d(x^{(n)}, y^{(n)}) \leq d(x^{(n)}, z^{(n)}) + d(z^{(n)}, y^{(n)}) \\
            d(z^{(n)}, y^{(n)}) \leq d(x^{(n)}, y^{(n)}) + d(x^{(n)}, z^{(n)})
        \end{cases}                                                                          & \text{(by Definition \ref{1.1.2}(c)(d))}                    \\
        \implies & \begin{cases}
            d(x^{(n)}, y^{(n)}) - d(z^{(n)}, y^{(n)}) \leq d(x^{(n)}, z^{(n)}) \\
            d(z^{(n)}, y^{(n)}) - d(x^{(n)}, y^{(n)}) \leq d(x^{(n)}, z^{(n)})
        \end{cases}                                                                                                                                        \\
        \implies & 0 \leq \abs*{d(x^{(n)}, y^{(n)}) - d(z^{(n)}, y^{(n)})} \leq d(x^{(n)}, z^{(n)})                                                                                  \\
        \implies & 0 = \lim_{n \to \infty} 0 \leq \lim_{n \to \infty} \abs*{d(x^{(n)}, y^{(n)}) - d(z^{(n)}, y^{(n)})}                                                               \\
                 & \quad \leq \lim_{n \to \infty} d(x^{(n)}, z^{(n)}) = 0                                              & \text{(by Exercise \ref{ex 1.4.8}(a))}                      \\
        \implies & \lim_{n \to \infty} \abs*{d(x^{(n)}, y^{(n)}) - d(z^{(n)}, y^{(n)})} = 0                            & \text{(by squeeze test)}                                    \\
        \implies & \lim_{n \to \infty} \big(d(x^{(n)}, y^{(n)}) - d(z^{(n)}, y^{(n)})\big) = 0                                                                                       \\
        \implies & \lim_{n \to \infty} d(x^{(n)}, y^{(n)}) = \lim_{n \to \infty} d(z^{(n)}, y^{(n)})                   & \text{(\(\lim_{n \to \infty} d(x^{(n)}, y^{(n)})\) exists)}
    \end{align*}
    and thus \(d_{\overline{X}}(\text{LIM}_{n \to \infty} x^{(n)}, \text{LIM}_{n \to \infty} y^{(n)}) = d_{\overline{X}}(\text{LIM}_{n \to \infty} z^{(n)}, \text{LIM}_{n \to \infty} y^{(n)})\).

    Now we show that \((\overline{X}, d_{\overline{X}})\) is a metric space.
    For identify:
    By Definition \ref{1.1.2}(a) we have
    \[
        d_{\overline{X}}(\text{LIM}_{n \to \infty} x^{(n)}, \text{LIM}_{n \to \infty} x^{(n)}) = \lim_{n \to \infty} d(x^{(n)}, x^{(n)}) = 0.
    \]
    For positivity:
    If \(\text{LIM}_{n \to \infty} x^{(n)} \neq \text{LIM}_{n \to \infty} y^{(n)}\), then by Exercise \ref{ex 1.4.8}(a) we have
    \[
        d_{\overline{X}}(\text{LIM}_{n \to \infty} x^{(n)}, \text{LIM}_{n \to \infty} y^{(n)}) = \lim_{n \to \infty} d(x^{(n)}, y^{(n)}) \neq 0.
    \]
    For symmetry:
    We have
    \begin{align*}
         & d_{\overline{X}}(\text{LIM}_{n \to \infty} x^{(n)}, \text{LIM}_{n \to \infty} y^{(n)})                                            \\
         & = \lim_{n \to \infty} d(x^{(n)}, y^{(n)})                                                 & \text{(by definition)}                \\
         & = \lim_{n \to \infty} d(y^{(n)}, x^{(n)})                                                 & \text{(by Definition \ref{1.1.2}(c))} \\
         & = d_{\overline{X}}(\text{LIM}_{n \to \infty} y^{(n)}, \text{LIM}_{n \to \infty} x^{(n)}). & \text{(by definition)}
    \end{align*}
    For transitive:
    We have
    \begin{align*}
         & d_{\overline{X}}(\text{LIM}_{n \to \infty} x^{(n)}, \text{LIM}_{n \to \infty} z^{(n)})                                                                                             \\
         & = \lim_{n \to \infty} d(x^{(n)}, z^{(n)})                                                                                                                                          \\
         & \leq \lim_{n \to \infty} \big(d(x^{(n)}, y^{(n)}) + d(y^{(n)}, z^{(n)})\big)                                                                                                       \\
         & = \lim_{n \to \infty} d(x^{(n)}, y^{(n)}) + \lim_{n \to \infty} d(y^{(n)}, z^{(n)})                                                                                                \\
         & = d_{\overline{X}}(\text{LIM}_{n \to \infty} x^{(n)}, \text{LIM}_{n \to \infty} y^{(n)}) + d_{\overline{X}}(\text{LIM}_{n \to \infty} y^{(n)}, \text{LIM}_{n \to \infty} z^{(n)}).
    \end{align*}
    Thus by Definition \ref{1.1.2} \((\overline{X}, d_{\overline{X}})\) is a metric space.
\end{proof}

\begin{proof}{(c)}
    Let \(I, J, K, N, k_1, k_2, n_1, n_2 \in \mathbf{Z}^+\).
    Let \((a^{(n)})_{n = m}^\infty\) be arbitrary Cauchy sequence in \((\overline{X}, d_{\overline{X}})\).
    Since \((a^{(n)})_{n = m}^\infty\) is a Cauchy sequence in \((\overline{X}, d_{\overline{X}})\), by Definition \ref{1.4.6} we know that
    \begin{align*}
                 & \forall\ \varepsilon \in \mathbf{R}^+, \exists\ N \geq m : \forall\ n_1, n_2 \geq N,                                                                               \\
                 & d_{\overline{X}}(a^{(n_1)}, a^{(n_2)}) \leq \frac{\varepsilon}{4}                                                                                                  \\
        \implies & \forall\ \varepsilon \in \mathbf{R}^+, \exists\ N \geq m : \forall\ n_1, n_2 \geq N,                                                                               \\
                 & d_{\overline{X}}(\text{LIM}_{k \to \infty} a_k^{(n_1)}, \text{LIM}_{k \to \infty} a_k^{(n_2)}) \leq \frac{\varepsilon}{4} & \text{(by Exercise \ref{ex 1.4.8}(a))} \\
        \implies & \forall\ \varepsilon \in \mathbf{R}^+, \exists\ N \geq m : \forall\ n_1, n_2 \geq N,                                                                               \\
                 & \lim_{k \to \infty} d(a_k^{(n_1)}, a_k^{(n_2)}) \leq \frac{\varepsilon}{4}.                                               & \text{(by Exercise \ref{ex 1.4.8}(b))}
    \end{align*}
    Since the choice of \(N\) depends on \(\varepsilon\), we denote such \(N\) as \(N_\varepsilon\).
    We can use axiom of choice to fix \(N_\varepsilon\) for each \(\varepsilon \in \mathbf{R}^+\), and we rewrite the above statement as
    \[
        \forall\ \varepsilon \in \mathbf{R}^+, \forall\ n_1, n_2 \geq N_\varepsilon, \lim_{k \to \infty} d(a_k^{(n_1)}, a_k^{(n_2)}) \leq \frac{\varepsilon}{4}.
    \]
    Let \(L = \lim_{k \to \infty} d(a_k^{(n_1)}, a_k^{(n_2)})\).
    Since \(\Big(d(a_k^{(n_1)}, a_k^{(n_2)})\Big)_{k = 1}^\infty\) is a sequence in \(\mathbf{R}\) and converges to \(L\), we have
    \begin{align*}
                 & \exists\ I \geq 1 : \forall\ k \geq I, \abs*{d(a_k^{(n_1)}, a_k^{(n_2)}) - L} \leq \frac{\varepsilon}{4}                                                                                           \\
        \implies & \exists\ I \geq 1 : \forall\ k \geq I, -\frac{\varepsilon}{4} \leq d(a_k^{(n_1)}, a_k^{(n_2)}) - L \leq \frac{\varepsilon}{4}                                                                      \\
        \implies & \exists\ I \geq 1 : \forall\ k \geq I, 0 \leq d(a_k^{(n_1)}, a_k^{(n_2)}) - L + \frac{\varepsilon}{4} \leq \frac{\varepsilon}{2}                                                                   \\
        \implies & \exists\ I \geq 1 : \forall\ k \geq I, 0 \leq d(a_k^{(n_1)}, a_k^{(n_2)}) \leq d(a_k^{(n_1)}, a_k^{(n_2)}) - L + \frac{\varepsilon}{4} \leq \frac{\varepsilon}{2} & (L \leq \frac{\varepsilon}{4})
    \end{align*}
    Since such \(I\) depends on the choice of \(\varepsilon\), we denote such \(I\) as \(I_\varepsilon\).
    Again we can use axiom of choice to fix \(I_\varepsilon\) for each \(\varepsilon \in \mathbf{R}^+\), and we rewrite the above statement as
    \[
        \forall\ \varepsilon \in \mathbf{R}^+, \forall\ n_1, n_2 \geq N_\varepsilon, \forall\ k \geq I_\varepsilon, d(a_k^{(n_1)}, a_k^{(n_2)}) \leq \frac{\varepsilon}{2}.
    \]
    If we let \(M_\varepsilon = \max(N_\varepsilon, I_\varepsilon)\), then we can further reduce the statement as
    \[
        \forall\ \varepsilon \in \mathbf{R}^+, \forall\ n_1, n_2, k \geq M_\varepsilon, d(a_k^{(n_1)}, a_k^{(n_2)}) \leq \frac{\varepsilon}{2}.
    \]

    Since \(a^{(n)} \in \overline{X}\) for each \(n \geq m\), by Exercise \ref{ex 1.4.8}(b) we know that there exists a Cauchy sequence \((a_k^{(n)})_{k = 1}^\infty\) in \((X, d)\) such that \(\text{LIM}_{k \to \infty} a_k^{(n)} = a^{(n)}\) for each \(n \geq m\).
    By Definition \ref{1.4.6} we know that
    \[
        \forall\ \varepsilon \in \mathbf{R}^+, \exists\ J \geq 1 : \forall\ k_1, k_2 \geq J, d(a_{k_1}^{(n)}, a_{k_2}^{(n)}) \leq \frac{\varepsilon}{2}.
    \]
    Since such \(J\) depends on the choice of \(n\) and \(\varepsilon\), we denote such \(J\) as \(J_\varepsilon^{(n)}\).
    We can use axiom of choice to fix \(J_\varepsilon^{(n)}\) for each \(\varepsilon \in \mathbf{R}^+\) and for each \(n \geq m\), and we rewrite the above statement as
    \[
        \forall\ \varepsilon \in \mathbf{R}^+, \forall\ k_1, k_2 \geq J_\varepsilon^{(n)}, d(a_{k_1}^{(n)}, a_{k_2}^{(n)}) \leq \frac{\varepsilon}{2}.
    \]

    We now define a sequence \((b_k)_{k = 1}^\infty\) in \((X, d)\) by setting \(b_k = a_k^{(m + k - 1)}\) for all \(k \geq 1\).
    Informally, \((b_k)_{k = 1}^\infty\) is consist of diagonal elements in \(\big((a_k^{(n)})_{k = 1}^\infty\big)_{n = m}^\infty\).
    We claim that \((b_k)_{k = 1}^\infty\) is a Cauchy sequence in \((X, d)\).
    By Definition \ref{1.4.6} it suffices to show that
    \[
        \forall\ \varepsilon \in \mathbf{R}^+, \exists\ K \geq 1 : \forall\ k_1, k_2 \geq K, d(b_{k_1}, b_{k_2}) \leq \varepsilon.
    \]
    For each \(\varepsilon \in \mathbf{R}^+\), we have
    \begin{align*}
         & \forall\ k_1, k_2 \geq M_\varepsilon, d(b_{k_1}, b_{k_2})                                                                                                                        \\
         & = d\Big(a_{k_1}^{(m + k_1 - 1)}, a_{k_2}^{(m + k_2 - 1)}\Big)                                                                                                                    \\
         & \leq d\Big(a_{k_1}^{(m + k_1 - 1)}, a_{k_2}^{(m + k_1 - 1)}\Big) + d\Big(a_{k_2}^{(m + k_1 - 1)}, a_{k_2}^{(m + k_2 - 1)}\Big) & \text{(by Definition \ref{1.1.2}(d))}           \\
         & \leq d\Big(a_{k_1}^{(m + k_1 - 1)}, a_{k_2}^{(m + k_1 - 1)}\Big) + \frac{\varepsilon}{2}.                                      & \text{(by the definition of \(M_\varepsilon\))}
    \end{align*}
    By choosing \(K = \max(M_\varepsilon, J_\varepsilon^{(M_\varepsilon)})\) we have
    \begin{align*}
         & \forall\ k_1, k_2 \geq K, d(b_{k_1}, b_{k_2})                                                                                                                \\
         & \leq d\Big(a_{k_1}^{(m + k_1 - 1)}, a_{k_2}^{(m + k_1 - 1)}\Big) + \frac{\varepsilon}{2}                                                                     \\
         & \leq \frac{\varepsilon}{2} + \frac{\varepsilon}{2}                                       & \text{(by the definition of \(J_\varepsilon^{(M_\varepsilon)}\))} \\
         & = \varepsilon.
    \end{align*}
    Since \(\varepsilon\) is arbitrary, we have showed that
    \[
        \forall\ \varepsilon \in \mathbf{R}^+, \exists\ K \geq 1 : \forall\ k_1, k_2 \geq K, d(b_{k_1}, b_{k_2}) \leq \varepsilon.
    \]
    Thus by Definition \ref{1.4.6} \((b_k)_{k = 1}^\infty\) is a Cauchy sequence.

    Now we show that \((a^{(n)})_{n = m}^\infty\) converges in \((\overline{X}, d_{\overline{X}})\).
    From the proof above we know that \((b_k)_{k = 1}^\infty\) is a Cauchy sequence, so \(\text{LIM}_{k \to \infty} b_k \in \overline{X}\).
    We claim that \((a^{(n)})_{n = m}^\infty\) converges to \(\text{LIM}_{k \to \infty} b_k\) with respect to \(d_{\overline{X}}\).
    It suffices to show that
    \begin{align*}
             & \lim_{n \to \infty} d_{\overline{X}}\Big(a^{(n)}, \text{LIM}_{k \to \infty} b_k\Big) = 0                             & \text{(by Definition \ref{1.1.14})}    \\
        \iff & \lim_{n \to \infty} d_{\overline{X}}\Big(\text{LIM}_{k \to \infty} a_k^{(n)}, \text{LIM}_{k \to \infty} b_k\Big) = 0 & \text{(by Exercise \ref{ex 1.4.8}(a))} \\
        \iff & \lim_{n \to \infty} \big(\lim_{k \to \infty} d(a_k^{(n)}, b_k)\big) = 0                                              & \text{(by Exercise \ref{ex 1.4.8}(b))} \\
        \iff & \forall\ \varepsilon \in \mathbf{R}^+, \exists\ N \geq m : \forall\ n \geq m,                                                                                 \\
             & \abs*{\lim_{k \to \infty} d(a_k^{(n)}, b_k) - 0} \leq \varepsilon.
    \end{align*}
    Since
    \begin{align*}
         & \forall\ \varepsilon \in \mathbf{R}^+, \exists\ M_\varepsilon \geq 1 : \forall\ k \geq M_\varepsilon, d(a_k^{(M_\varepsilon)}, b_k)                                                   \\
         & = d\Big(a_k^{(M_\varepsilon)}, a_k^{(m + k - 1)}\Big)                                                                                                                                 \\
         & \leq \frac{\varepsilon}{2}                                                                                                          & \text{(by the definition of \(M_\varepsilon\))} \\
         & < \varepsilon,
    \end{align*}
    we know that \(\lim_{k \to \infty} d(a_k^{(M_\varepsilon)}, b_k)\) exists.
    Since
    \begin{align*}
                 & \forall\ n, k \geq M_\varepsilon, 0 \leq d(a_k^{(n)}, b_k) \leq \frac{\varepsilon}{2} < \varepsilon & \text{(by the definition of \(M_\varepsilon\))} \\
        \implies & \forall\ n \geq M_\varepsilon, 0 \leq \lim_{k \to \infty} d(a_k^{(n)}, b_k) \leq \varepsilon        & \text{(by comparison test)}                     \\
        \implies & \forall\ n \geq M_\varepsilon, \abs*{\lim_{k \to \infty} d(a_k^{(n)}, b_k) - 0} \leq \varepsilon.
    \end{align*}
    by setting \(N = M_\varepsilon\) we are done.

    Since for arbitrary Cauchy sequence \((a^{(n)})_{n = m}^\infty\) in \((\overline{X}, d_{\overline{X}})\) converges to \((\overline{X}, d_{\overline{X}})\), by Definition \ref{1.4.10} we know that \((\overline{X}, d_{\overline{X}})\) is complete.
\end{proof}

\begin{proof}{(d)}
    Since for any \(x, y \in X\), we have
    \begin{align*}
             & x = y                                                                                                                   \\
        \iff & d(x, y) = 0                                                                    & \text{(by Definition \ref{1.1.2}(a))}  \\
        \iff & \lim_{n \to \infty} d(x, y) = 0                                                                                         \\
        \iff & d_{\overline{X}}(\text{LIM}_{n \to \infty} x, \text{LIM}_{n \to \infty} y) = 0 & \text{(by Exercise \ref{ex 1.4.8}(b))} \\
        \iff & \text{LIM}_{n \to \infty} x = \text{LIM}_{n \to \infty} y.                     & \text{(by Definition \ref{1.1.2}(a))}
    \end{align*}
    Thus
    \begin{align*}
        d_{\overline{X}}(x, y) & = d_{\overline{X}}(\text{LIM}_{n \to \infty} x, \text{LIM}_{n \to \infty} y)                                          \\
                               & = \lim_{n \to \infty} d(x, y)                                                & \text{(by Exercise \ref{ex 1.4.8}(b))} \\
                               & = d(x, y).
    \end{align*}
\end{proof}

\begin{proof}{(e)}
    From Exercise \ref{ex 1.4.8}(d) have \(d = d_{\overline{X}}|_{X \times X}\).
    Let \(Y\) be the closure of \(X\) in \((\overline{X}, d_{\overline{X}})\).
    We want to show that \(Y = \overline{X}\).
    By Definition \ref{1.2.9} we know that \(Y \subseteq \overline{X}\).
    Thus we only need to show that \(\overline{X} \subseteq Y\).

    Let \(x_0 \in \overline{X}\).
    By Exercise \ref{ex 1.4.8}(b) there exists a Cauchy sequence \((a_n)_{n = 1}^\infty\) in \((X, d_{\overline{X}})\) such that \(\text{LIM}_{n \to \infty} a_n = x_0\).
    Since \((\overline{X}, d_{\overline{X}})\) is complete, by Definition \ref{1.4.10} we know that \((a_n)_{n = 1}^\infty\) converges in \(\overline{X}\) with respect to \(d_{\overline{X}}\).
    But we know that \((a_n)_{n = 1}^\infty\) is a Cauchy sequence in \((X, d_{\overline{X}})\), thus by Proposition \ref{1.2.10}(c) \(x_0\) is an adherent point of \(X\) and \(x_0 \in Y\).
    Since \(x_0\) is arbitrary, we thus have \(\overline{X} \subseteq Y\).
\end{proof}

\begin{proof}{(f)}
    By Exercise \ref{ex 1.4.8}(d) we know that \((x_n)_{n = 1}^\infty\) is also in \((\overline{X}, d_{\overline{X}})\).
    Thus by Exercise \ref{ex 1.4.8}(c)(e) we have \(\lim_{n \to \infty} x_n = \text{LIM}_{n \to \infty} x_n\).
\end{proof}