\section{Compact metric spaces}\label{sec:1.5}

\begin{defn}[Compactness]\label{1.5.1}
  A metric space \((X, d)\) is said to be \emph{compact} iff every sequence in \((X, d)\) has at least one convergent subsequence.
  A subset \(Y\) of a metric space \(X\) is said to be \emph{compact} if the subspace \((Y, d|_{Y \times Y})\) is compact.
\end{defn}

\begin{rmk}\label{1.5.2}
  The notion of a set \(Y\) being compact is \emph{intrinsic}, in the sense that it only depends on the metric function \(d|_{Y \times Y}\) restricted to \(Y\), and not on the choice of the ambient space \(X\).
  The notions of completeness in \cref{1.4.10}, and of boundedness in \cref{1.5.3}, are also intrinsic, but the notions of open and closed are not
  (see the discussion in \cref{sec:1.3}).
\end{rmk}

\begin{note}
  The notion of a set \(Y\) being compact only depends on the metric function \(d|_{Y \times Y}\) but not ambient space \(X\) since the elements of a sequence in \(Y\) stays the same no matter which spaces \(Y\) is subset to.
  But the notion of a set being open or closed depends on the definition of a metric ball, which may be different when given different ambient spaces.
\end{note}

\begin{note}
  Heine-Borel theorem shows that in the real line \(\R\) with the usual metric, every closed and bounded set is compact, and conversely every compact set is closed and bounded.
\end{note}

\begin{defn}[Bounded sets]\label{1.5.3}
  Let \((X, d)\) be a metric space, and let \(Y\) be a subset of \(X\).
  We say that \(Y\) is \emph{bounded} iff for every \(x \in X\) there exists a ball \(B(x, r)\) in \(X\) which contains \(Y\).
  We call \((X, d)\) bounded if \(X\) is bounded.
\end{defn}

\begin{rmk}\label{1.5.4}
  \cref{1.5.3} is compatible with the definition of a bounded set in \((\R, d_{l^1}|_{\R \times \R})\).
\end{rmk}

\begin{prop}\label{1.5.5}
  Let \((X, d)\) be a compact metric space.
  Then \((X, d)\) is both complete and bounded.
\end{prop}

\begin{proof}
  We first show that \((X, d)\) is complete.
  Let \((a_n)_{n = 1}^\infty\) be a Cauchy sequence in \((X, d)\).
  Since \((X, d)\) is compact, by \cref{1.5.1} we know that there exists a subsequence of \((a_n)_{n = 1}^\infty\) which converges to some \(x_0 \in X\) with respect to \(d\).
  Since \((a_n)_{n = 1}^\infty\) is a Cauchy sequence in \((X, d)\), by \cref{1.4.9} we know that \((a_n)_{n = 1}^\infty\) must converge to \(x_0\) with respect to \(d\).
  Since \((a_n)_{n = 1}^\infty\) is arbitrary, by \cref{1.4.10} we know that \((X, d)\) is complete.

  Now we show that \((X, d)\) is bounded by contradiction.
  Suppose for sake of contradiction that \((X, d)\) is not bounded.
  Then by \cref{1.5.3} we have
  \begin{align*}
             & \lnot\big(\forall x_0 \in X, \exists\ r \in \R^+ : X \subseteq B_{(X, d)}(x_0, r)\big)            \\
    \implies & \exists\ x_0 \in X : \forall r \in \R^+, X \not\subseteq B_{(X, d)}(x_0, r)                       \\
    \implies & \exists\ x_0 \in X : \forall r \in \R^+, \big(X \setminus B_{(X, d)}(x_0, r)\big) \neq \emptyset  \\
    \implies & \exists\ x_0 \in X : \forall n \in \Z^+, \big(X \setminus B_{(X, d)}(x_0, n)\big) \neq \emptyset.
  \end{align*}
  If \(X = \emptyset\), then \((\emptyset, d)\) is bound since \(x_0 \notin \emptyset\).
  So we only considered the cases \(X \neq \emptyset\).
  Let \((a_n)_{n = 1}^\infty\) be the sequence where \(a_n \in X \setminus B_{(X, d)}(x_0, n)\) for all \(n \in \Z^+\).
  Note that such sequence is well-defined by axiom of choice.
  Since \(a_n \in X \setminus B_{(X, d)}(x_0, n)\), by \cref{1.2.1} we know that \(d(a_n, x_0) \geq n\) for all \(n \in \Z^+\).
  Since \((X, d)\) is compact, by \cref{1.5.1} there exists a subsequence \((a_{n_j})_{j = 1}^\infty\) of \((a_n)_{n = 1}^\infty\) such that \((a_{n_j})_{j = 1}^\infty\) converges in \(X\).
  Let \(\lim_{j \to \infty} d(a_{n_j}, L) = 0\) for some \(L \in X\).
  By \cref{1.1.14} we know that
  \[
    \forall \varepsilon \in \R^+, \exists\ J \in \Z^+ : \forall j \geq J, d(a_{n_j}, L) \leq \varepsilon.
  \]
  In particular,
  \[
    \exists\ J \in \Z^+ : \forall j \geq J, d(a_{n_j}, L) \leq 1.
  \]
  Now we fix such \(J\) and let \(i = \max\big(J + 1, \ceil{d(L, x_0)} + 2\big)\).
  Then we have
  \begin{align*}
    d(a_{n_i}, x_0) & \leq d(a_{n_i}, L) + d(L, x_0) & \text{(by \cref{1.1.2}(d))} \\
                    & \leq 1 + d(L, x_0)             & (i > J)
  \end{align*}
  and
  \begin{align*}
    d(a_{n_i}, x_0) & \geq n_i            & \text{(by the definition of \(a_{n_i}\))} \\
                    & \geq i                                                          \\
                    & \geq d(L, x_0) + 2. & \text{(by the definition of \(i\))}
  \end{align*}
  But this means \(d(L, x_0) + 2 \leq d(L, x_0) + 1\), a contradiction.
  Thus \((X, d)\) is bounded.
\end{proof}

\begin{cor}[Compact sets are closed and bounded]\label{1.5.6}
  Let \((X, d)\) be a metric space, and let \(Y\) be a compact subset of \(X\).
  Then \(Y\) is closed and bounded.
\end{cor}

\begin{proof}
  Since \((Y, d|_{Y \times Y})\) is compact, by \cref{1.5.5} we know that \((Y, d)\) is complete and bounded.
  Thus by \cref{1.4.12}(a) we know that \(Y\) is closed in \((X, d)\).
\end{proof}

\begin{thm}[Heine-Borel theorem]\label{1.5.7}
  Let \((\R^n, d)\) be a Euclidean space with either the Euclidean metric, the taxicab metric, or the supnorm metric.
  Let \(E\) be a subset of \(\R^n\).
  Then \(E\) is compact if and only if it is closed and bounded.
\end{thm}

\begin{proof}
  We first show that for any \(E \subseteq \R^n\), \(E\) is closed and bounded in \((\R^n, d_{l^1}|_{\R^n \times \R^n})\) iff \((E, d_{l^1}|_{E \times E})\) is compact.
  By \cref{ex:1.1.7} we know that \((\R^n, d_{l^1}|_{\R^n \times \R^n})\) is a metric space, and by \cref{1.5.6} we know that if \((E, d_{l^1}|_{E \times E})\) is compact, then \(E\) is closed and bounded in \((\R^n, d_{l^1}|_{\R^n \times \R^n})\).
  So we only need to show that if \(E\) is closed and bounded in \((\R^n, d_{l^1}|_{\R^n \times \R^n})\), then \((E, d_{l^1}|_{E \times E})\) is compact.
  Suppose that \(E \subseteq \R^n\), \(E\) is closed and bounded in \((\R^n, d_{l^1}|_{\R^n \times \R^n})\).
  Since \(E \subseteq \R^n\), we know that for every \(x \in E\), \(x\) is in the form \(x = (x_1, \dots, x_n) = (x_i)_{i = 1}^n \in \R^n\).
  Let \(I_n = \{i \in \N : 1 \leq i \leq n\}\).
  For each \(i \in I_n\), let \(E_i\) be the set
  \[
    E_i = \{y \in \R | \exists\ x \in E : x_i = y\},
  \]
  i.e., \(E_i\) is the collection of \(i^{\text{th}}\) coordinate of all element \(x \in E\).
  We claim that for every \(i \in I_n\), \(E_i\) is a subset of some closed interval and thus \(E_i\) is bounded in \((\R, d_{l^1}|_{\R \times \R})\).
  This is true since
  \begin{align*}
             & E \text{ is bounded in } (\R^n, d_{l^1}|_{\R^n \times \R^n})                                                                              \\
    \implies & \forall y \in \R^n, \exists\ r \in \R^+ : E \subseteq B_{(\R^n, d_{l^1}|_{\R^n \times \R^n})}(y, r)            & \text{(by \cref{1.5.3})} \\
    \implies & \forall y \in \R^n, \exists\ r \in \R^+ : \forall x \in E, x \in B_{(\R^n, d_{l^1}|_{\R^n \times \R^n})}(y, r)                            \\
    \implies & \forall y \in \R^n, \exists\ r \in \R^+ : \forall x \in E,                                                                                \\
             & d_{l^1}|_{\R^n \times \R^n}(x, y) = \sum_{i = 1}^n \abs{x_i - y_i} < r                                         & \text{(by \cref{1.2.1})} \\
    \implies & \forall y \in \R^n, \exists\ r \in \R^+ : \forall x \in E, \forall i \in I_n,                                                             \\
             & d_{l^1}|_{\R \times \R}(x_i, y_i) = \abs{x_i - y_i} < r                                                                                   \\
    \implies & \forall y \in \R^n, \exists\ r \in \R^+ : \forall x \in E, \forall i \in I_n,                                                             \\
             & x_i \in (y_i - r, y_i + r)                                                                                                                \\
    \implies & \forall y \in \R^n, \exists\ r \in \R^+ : \forall i \in I_n,                                                                              \\
             & E_i \subseteq (y_i - r, y_i + r) \subseteq [y_i - r, y_i + r]                                                                             \\
    \implies & \forall y \in \R^n, \exists\ r \in \R^+ : \forall i \in I_n,                                                                              \\
             & E_i \subseteq B_{(\R, d_{l^1}|_{\R \times \R})}(y_i, r)                                                        & \text{(by \cref{1.2.1})} \\
    \implies & \forall i \in I_n, E_i \text{ is bounded in } (\R, d_{l^1}|_{\R \times \R}).                                   & \text{(by \cref{1.5.3})}
  \end{align*}

  Let \(P(n)\) be the statement ``If \(F \subseteq \R^n\) such that for every \(i \in I_n\), \(F_i\) is bounded in \((\R, d_{l^1}|_{\R \times \R})\) and \(F_i \subseteq C_i\) for some closed interval \(C_i \subseteq \R\), then for any sequence in \(F\) there exists a subsequence which converges in \(\R^n\) with respect to \(d_{l^1}|_{\R^n \times \R^n}\)''.
  We use induction on \(n\) to show that \(P(n)\) is true for all \(n \in \Z^+\).

  For \(n = 1\), by hypothesis we have \(F = F_1\) is bounded in \((\R, d_{l^1}|_{\R \times \R})\) and \(F_1 \subseteq C_1\) for some closed interval \(C_1 \subseteq \R\).
  By Heine-Borel theorem on real line (Theorem 9.1.24 in Analysis I) we know that for every sequence \((a^{(k)})_{k = 1}^\infty\) in \(F\), there exists a subsequence \((a^{(k_j)})_{j = 1}^\infty\) which converges in \(C_1 \subseteq \R\) with respect to \(d_{l^1}|_{\R \times \R}\).
  Thus the base case holds.

  Suppose inductively that \(P(n)\) is true for some \(n \geq 1\).
  Then we need to show that \(P(n + 1)\) is true.
  Let \(F \subseteq \R^{n + 1}\) such that for every \(i \in I_{n + 1}\), \(F_i\) is bounded in \((\R, d_{l^1}|_{\R \times \R})\) and \(F_i \subseteq C_i\) for some closed interval \(C_i \in \R\).
  Let \((a^{(k)})_{k = 1}^\infty\) be arbitrary sequence in \(F\).
  We define \((b^{(k)})_{k = 1}^\infty\) by setting \(b^{(k)} = (a_1^{(k)}, \dots, a_n^{(k)})\) for each \(k \geq 1\), i.e., \(b^{(k)}\) is the first \(n\) coordinates of \(a^{(k)}\).
  Since for all \(k \geq 1\), \(b^{(k)} \in \R^n\) and \(b_i^{(k)} \in F_i\) for all \(i \in I_n\), by induction hypothesis there exists a subsequence \((b^{(k_j)})_{j = 1}^\infty\) which converges in \(\R^n\) with respect to \(d_{l^1}|_{\R^n \times \R^n}\).
  Since \((a_{n + 1}^{(k_j)})_{j = 1}^\infty\) is in \(F_{n + 1}\) and \(F_{n + 1} \subseteq C_{n + 1}\) for some closed interval \(C_{n + 1} \subseteq \R\), by Heine-Borel theorem on real line (Theorem 9.1.24 in Analysis I) we know that there exists a subsequence \((a^{(k_{j_p})})_{p = 1}^\infty\) which converges in \(C_{n + 1}\) with respect to \(d_{l^1}|_{\R \times \R}\).
  But by \cref{1.4.3} we know that every subsequence of \((b^{(k_j)})_{j = 1}^\infty\) also converges in \(\R^n\) with respect to \(d_{l^1}|_{\R^n \times \R^n}\).
  In particular, \((b^{k_{j_p}})_{p = 1}^\infty\) converges in \(\R^n\) with respect to \(d_{l^1}|_{\R^n \times \R^n}\).
  Thus by \cref{1.1.18}(b)(d) we know that \((a^{(k_{j_p})})_{p = 1}^\infty\) converges in \(\R^{n + 1}\) with respect to \(d_{l^1}|_{\R^{n + 1} \times \R^{n + 1}}\), and this closes the induction.

  From the proof above we know that if \(E \subseteq \R^n\) such that \(E\) is closed and bounded in \((\R^n, d_{l^1}|_{\R^n \times \R^n})\), then for every \(i \in I_n\), \(E_i\) is bounded in \((\R, d_{l^1}|_{\R \times \R})\) and \(E_i \subseteq C_i\) for some closed interval \(C_i \subseteq \R\).
  Thus we know that for every sequence in \(E\) there exists a subsequence which converges in \(\R^n\) with respect to \(d_{l^1}|_{\R^n \times \R^n}\).
  Since \(E\) is closed in \((\R^n, d_{l^1}|_{\R^n \times \R^n})\), by \cref{1.2.15}(b) we know that such subsequence must converges in \(E\) with respect to \(d_{l^1}|_{\R^n \times \R^n}\).
  Thus by \cref{1.5.1} \((E, d_{l^1}|_{E \times E})\) is compact.

  Since every sequence in \(E\) has a subsequence which converges in \(E\) with respect to \(d_{l^1}|_{E \times E}\), by \cref{1.1.18} we know that such subsequence also converges with respect to \(d_{l^2}|_{E \times E}\) and \(d_{l^\infty}|_{E \times E}\).
  Thus \((E, d_{l^2}|_{E \times E})\) and \((E, d_{l^\infty}|_{E \times E})\) are compact iff \((E, d_{l^1}|_{\R^n \times \R^n})\) is compact.
\end{proof}

\begin{note}
  The Heine-Borel theorem is not true for more general metric spaces.
  However, a version of the Heine-Borel theorem is available if one is willing to replace closedness with the stronger notion of completeness, and boundedness with the stronger notion of \emph{total boundedness}.
\end{note}

\begin{note}
  One can characterize compactness topologically via \cref{1.5.8}:
  every open cover of a compact set has a finite subcover.
\end{note}

\begin{thm}\label{1.5.8}
  Let \((X, d)\) be a metric space, and let \(Y\) be a compact subset of \(X\).
  Let \((V_{\alpha})_{\alpha \in A}\) be a collection of open sets in \(X\), and suppose that
  \[
    Y \subseteq \bigcup_{\alpha \in A} V_{\alpha}.
  \]
  (i.e., the collection \((V_{\alpha})_{\alpha \in A}\) \emph{covers} \(Y\)).
  Then there exists a \emph{finite} subset \(F\) of \(A\) such that
  \[
    Y \subseteq \bigcup_{\alpha \in F} V_{\alpha}.
  \]
\end{thm}

\begin{proof}
  We assume for sake of contradiction that there does not exist any finite subset \(F\) of \(A\) for which \(Y \subseteq \bigcup_{\alpha \in F} V_{\alpha}\).

  Let \(y\) be any element of \(Y\).
  Then \(y\) must lie in at least one of the sets \(V_{\alpha}\).
  Since each \(V_{\alpha}\) is open in \((X, d)\), by \cref{1.2.15}(a) there must therefore be an \(r > 0\) such that \(B_{(X, d)}(y, r) \subseteq V_{\alpha}\).
  Now let \(r(y)\) denote the quantity
  \[
    r(y) \coloneqq \sup\big\{r \in (0, \infty) : B_{(X, d)}(y, r) \subseteq V_{\alpha} \text{ for some } \alpha \in A\big\}.
  \]
  By the above discussion, we know that \(r(y) > 0\) for all \(y \in Y\).
  Now, let \(r_0\) denote the quantity
  \[
    r_0 \coloneqq \inf\big\{r(y) : y \in Y\big\}.
  \]
  Since \(r(y) > 0\) for all \(y \in Y\), we have \(r_0 \geq 0\).
  There are three cases: \(r_0 = 0\), \(0 < r_0 < \infty\) and \(r_0 = \infty\).
  \begin{itemize}
    \item Case 1:
          \(r_0 = 0\).
          Then for every integer \(n \geq 1\), there is at least one point \(y\) in \(Y\) such that \(r(y) < 1 / n\) (otherwise the infimum cannot be \(0\)).
          We thus choose, for each \(n \geq 1\), a point \(y^{(n)}\) in \(Y\) such that \(r(y^{(n)}) < 1 / n\)
          (we can do this because of the axiom of choice, see Proposition 8.4.7 in Analysis I).
          In particular we have \(\lim_{n \to \infty} r(y^{(n)}) = 0\), by the squeeze test.
          The sequence \((y^{(n)})_{n = 1}^\infty\) is a sequence in \(Y\);
          since \((Y, d|_{Y \times Y})\) is compact, we can thus find a subsequence \((y^{(n_j)})_{j = 1}^\infty\) which converges to a point \(y_0 \in Y\) with respect to \(d|_{Y \times Y}\).

          As before, we know that there exists some \(\alpha \in A\) such that \(y_0 \in V_{\alpha}\), and hence (since \(V_{\alpha}\) is open in \((X, d)\)) there exists some \(\varepsilon > 0\) such that \(B_{(X, d)}(y_0, \varepsilon) \subseteq V_{\alpha}\).
          Since \((y^{(n_j)})_{j = 1}^\infty\) converges to \(y_0\), there must exist an \(N \geq 1\) such that \(y^{(n_j)} \in B_{(X, d)}(y_0, \varepsilon / 2)\) for all \(n_j \geq N\).
          In particular, by the triangle inequality we have \(B_{(X, d)}(y^{(n_j)}, \varepsilon / 2) \subseteq B_{(X, d)}(y_0, \varepsilon)\), and thus \(B_{(X, d)}(y^{(n_j)}, \varepsilon / 2) \subseteq V_{\alpha}\).
          By definition of \(r(y^{(n_j)})\), this implies that \(r(y^{(n_j)}) \geq \varepsilon / 2\) for all \(n_j \geq N\).
          But this contradicts the fact that \(\lim_{n \to \infty} r(y^{(n)}) = 0\).
    \item Case 2:
          \(0 < r_0 < \infty\).
          In this case we now have \(r(y) > r_0 / 2\) for all \(y \in Y\).
          This implies that for every \(y \in Y\) there exists an \(\alpha \in A\) such that \(B_{(X, d)}(y, r_0 / 2) \subseteq V_{\alpha}\) (by the definition of \(r(y)\)).

          We now construct a sequence \(y^{(1)}, y^{(2)}, \dots\) by the following recursive procedure.
          We let \(y^{(1)}\) be any point in \(Y\).
          The ball \(B_{(X, d)}(y^{(1)}, r_0 / 2)\) is contained in one of the \(V_{\alpha}\) and thus cannot cover all of \(Y\), since we would then obtain a finite cover, a contradiction.
          Thus there exists a point \(y^{(2)}\) which does not lie in \(B_{(X, d)}(y^{(1)}, r_0 / 2)\), so in particular \(d(y^{(2)}, y^{(1)}) \geq r_0 / 2\).
          Choose such a point \(y^{(2)}\).
          The set \(B_{(X, d)}(y^{(1)}, r_0 / 2) \cup B_{(X, d)}(y^{(2)}, r_0 / 2)\) cannot cover all of \(Y\), since we would then obtain two sets \(V_{\alpha_1}\) and \(V_{\alpha_2}\) which covered \(Y\), a contradiction again.
          So we can choose a point \(y^{(3)}\) which does not lie in \(B_{(X, d)}(y^{(1)}, r_0 / 2) \cup B_{(X, d)}(y^{(2)}, r_0 / 2)\), so in particular \(d(y^{(3)}, y^{(1)}) \geq r_0 / 2\) and \(d(y^{(3)}, y^{(2)}) \geq r_0 / 2\).
          Continuing in this fashion we obtain a sequence \((y^{(n)})_{n = 1}^\infty\) in \(Y\) with the property that \(d(y^{(k)}, y^{(j)}) \geq r_0 / 2\) for all \(k > j\).
          In particular the sequence \((y^{(n)})_{n = 1}^\infty\) is not a Cauchy sequence, and in fact no subsequence of \((y^{(n)})_{n = 1}^\infty\) can be a Cauchy sequence either.
          But this contradicts the assumption that \((Y, d|_{Y \times Y})\) is compact (by \cref{1.4.7}).
    \item Case 3:
          \(r_0 = \infty\).
          For this case we argue as in Case 2, but replacing the role of \(r_0 / 2\) by (say) \(1\).
  \end{itemize}
\end{proof}

\begin{note}
  It turns out that \cref{1.5.8} has a converse:
  if \(Y\) has the property that every open cover has a finite sub-cover, then it is compact.
  In fact, this property is often considered the more fundamental notion of compactness than the sequence-based one.
  (For metric spaces, the two notions, that of compactness and sequential compactness, are equivalent, but for more general \emph{topological spaces}, the two notions are slightly diﬀerent.)
\end{note}

\begin{cor}\label{1.5.9}
  Let \((X, d)\) be a metric space, and let \(K_1, K_2, K_3, \dots\) be a sequence of non-empty compact subsets of \(X\) such that
  \[
    K_1 \supseteq K_2 \supseteq K_3 \supseteq \dots.
  \]
  Then the intersection \(\bigcap_{n = 1}^\infty K_n\) is non-empty.
\end{cor}

\begin{proof}
  Since \((K_1, d|_{K_1 \times K_1})\) is compact and \(K_1 \subseteq X\), by \cref{1.5.6} we know that \(K_1\) is closed in \((X, d)\).
  Since \(K_1\) is closed in \((X, d)\) and \(K_1 \cap K_n = K_n\) for every \(n \geq 1\), by \cref{1.3.4}(b) we know that \(K_n\) is relatively closed in \((K_1, d|_{K_1 \times K_1})\).
  Let \(V_n = K_1 \setminus K_n\) for every \(n \geq 1\).
  Then for every \(n \geq 1\), we have \(V_n \subseteq K_1\) and by \cref{1.2.15}(e) \(V_n\) is open in \((K_1, d|_{K_1 \times K_1})\).

  Suppose for sake of contradiction that \(\bigcap_{n = 1}^\infty K_n = \emptyset\).
  Since
  \[
    \bigcup_{n = 1}^\infty V_n = \bigcup_{n = 1}^\infty (K_1 \setminus K_n) = K_1 \setminus \bigg(\bigcap_{n = 1}^\infty K_n\bigg) = K_1
  \]
  and \((K_1, d)\) is compact, by \cref{1.5.8} we know that there exists a finite set \(F \subseteq \Z^+\) such that
  \[
    K_1 \subseteq \bigcup_{i \in F} V_i.
  \]
  Since \(F\) is finite subset of \(\Z^+\), we know that \(\min(F)\) is well-defined.
  Then we have
  \begin{align*}
             & K_1 \subseteq \bigcup_{i \in F} V_i \subseteq \bigcup_{n = 1}^\infty V_i = K_1                                                          \\
    \implies & K_1 = \bigcup_{i \in F} V_i                                                                                                             \\
    \implies & K_1 = \bigcup_{i \in F} (K_1 \setminus K_i)                                                                                             \\
    \implies & K_1 = K_1 \setminus \bigg(\bigcap_{i \in F} K_i\bigg)                                                                                   \\
    \implies & \bigcap_{i \in F} K_i = \emptyset                                              & \text{(since \(\bigcap_{i \in F} K_i \subseteq K_1)\)} \\
    \implies & K_{\min(F)} = \emptyset.                                                       & \text{(since \(K_{\min(F)} = \bigcap_{i \in F} K_i)\)}
  \end{align*}
  But by hypothesis we know that \(K_{\min(F)} \neq \emptyset\), a contradiction.
  Thus \(\bigcap_{n = 1}^\infty K_n \neq \emptyset\).
\end{proof}

\begin{thm}\label{1.5.10}
  Let \((X, d)\) be a metric space.
  \begin{enumerate}
    \item If \(Y\) is a compact subset of \(X\), and \(Z \subseteq Y\), then \(Z\) is compact if and only if \(Z\) is closed.
    \item If \(Y_1, \dots, Y_n\) are a finite collection of compact subsets of \(X\), then their union \(Y_1 \cup \dots \cup Y_n\) is also compact.
    \item Every finite subset of \(X\) (including the empty set) is compact.
  \end{enumerate}
\end{thm}

\begin{proof}{(a)}
  By \cref{1.5.6} we know that if \((Z, d|_{Z \times Z})\) is compact then \(Z\) is closed in \((Y, d|_{Y \times Y})\).
  Now we show that if \(Z\) is closed in \((Y, d|_{Y \times Y})\) then \((Z, d|_{Z \times Z})\) is compact.
  Since \((Y, d|_{Y \times Y})\) is compact and \(Z \subseteq Y\), by \cref{1.5.1} we know that every sequence \((z^{(n)})_{n = 1}^\infty\) in \(Z\) has a convergent subsequence \((z^{(n_j)})_{j = 1}^\infty\) which converges in \(Y\) with respect to \(d|_{Y \times Y}\).
  Since \(Z\) is closed in \((Y, d|_{Y \times Y})\), by \cref{1.2.15}(b) we know that \((z^{(n_j)})_{j = 1}^\infty\) converges in \(Z\) with respect to \(d|_{Y \times Y}\).
  Since \((z^{(n)})_{n = 1}^\infty\) is arbitrary, by \cref{1.5.1} we know that \((Z, d|_{Z \times Z})\) is compact.
\end{proof}

\begin{proof}{(b)}
  We use induction on \(n\) to show that \((\bigcup_{i = 1}^n Y_i, d)\) is compact for every \(n \in \Z^+\).
  For \(n = 1\), we know that \(\bigcup_{i = 1}^1 Y_i = Y_1\) and by hypothesis \((Y_1, d)\) is compact, thus the base case holds.
  Suppose inductively that \((\bigcup_{i = 1}^n Y_i, d)\) is compact for some \(n \geq 1\).
  Then for \(n + 1\), we need to show that \((\bigcup_{i = 1}^{n + 1} Y_i, d)\) is compact.
  Let \((y^{(k)})_{k = 1}^\infty\) be a sequence in \(\bigcup_{i = 1}^{n + 1} Y_i\).
  We now split into two cases:
  \begin{itemize}
    \item If there exists a subsequence \((y^{(k_j)})_{j = 1}^\infty\) whose elements are in \(\bigcup_{i = 1}^n Y_i\), then by induction hypothesis we know that \((y^{(k_j)})_{j = 1}^\infty\) converges in \(\bigcup_{i = 1}^n Y_i\) with respect to \(d\).
          This means \((y^{(k_j)})_{j = 1}^\infty\) also converges in \(\bigcup_{i = 1}^{n + 1} Y_i\) with respect to \(d\) since \(\bigcup_{i = 1}^n Y_i \subseteq \bigcup_{i = 1}^{n + 1} Y_i\).
    \item If there does not exist a subsequence \((y^{(k_j)})_{j = 1}^\infty\) whose elements are in \(\bigcup_{i = 1}^n Y_i\), then there is only finitely many elements in \((y^{(k)})_{k = 1}^\infty\) which are in \(\bigcup_{i = 1}^n Y_i\).
          This means there exists a subsequence \((y^{(k_j)})_{j = 1}^\infty\) whose elements are in \(Y_{n + 1}\).
          By hypothesis we know that \((Y_{n + 1}, d)\) is compact, thus by \cref{1.5.1} there exists a subsequence \((y^{(k_{j_p})})_{p = 1}^\infty\) of \((y^{(k_j)})_{j = 1}^\infty\) converges in \(Y_{n + 1}\) with respect to \(d\).
          Since \(Y_{n + 1} \subseteq \bigcup_{i = 1}^{n + 1} Y_i\), we know that \((y^{(k_{j_p})})_{p = 1}^\infty\) also converges in \(\bigcup_{i = 1}^{n + 1} Y_i\) with respect to \(d\).
  \end{itemize}
  From all cases above we conclude that there exists a subsequence of \((y^{(k)})_{k = 1}^\infty\) which converges in \(\bigcup_{i = 1}^{n + 1} Y_i\) with respect to \(d\).
  Since \((y^{(k)})_{k = 1}^\infty\) is arbitrary, by \cref{1.5.1} \((\bigcup_{i = 1}^{n + 1} Y_i, d)\) is compact.
  This closes the induction.
\end{proof}

\begin{proof}{(c)}
  Let \(Y \subseteq X\) and \(\#(Y) = n\).
  Let \(P(n)\) be the statement ``\(\#(Y) = n\) and for every sequence \((y^{(k)})_{k = 1}^\infty\) in \(Y\), there exists a subsequence of \((y^{(k)})_{k = 1}^\infty\) which converges in \(Y\) with respect to \(d\)''.
  We use induction on \(n\) to show that \(P(n)\) is true for all \(n \in \N\).
  For \(n = 0\), we have \(Y = \emptyset\) and the statement \(P(0)\) is trivially true.
  Thus by \cref{1.5.1} \((\emptyset, d)\) is compact and the base case holds.
  Suppose inductively that \(P(n)\) is true for some \(n \geq 0\).
  Then we need to show that \(P(n + 1)\) is true.
  Let \(Y \subseteq X\) such that \(\#(Y) = n + 1\) and let \(x_0 \in Y\).
  Let \((y^{(k)})_{k = 1}^\infty\) be arbitrary sequence in \(Y\).
  Now we split into two cases:
  \begin{itemize}
    \item If the set \(\{k \in \N : y^{(k)} = x_0\}\) is finite, then we can have a subsequence \((y^{(k_j)})_{j = 1}^\infty\) whose elements are in \(Y \setminus \{x_0\}\).
          Since \((y^{(k_j)})_{j = 1}^\infty\) is in \(Y \setminus \{x_0\}\) and \(\#(Y \setminus \{x_0\}) = n\), by induction hypothesis we know that there exists a subsequence of \((y^{(k_j)})_{j = 1}^\infty\) which converges in \(Y \setminus \{x_0\}\) with respect to \(d\).
          But \((y^{(k_j)})_{j = 1}^\infty\) is also in \(Y\), thus we know that there exists a subsequence of \((y^{(k_j)})_{j = 1}^\infty\) which converges in \(Y\) with respect to \(d\).
    \item If the set \(\{k \in \N : y^{(k)} = x_0\}\) is infinite, then we can have a subsequence \((y^{(k_j)})_{j = 1}^\infty\) whose elements are all \(x_0\) and obviously \((y^{(k_j)})_{j = 1}^\infty\) converges to \(x_0\) with respect to \(d\).
          Since \(x_0 \in Y\), we know that \((y^{(k_j)})_{j = 1}^\infty\) converges in \(Y\) with respect to \(d\).
  \end{itemize}
  From all cases above we conclude that there exists a subsequence of \((y^{(k)})_{k = 1}^\infty\) which converges in \(Y\) with respect to \(d\).
  Since \((y^{(k)})_{k = 1}^\infty\) is arbitrary, we conclude that \(P(n + 1)\) is true and this closes the induction.
  Since \(P(n)\) is true for every \(n \in \N\), by \cref{1.5.1} we know that if \(Y\) is a finite subset of \(X\), then \((Y, d|_{Y \times Y})\) is compact.
\end{proof}

\exercisesection

\begin{ex}\label{ex:1.5.1}
  Show that Definitions 9.1.22 in Analysis I and \cref{1.5.3} match when talking about subsets of the real line with the standard metric.
\end{ex}

\begin{proof}
  Let \(X \subseteq \R\) and let \(d = d_{l^1}|_{\R \times \R}\).
  By \cref{ex:1.1.2} we know that \((\R, d)\) is a metric space.
  Then we have
  \begin{align*}
         & X \text{ is bounded in the sense of Definition 9.1.22 in Analysis I}                \\
    \iff & \exists\ M \in \R^+ : X \subseteq [-M, M] \subseteq (-M - 1, M + 1)                 \\
    \iff & \exists\ M \in \R^+ : \forall x \in X, \abs{x - 0} < M + 1                          \\
    \iff & \forall y \in \R, \exists\ M \in \R^+ : \forall x \in X,                            \\
         & \abs{x - y} \leq \abs{x - 0} + \abs{y - 0} < M + 1 + \abs{y}                        \\
    \iff & \forall y \in \R, \exists\ M \in \R^+ : X \subseteq B_{(\R, d)}(y, M + 1 + \abs{y}) \\
    \iff & X \text{ is bounded in the sense of \cref{1.5.3}}.
  \end{align*}
\end{proof}

\begin{ex}\label{ex:1.5.2}
  Prove \cref{1.5.5}.
\end{ex}

\begin{proof}
  See \cref{1.5.5}.
\end{proof}

\begin{ex}\label{ex:1.5.3}
  Prove \cref{1.5.7}.
\end{ex}

\begin{proof}
  See \cref{1.5.7}.
\end{proof}

\begin{ex}\label{ex:1.5.4}
  Let \((\R, d)\) be the real line with the standard metric.
  Give an example of a continuous function \(f : \R \to \R\), and an open set \(V \subseteq \R\), such that the image \(f(V) \coloneqq \{f(x) : x \in V\}\) of \(V\) is \emph{not} open.
\end{ex}

\begin{proof}
  Let \(f(x) = 1\) for all \(x \in \R\).
  Since \(f\) is a constant function, we know that \(f\) is continuous.
  By \cref{1.2.1} and \cref{1.2.15}(c) we know that
  \[
    B_{(\R, d_{l^1}|_{\R \times \R})}(1, 1) = (0, 2)
  \]
  is open in \((\R, d_{l^1}|_{\R \times \R})\) and by \cref{1.2.15}(d)
  \[
    f\big((0, 2)\big) = \{1\}
  \]
  is closed in \((\R, d_{l^1}|_{\R \times \R})\).
  By \cref{1.2.15}(a) we know that \(\{1\}\) is not open in \((\R, d_{l^1}|_{\R \times \R})\) since we cannot find an \(r \in \R^+\) such that \(B_{(\R, d_{l^1})}(1, r) \subseteq \{1\}\).
  Thus \(f\) satisfys the requirements.
\end{proof}

\begin{ex}\label{ex:1.5.5}
  Let \((\R, d)\) be the real line with the standard metric.
  Give an example of a continuous function \(f : \R \to \R\), and a closed set \(F \subseteq \R\), such that \(f(F)\) is \emph{not} closed.
\end{ex}

\begin{proof}
  Let \(f(x) = 2^x\) for all \(x \in \R\).
  We know that \(2^x\) is continuous on \((-\infty, \infty)\).
  Let \(F = (-\infty, 0]\).
  If \(x_0 \in \partial_{(\R, d)}(F)\), then by \cref{1.2.5} we must have \(B(x_0, r) \not\subseteq F\) and \(B(x_0, r) \cap F \neq \emptyset\) for every \(r \in \R^+\).
  Thus we must have \(\partial_{(\R, d)}(F) = \{0\}\).
  Since \(0 \in F\), by \cref{1.2.12} we know that \(F\) is closed in \((\R, d)\).
  Since
  \[
    f(F) = f\big((-\infty, 0]\big) = (0, 1]
  \]
  and \(0 \notin (0, 1]\), by \cref{1.2.12} we know that \(f(F)\) is not closed in \((\R, d)\).
  Thus the function \(f\) and the closed set \(F\) satisfy the requirements.
\end{proof}

\begin{ex}\label{ex:1.5.6}
  Prove \cref{1.5.9}.
\end{ex}

\begin{proof}
  See \cref{1.5.9}.
\end{proof}

\begin{ex}\label{ex:1.5.7}
  Prove \cref{1.5.10}.
\end{ex}

\begin{proof}
  See \cref{1.5.10}.
\end{proof}

\begin{ex}\label{ex:1.5.8}
  Let \((X, d_{l^1})\) be the metric space from \cref{ex:1.1.15}.
  For each natural number \(n\), let \(e^{(n)} = (e_j^{(n)})_{j = 0}^\infty\) be the sequence in \(X\) such that \(e_j^{(n)} \coloneqq 1\) when \(n = j\) and \(e_j^{(n)} \coloneqq 0\) when \(n \neq j\).
  Show that the set \(\{e^{(n)} : n \in \N\}\) is a closed and bounded subset of \(X\), but is not compact.
  (This is despite the fact that \((X, d_{l^1})\) is even a complete metric space
  - a fact which we will not prove here.
  The problem is that not that \(X\) is incomplete, but rather that it is ``infinite-dimensional'', in a sense that we will not discuss here.)
\end{ex}

\begin{proof}
  Let \(E = \{e^{(n)} : n \in \N\}\).
  We first show that \(E\) is bounded in \((X, d_{l^1})\).
  Since
  \[
    \sum_{j = 0}^\infty \abs{e_j^{(n)}} = 1
  \]
  for every \(n \in \N\), we know that \(e^{(n)}\) is absolutely convergent and by \cref{ex:1.1.15} \(e^{(n)} \in X\).
  By \cref{ex:1.1.15} we know that
  \[
    \forall (a_j)_{j = 0}^\infty, (b_j)_{j = 0}^\infty \in X, d_{l^1}\big((a_j)_{j = 0}^\infty, (b_j)_{j = 0}^\infty\big)
  \]
  is well-defined, thus \(d_{l^1}\big((a_j)_{j = 0}^\infty, (e_j^{(n)})_{j = 0}^\infty\big)\) is well-defined for every \(n \in \N\).
  Since
  \begin{align*}
     & \forall (a_j)_{j = 0}^\infty \in X, d_{l^1}\big((a_j)_{j = 0}^\infty, (e_j^{(n)})_{j = 0}^\infty\big)                                                              \\
     & = \sum_{j = 0}^\infty \abs{a_j - e_j^{(n)}}                                                                                                                        \\
     & = \sum_{j = 0 : j \neq n}^\infty \abs{a_j} + \abs{a_n - 1}                                                                                                         \\
     & \leq \sum_{j = 0}^\infty \abs{a_j} + \abs{a_n - 1}                                                    & \text{(well-defined since \((a_j)_{j = 0}^\infty \in X\))} \\
     & \leq \sum_{j = 0}^\infty \abs{a_j} + \sup_{n \in \N}\abs{a_n - 1}                                     & \text{(well-defined since \((a_j)_{j = 0}^\infty \in X\))} \\
     & < \sum_{j = 0}^\infty \abs{a_j} + \sup_{n \in \N}\abs{a_n - 1} + 1,
  \end{align*}
  by \cref{1.2.1} we know that the ball
  \[
    B_{(X, d_{l^1})}\bigg((a_j)_{j = 0}^\infty, \sum_{j = 0}^\infty \abs{a_j} + \sup_{n \in \N}\abs{a_n - 1} + 1\bigg)
  \]
  contains the set \(E\) for every \((a_j)_{j = 0}^\infty \in X\).
  Thus by \cref{1.5.3} we know that \(E\) is bounded in \((X, d_{l^1})\).

  Next we show that \(E\) is closed in \((X, d_{l^1})\).
  Let \(\overline{E}_{(X, d_{l^1})}\) be the closure of \(E\) and let \(x \in \overline{E}_{(X, d_{l^1})}\).
  By \cref{1.2.10}(c) we know that there exists a sequence \((a^{(k)})_{k = 0}^\infty\) in \(E\) such that \(\lim_{k \to \infty} d_{l^1}\big((a_j^{(k)})_{j = 0}^\infty, x\big) = 0\).
  By \cref{1.4.7} we know that \((a^{(k)})_{k = 0}^\infty\) is a Cauchy sequence in \((X, d_{l^1})\).
  Let \(k, k' \in \N\).
  By \cref{1.4.6} we have
  \[
    \forall \varepsilon \in \R^+, \exists\ N \in \N : \forall k, k' \geq N, d_{l^1}(a^{(k)}, a^{(k')}) \leq \varepsilon.
  \]
  In particular,
  \[
    \exists\ N \in \N : \forall k, k' \geq N, d_{l^1}(a^{(k)}, a^{(k')}) \leq \frac{1}{2} < 1.
  \]
  Since \(a^{(k)}, a^{(k')} \in E\), we know that \(a^{(k)} = e^{(n)}\) and \(a^{(k')} = e^{(n')}\) for some \(n, n' \in \N\) and
  \[
    d_{l^1}(a^{(k)}, a^{(k')}) = d_{l^1}(e^{(n)}, e^{(n')}) = \begin{cases}
      0 & \text{if } n = n';    \\
      2 & \text{if } n \neq n'.
    \end{cases}
  \]
  This means
  \begin{align*}
             & \exists\ N \in \N : \forall k, k' \geq N, d_{l^1}(a^{(k)}, a^{(k')}) \leq \frac{1}{2} < 1                             \\
    \implies & \exists\ N \in \N : \forall k, k' \geq N, d_{l^1}(a^{(k)}, a^{(k')}) = 0                                              \\
    \implies & \exists\ N \in \N : \forall k \geq N, d_{l^1}(a^{(k)}, a^{(N)}) = 0                                                   \\
    \implies & \exists\ N \in \N : \lim_{k \to \infty} d_{l^1}(a^{(k)}, a^{(N)}) = 0                     & \text{(by \cref{1.1.14})} \\
    \implies & \exists\ N \in \N : a^{(N)} = x                                                           & \text{(by \cref{1.1.20})} \\
    \implies & x \in E.
  \end{align*}
  Since \(x\) is arbitrary adherent point of \(E\) in \((X, d_{l^1})\), we have
  \begin{align*}
             & \overline{E}_{(X, d_{l^1})} \subseteq E                                \\
    \implies & \overline{E}_{(X, d_{l^1})} = E         & \text{(by \cref{1.2.10}(c))} \\
    \implies & E \text{ is closed in } (X, d_{l^1}).   & \text{(by \cref{1.2.15}(b))}
  \end{align*}

  Finally we show that \((E, d_{l^1}|_{E \times E})\) is not compact.
  Let \((e^{(n)})_{n = 0}^\infty\) be a sequence and let \(n, n' \in \N\).
  Since \((e^{(n)})_{n = 0}^\infty\) is a sequence in \(E\), we know that
  \[
    \forall N \in \N, \forall n, n' \geq N, d_{l^1}(e^{(n)}, e^{(n')}) = 2.
  \]
  Thus by \cref{1.4.6} \((e^{(n)})_{n = 0}^\infty\) is not a Cauchy sequence in \((E, d_{l^1}|_{E \times E})\).
  Similarly any subsequence of \((e^{(n)})_{n = 0}^\infty\) is not a Cauchy sequence in \((E, d_{l^1}|_{E \times E})\).
  By \cref{1.4.7} this means no subsequence is convergent in \((E, d_{l^1}|_{E \times E})\), and by \cref{1.5.1} \((E, d_{l^1}|_{E \times E})\) is not compact.
\end{proof}

\begin{ex}\label{ex:1.5.9}
  Show that a metric space \((X, d)\) is compact if and only if every sequence in \(X\) has at least one limit point.
\end{ex}

\begin{proof}
  \begin{align*}
         & (X, d) \text{ is compact}                                                                    \\
    \iff & \text{every sequence in } X \text{ has a convergent subsequence}                             \\
         & \text{which converges in } X                                      & \text{(by \cref{1.5.1})} \\
    \iff & \text{every sequence in } X \text{ has at least one limit point}. & \text{(by \cref{1.4.5})}
  \end{align*}
\end{proof}

\begin{ex}\label{ex:1.5.10}
  A metric space \((X, d)\) is called \emph{totally bounded} if for every \(\varepsilon > 0\), there exists a natural number \(n\) and a finite number of balls \(B(x^{(1)}, \varepsilon), \dots, B(x^{(n)}, \varepsilon)\) which cover \(X\) (i.e., \(X = \bigcup_{i = 1}^n B(x^{(i)}, \varepsilon)\)).
  (Note that \(x^{(1)}, \dots, x^{(n)} \in X\))
  \begin{enumerate}
    \item Show that every totally bounded space is bounded.
    \item Show the following stronger version of \cref{1.5.5}:
          if \((X, d)\) is compact, then complete and totally bounded.
    \item Conversely, show that if \(X\) is complete and totally bounded, then \(X\) is compact.
  \end{enumerate}
\end{ex}

\begin{proof}{(a)}
  Suppose that \(X\) is totally bounded in \((X, d)\).
  Let \(i, j, n \in \N\).
  Then by definition we have
  \[
    \forall \varepsilon \in \R^+, \exists\ n \in \N : X = \bigcup_{i = 1}^n B_{(X, d)}(x^{(i)}, \varepsilon).
  \]
  In particular, we have
  \[
    \exists\ n \in \N : X = \bigcup_{i = 1}^n B_{(X, d)}(x^{(i)}, 1).
  \]
  If \(n = 0\), then we have \(X = \emptyset\) and by \cref{1.5.3} \(\emptyset\) is bounded in \((\emptyset, d)\).
  So suppose that \(n > 0\).
  Now we use axiom of choice to choose one finite collection of \(x^{(1)}, \dots, x^{(n)} \in X\).
  Let \(I_n = \{i \in \N : 1 \leq i \leq n\}\) and let \(r = \max\{d(x^{(i)}, x^{(1)}) : i \in I_n\}\).
  We know that \(r\) is well-defined since \(I_n\) is finite.
  Then have
  \begin{align*}
             & \forall i \in I_n, \forall y \in B_{(X, d)}(x^{(i)}, 1), d(y, x^{(i)}) < 1    & \text{(by \cref{1.2.1})}                       \\
    \implies & \forall i \in I_n, \forall y \in B_{(X, d)}(x^{(i)}, 1),                                                                       \\
             & d(y, x^{(1)}) \leq d(y, x^{(i)}) + d(x^{(i)}, x^{(1)}) < 1 + r                & \text{(by \cref{1.1.2}(d))}                    \\
    \implies & \forall y \in \bigcup_{i = 1}^n B_{(X, d)}(x^{(i)}, 1), d(y, x^{(1)}) < 1 + r                                                  \\
    \implies & \forall y \in X, d(y, x^{(1)}) < 1 + r                                        & (X = \bigcup_{i = 1}^n B_{(X, d)}(x^{(i)}, 1)) \\
    \implies & \forall y \in X, y \in B_{(X, d)}(x^{(1)}, 1 + r)                             & \text{(by \cref{1.2.1})}                       \\
    \implies & X \subseteq B_{(X, d)}(x^{(1)}, 1 + r)                                                                                         \\
    \implies & \forall y \in X, X \subseteq B_{(X, d)}\big(y, d(y, x^{(1)}) + 1 + r\big)     & \text{(by \cref{1.1.2}(d))}                    \\
    \implies & X \text{ is bounded in } (X, d).                                              & \text{(by \cref{1.5.3})}
  \end{align*}
\end{proof}

\begin{proof}{(b)}
  By \cref{1.5.5} we know that if \((X, d)\) is compact then \((X, d)\) is complete.
  Thus we only need to show that \(X\) is totally bound in \((X, d)\).
  Let \(\varepsilon \in \R^+\).
  Then we have
  \begin{align*}
             & \forall y \in X, y \in B_{(X, d)}(y, \varepsilon)                                                                          & \text{(by \cref{1.2.1})}        \\
    \implies & X \subseteq \bigcup_{y \in X} B_{(X, d)}(y, \varepsilon)                                                                                                     \\
    \implies & \bigcup_{y \in X} B_{(X, d)}(y, \varepsilon) \text{ is an open cover of } X \text{ in } (X, d)                             & \text{(by \cref{1.2.15}(c)(g))} \\
    \implies & \exists\ F \subseteq X : (F \text{ is finite}) \land \bigg(X \subseteq \bigcup_{y \in F} B_{(X, d)}(y, \varepsilon)\bigg). & \text{(by \cref{1.5.8})}
  \end{align*}
  Since \(\varepsilon\) is arbitrary, by definition we know that \(X\) is totally bounded in \((X, d)\).
\end{proof}

\begin{proof}{(c)}
  Suppose that \((X, d)\) is complete and \(X\) is totally bounded in \((X, d)\).
  If \(X = \emptyset\), then we know that \((\emptyset, d)\) is compact is trivially true.
  So suppose that \(X \neq \emptyset\).
  To show that \((X, d)\) is compact, by \cref{1.5.1} we need to show that every sequence in \(X\) has a convergent subsequence which converges in \(X\) with respect to \(d\).
  So let \((x^{(n)})_{n = 1}^\infty\) be a sequence in \(X\).
  If we have
  \[
    \exists\ N \in \Z^+ : \big\{n \in \Z^+ : x^{(n)} = x^{(N)}\big\} \text{ is infinite},
  \]
  then there must exist a subsequence of \((x^{(n)})_{n = 1}^\infty\) which converges to \(x^{(N)}\) with respect to \(d\) for some \(N \in \Z^+\).
  So suppose that
  \[
    \forall N \in \Z^+, \big\{n \in \Z^+ : x^{(n)} = x^{(N)}\big\} \text{ is finite}.
  \]
  Let \(E = \{x^{(n)} : n \in \Z^+\}\).
  For each \(\varepsilon \in \R^+\), we define \(F_\varepsilon\) to be the set
  \[
    F_\varepsilon = \Bigg\{F \subseteq X : (F \text{ is finite}) \land \bigg(X \subseteq \bigcup_{y \in F} B_{(X, d)}(y, \varepsilon)\bigg)\Bigg\}.
  \]
  Since \(X\) is totally bounded in \((X, d)\), by definition we know that \(F_\varepsilon \neq \emptyset\) for every \(\varepsilon \in \R^+\).
  For arbitrary \(\varepsilon \in \R^+\) and arbitrary \(F \in F_\varepsilon\), we claim that
  \[
    \exists\ y \in F : E \cap B_{(X, d)}(y, \varepsilon) \text{ is infinite}.
  \]
  Suppose for sake of contradiction that the claim is false.
  Then we have
  \begin{align*}
             & \forall y \in F, E \cap B_{(X, d)}(y, \varepsilon) \text{ is finite}                                                    \\
    \implies & \bigcup_{y \in F} \bigg(E \cap B_{(X, d)}(y, \varepsilon)\bigg) \text{ is finite}      & \text{(since \(F\) is finite)} \\
    \implies & E \cap \bigg(\bigcup_{y \in F} B_{(X, d)}(y, \varepsilon)\bigg) \text{ is finite}                                       \\
    \implies & E = E \cap X \subseteq E \cap \bigg(\bigcup_{y \in F} B_{(X, d)}(y, \varepsilon)\bigg)
  \end{align*}
  and thus \(E\) is finite.
  But by the definition of \(E\) we know that \(E\) is infinite, a contradiction.
  Thus the claim is true.

  Using the claim above we can now define the set \(A_\varepsilon\) for each \(\varepsilon \in \R^+\).
  \[
    A_\varepsilon = \big\{y \in F : (F \in F_\varepsilon) \land (E \cap B_{(X, d)}(y, \varepsilon) \text{ is infinite})\big\}
  \]
  From the claim above we know that \(A_\varepsilon \neq \emptyset\) for every \(\varepsilon \in \R^+\).
  Now we claim that
  \[
    \forall \delta, \varepsilon \in \R^+, \delta < \varepsilon \implies A_\delta \subseteq A_\varepsilon.
  \]
  Let \(\delta, \varepsilon \in \R^+\) and \(\delta < \varepsilon\).
  Then we have
  \begin{align*}
             & \forall y \in A_\delta, E \cap B_{(X, d)}(y, \delta) \text{ is infinite}                                                            \\
    \implies & \forall y \in A_\delta, B_{(X, d)}(y, \delta) \text{ is infinite}                                & \text{(since \(E\) is infinite)} \\
    \implies & \forall y \in A_\delta, B_{(X, d)}(y, \delta) \subseteq B_{(X, d)}(y, \varepsilon)               & \text{(by \cref{1.2.1})}         \\
    \implies & \forall y \in A_\delta, E \cap B_{(X, d)}(y, \delta) \subseteq E \cap B_{(X, d)}(y, \varepsilon)                                    \\
    \implies & \forall y \in A_\delta, E \cap B_{(X, d)}(y, \varepsilon) \text{ is infinite}                                                       \\
    \implies & \forall y \in A_\delta, y \in A_\varepsilon                                                                                         \\
    \implies & A_\delta \subseteq A_\varepsilon
  \end{align*}
  and thus the claim is true.

  Now we construct a subsequence of \(x^{(n)}\).
  For each \(j \in \Z^+\) we define \(N_j\) as follow:
  \[
    N_j = \bigg\{n \in \Z^+ : x^{(n)} \in E \cap B_{(X, d)}(y, \frac{1}{j}) \text{ for some } y \in A_{\frac{1}{j}}\bigg\}.
  \]
  From the claim above we know that \(A_{\frac{1}{j}}\) is infinite for every \(j \in \Z^+\), thus \(N_j\) is infinite and we have
  \[
    \forall i, j \in \Z^+, i < j \implies \frac{1}{j} < \frac{1}{i} \implies A_{\frac{1}{j}} \subseteq A_{\frac{1}{i}} \implies N_j \subseteq N_i.
  \]
  Now we recursively define \(n_j\) for each \(j \in \Z^+\) as follow:
  \[
    n_j = \begin{cases}
      \min N_1                                & \text{if } j = 1 \\
      \min\{n \in N_{j - 1} : n > n_{j - 1}\} & \text{if } j > 1
    \end{cases}
  \]
  Since \(N_j\) is infinite for each \(j \in \Z^+\), we know that the set \(\{n \in N_{j - 1} : n > n_{j - 1}\}\) is also infinite for every \(j \geq 2\).
  If not, then the maximum element of \(N_{j - 1}\) would be \(n_{j - 1}\) and \(N_{j - 1}\) is finite, a contradiction.
  Since \(\{n \in N_j : n > n_{j - 1}\} \subseteq \Z^+\) for each \(j \in \Z^+\) and \(j \geq 2\), by well-ordering principle we know that \(\min\{n \in N_j : n > n_{j - 1}\}\) is well-defined.
  Thus \(n_j\) is well-defined for each \(j \in \Z^+\) and \((x^{(n_j)})_{j = 1}^\infty\) is a subsequence of \((x^{(n)})_{n = 1}^\infty\).

  Now we claim that the subsequence \((x^{(n_j)})_{j = 1}^\infty\) converges in \(X\) with respect to \(d\).
  We have
  \begin{align*}
             & \forall \varepsilon \in \R^+, \exists\ J \in \Z^+ : \frac{1}{J} < \varepsilon                                                    & \text{(by Archimedean property)}                  \\
    \implies & \forall \varepsilon \in \R^+, \exists\ J \in \Z^+ :                                                                                                                                  \\
             & (\frac{1}{J} < \varepsilon) \land (\forall j \geq J, N_j \subseteq N_J)                                                          & \text{(from the claim above)}                     \\
    \implies & \forall \varepsilon \in \R^+, \exists\ J \in \Z^+ :                                                                                                                                  \\
             & (\frac{1}{J} < \varepsilon)                                                                                                                                                          \\
             & \land \big(\exists\ y \in A_{\frac{1}{j}} : \forall j \geq J, d(x^{(n_j)}, y) < \frac{1}{j} \leq \frac{1}{J} < \varepsilon\big). & \text{(by the definition of \(A_{\frac{1}{j}}\))}
  \end{align*}
  If we fix \(J\) for each \(\varepsilon\), then we have
  \begin{align*}
             & \forall i, j \geq 2J, \exists\ y \in A_{\frac{1}{2J}} :                                                                      \\
             & d(x^{(n_i)}, y) + d(x^{(n_j)}, y) < \frac{1}{2J} + \frac{1}{2J} = \frac{1}{J} < \varepsilon                                  \\
    \implies & \forall i, j \geq 2J, \exists\ y \in A_{\frac{1}{2J}} :                                                                      \\
             & d(x^{(n_i)}, x^{(n_j)}) \leq d(x^{(n_i)}, y) + d(x^{(n_j)}, y) < \varepsilon                & \text{(by \cref{1.1.2}(c)(d))} \\
    \implies & \forall i, j \geq 2J, d(x^{(n_i)}, x^{(n_j)}) < \varepsilon.
  \end{align*}
  Thus we conclude that
  \[
    \forall \varepsilon \in \R^+, \exists\ J \in \Z^+ : \forall i, j \geq J, d(x^{(n_i)}, d^{(n_j)}) < \varepsilon
  \]
  and by \cref{1.4.6} \((x^{(n_j)})_{j = 1}^\infty\) is a Cauchy sequence in \((X, d)\).
  Since \((X, d)\) is compatible, by \cref{1.4.10} we know that \((x^{(n_j)})_{j = 1}^\infty\) converges in \(X\) with respect to \(d\).
  Since \((x^{(n)})_{n = 1}^\infty\) is arbitrary, we know that any sequence in \(X\) has a subsequence converges in \(X\) with respect to \(d\), and by \cref{1.5.1} \((X, d)\) is compact.
\end{proof}

\begin{ex}\label{ex:1.5.11}
  Let \((X, d)\) have the property that every open cover of \(X\) has a finite subcover.
  Show that \(X\) is compact.
\end{ex}

\begin{proof}
  Suppose that \((X, d)\) is a metric space such that every open cover of \(X\) has a finite subcover.
  We want to show that \((X, d)\) is compact.
  Suppose for sake of contradiction that \((X, d)\) is not compact.
  Then by \cref{1.5.1} we know that there exists a sequence \((x^{(n)})_{n = 1}^\infty\) in \(X\) which has no convergent subsequence which converges in \(X\) with respect to \(d\).
  We know that
  \[
    \forall N \in \Z^+, \big\{n \in \Z^+ : x^{(n)} = x^{(N)}\big\} \text{ is finite}.
  \]
  If not, then we would have a subsequence which converges to \(x^{(N)}\) with respect to \(d\) for some \(N \in \Z^+\), a contradiction.
  Thus we know that \(\big\{x^{(n)} : n \in \Z^+\big\}\) is infinite.
  By \cref{ex:1.5.9} we know that \((x^{(n)})_{n = 1}^\infty\) cannot have a limit point in \(X\) with respect to \(d\).
  By \cref{1.4.4} this means
  \begin{align*}
             & \lnot\big(\exists\ y \in X : \forall \varepsilon \in \R^+, \forall N \in \Z^+, \exists\ n \geq N : d(x^{(n)}, y) \leq \varepsilon\big) \\
    \implies & \forall y \in X, \exists\ \varepsilon \in \R^+ : \exists\ N \in \Z^+ : \forall n \geq N, d(x^{(n)}, y) > \varepsilon                   \\
    \implies & \forall y \in X, \exists\ \varepsilon \in \R^+ : \big\{x^{(n)} : n \in \Z^+\big\} \cap B_{(X, d)}(y, \varepsilon) \text{ is finite}.
  \end{align*}
  Now we fix such \(\varepsilon\).
  Since
  \begin{align*}
             & \forall y \in X, B_{(X, d)}(y, \varepsilon) \text{ is open in } (X, d) & \text{(by \cref{1.2.15}(c))} \\
    \implies & X = \bigcup_{y \in X} B_{(X, d)}(y, \varepsilon)                       & \text{(by \cref{1.2.1})}
  \end{align*}
  and \(\bigcup_{y \in X} B_{(X, d)}(y, \varepsilon)\) is an open cover of \(X\) in \((X, d)\), we know that
  \[
    \exists\ F \subseteq X : (F \text{ is finite}) \land \bigg(X = \bigcup_{y \in F} B_{(X, d)}(y, \varepsilon)\bigg).
  \]
  Now we fix such \(F\).
  Then we have
  \begin{align*}
    \{x^{(n)} : n \in \Z^+\} & = \{x^{(n)} : n \in \Z^+\} \cap X                                                        \\
                             & = \{x^{(n)} : n \in \Z^+\} \cap \bigg(\bigcup_{y \in F} B_{(X, d)}(y, \varepsilon)\bigg) \\
                             & = \bigcup_{y \in F} \big(\{x^{(n)} : n \in \Z^+\} \cap B_{(X, d)}(y, \varepsilon)\big).
  \end{align*}
  But we know that \(\bigcup_{y \in F} \big(\{x^{(n)} : n \in \Z^+\} \cap B_{(X, d)}(y, \varepsilon)\big)\) is finite since \(F\) is finite and \(\{x^{(n)} : n \in \Z^+\} \cap B_{(X, d)}(y, \varepsilon)\) is finite for every \(y \in F\).
  This means \(\{x^{(n)} : n \in \Z^+\}\) is finite, a contradiction.
  Thus \((X, d)\) is compact.
\end{proof}

\begin{ex}\label{ex:1.5.12}
  Let \((X, d_{\text{disc}})\) be a metric space with the discrete metric \(d_{\text{disc}}\).
  \begin{enumerate}
    \item Show that \(X\) is always complete.
    \item When is \(X\) compact, and when is \(X\) not compact?
          Prove your claim.
  \end{enumerate}
\end{ex}

\begin{proof}{(a)}
  Let \((x^{(n)})_{n = 1}^\infty\) be a Cauchy sequence in \((X, d_{\text{disc}})\).
  Let \(i, j \in \Z^+\).
  By \cref{1.4.6} we know that
  \[
    \forall \varepsilon \in \R^+, \exists\ N \in \Z^+ : \forall i, j \geq N, d_{\text{disc}}(x^{(i)}, x^{(j)}) \leq \varepsilon.
  \]
  In particular, we have
  \[
    \exists\ N \in \Z^+ : \forall i, j \geq N, d_{\text{disc}}(x^{(i)}, x^{(j)}) \leq \frac{1}{2}.
  \]
  But by \cref{1.1.11} we know that
  \[
    d_{\text{disc}}(x^{(i)}, x^{(j)}) \leq \frac{1}{2} \iff x^{(i)} = x^{(j)}.
  \]
  Thus we have
  \[
    \forall \varepsilon \in \R^+, \exists\ N \in \Z^+ : \forall i \geq N, x^{(i)} = x^{(N)}.
  \]
  and by \cref{1.1.14} we have \(\lim_{n \to \infty} d_{\text{disc}}(x^{(n)}, x^{(N)}) = 0\).
  This means \((x^{(n)})_{n = 1}^\infty\) converges to some \(x^{(N)} \in X\) with respect to \(d_{\text{disc}}\).
  Since \((x^{(n)})_{n = 1}^\infty\) is arbitrary, by \cref{1.4.10} we know that \((X, d_{\text{disc}})\) is complete.
\end{proof}

\begin{proof}{(b)}
  We claim that \((X, d_{\text{disc}})\) is compact iff \(X\) is finite.
  By \cref{1.5.10}(c) we know that if \(X\) is finite then \((X, d_{\text{disc}})\) is compact.
  So we only need to show that if \((X, d_{\text{disc}})\) is compact then \(X\) is finite.
  Suppose that \((X, d_{\text{disc}})\) is compact.
  By \cref{1.5.8} we know that every open cover of \(X\) in \((X, d_{\text{disc}})\) has a finite subcover.
  In particular, we know that
  \[
    \exists\ F \subseteq X : (F \text{ is finite}) \land \bigg(X = \bigcup_{y \in F} B_{(X, d_{\text{disc}})}(y, \frac{1}{2})\bigg).
  \]
  Now we fix such \(F\).
  We know that
  \begin{align*}
             & \forall y \in F, B_{(X, d)}(y, \frac{1}{2}) = \{z \in X : d(z, y) < \frac{1}{2}\} & \text{(by \cref{1.2.1})}  \\
    \implies & \forall y \in F, B_{(X, d)}(y, \frac{1}{2}) = \{y\}                               & \text{(by \cref{1.1.11})} \\
    \implies & X = \bigcup_{y \in F} B_{(X, d)}(y, \frac{1}{2}) = F
  \end{align*}
  Thus \(X\) is finite.
\end{proof}

\begin{ex}\label{ex:1.5.13}
  Let \(E\) and \(F\) be two compact subsets of \(\R\) (with the standard metric \(d(x, y) = \abs{x - y}\)).
  Show that the Cartesian product \(E \times F \coloneqq \{(x, y) : x \in E, y \in F\}\) is a compact subset of \(\R^2\) (with the Euclidean metric \(d_{l^2}\)).
\end{ex}

\begin{proof}
  Since \(E \times F \subseteq \R^2\), we know that every element \(x \in E \times F\) is in the form \(x = (x_1, x_2)\) where \(x_1 \in E\) and \(x_2 \in F\).
  Let \((x^{(n)})_{n = 1}^\infty\) be a sequence in \(E \times F\).
  Then we know that \((x_1^{(n)})_{n = 1}^\infty\) is a sequence in \(E\) and \((x_2^{(n)})_{n = 1}^\infty\) is a sequence in \(F\).
  Since \((E, d_{l^1}|_{\R \times \R})\) is compact, by \cref{1.5.1} we know that there exists a subsequence \((x_1^{(n_j)})_{j = 1}^\infty\) which converges to some \(L_E \in E\) with respect to \(d_{l^1}|_{\R \times \R}\).
  Since \((x_2^{(n_j)})_{j = 1}^\infty\) is a subsequence of \((x_2^{(n)})_{n = 1}^\infty\) and \((F, d_{l^1}|_{\R \times \R})\) is compact, by \cref{1.5.1} we know that there exists a subsequence \((x_2^{(n_{j_p})})_{p = 1}^\infty\) which converges to some \(L_F \in F\) with respect to \(d_{l^1}|_{\R \times \R}\).
  Since \(\lim_{j \to \infty} d_{l^1}|_{\R \times \R}(x_1^{(n_j)}, L_E) = 0\), by \cref{1.4.9} we know that \(\lim_{p \to \infty} d_{l^1}|_{\R \times \R}(x_1^{(n_{j_p})}, L_E) = 0\).
  Thus by \cref{1.1.18}(a)(b)(d) we have
  \[
    \lim_{p \to \infty} d_{l^1}|_{\R^2 \times \R^2}\big(x^{(n_{j_p})}, (L_E, L_F)\big) = \lim_{p \to \infty} d_{l^2}|_{\R^2 \times \R^2}\big(x^{(n_{j_p})}, (L_E, L_F)\big) = 0.
  \]
  Since \((x^{(n)})_{n = 1}^\infty\) is arbitrary, by \cref{1.5.1} we know that \((E \times F, d_{l^2}|_{\R^2 \times \R^2})\) is compact.
\end{proof}

\begin{ex}\label{ex:1.5.14}
  Let \((X, d)\) be a metric space, let \(E\) be a non-empty compact subset of \(X\), and let \(x_0\) be a point in \(X\).
  Show that there exists a point \(x \in E\) such that
  \[
    d(x_0, x) = \inf\{d(x_0, y) : y \in E\},
  \]
  i.e., \(x\) is the closest point in \(E\) to \(x_0\).
\end{ex}

\begin{proof}
  Let \(R = \inf\{d(x_0, y) : y \in E\}\).
  Since \(d(x_0, y) \geq 0\) for every \(y \in E\), we know that the set \(\{d(x_0, y) : y \in E\}\) is bounded below and thus \(R\) is well-defined.
  Since \(R\) is well-defined, we can construct a sequence \((x^{(n)})_{n = 1}^\infty\) in \(E\) such that \(d(x_0, x^{(n)}) \leq R + \frac{1}{n}\) for every \(n \in \Z^+\).
  If not, then we would have
  \[
    \exists\ n \in \Z^+ : \forall y \in E, d(x_0, y) > R + \frac{1}{n}
  \]
  and thus \(\inf\{d(x_0, y) : y \in E\} \geq R + \frac{1}{n} > R\), a contradiction.
  Since \(\big(d(x_0, x^{(n)})\big)_{n = 1}^\infty\) is a sequence in \(\R\), we have
  \begin{align*}
             & \forall n \in \Z^+, R \leq d(x_0, x^{(n)}) \leq R + \frac{1}{n}                            \\
    \implies & \lim_{n \to \infty} d(x_0, x^{(n)}) = R.                        & \text{(by squeeze test)}
  \end{align*}
  Since \((E, d)\) is compact, by \cref{1.5.1} we know that there exists a subsequence \((x^{(n_j)})_{j = 1}^\infty\) such that
  \[
    \lim_{j \to \infty} d(x^{(n_j)}, x) = 0
  \]
  for some \(x \in E\).
  Then we have
  \begin{align*}
             & \lim_{n \to \infty} d(x_0, x^{(n)}) = R                                                                                                \\
    \implies & \lim_{j \to \infty} d(x_0, x^{(n_j)}) = R                                                                & \text{(by \cref{1.4.3})}    \\
    \implies & \lim_{j \to \infty} d(x_0, x^{(n_j)}) + \lim_{j \to \infty} d(x^{(n_j)}, x) = R + 0 = R                                                \\
    \implies & \lim_{j \to \infty} \big(d(x_0, x^{(n_j)}) + d(x^{(n_j)}, x)\big) = R                                                                  \\
    \implies & \lim_{j \to \infty} d(x_0, x) \leq \lim_{j \to \infty} \big(d(x_0, x^{(n_j)}) + d(x^{(n_j)}, x)\big) = R & \text{(by \cref{1.1.2}(d))} \\
    \implies & R \leq d(x_0, x) \leq R                                                                                                                \\
    \implies & d(x_0, x) = R
  \end{align*}
  and thus \(x\) is the closest point in \(E\) to \(x_0\).
\end{proof}

\begin{ex}\label{ex:1.5.15}
  Let \((X, d)\) be a compact metric space.
  Suppose that \((K_{\alpha})_{\alpha \in I}\) is a collection of closed sets in \(X\) with the property that any finite subcollection of these sets necessarily has non-empty intersection, thus \(\bigcap_{\alpha \in F} K_{\alpha} \neq \emptyset\) for all finite \(F \subseteq I\).
  (This property is known as the \emph{finite intersection property}.)
  Show that the \emph{entire} collection has non-empty intersection, thus \(\bigcap_{\alpha \in I} K_{\alpha} \neq \emptyset\).
  Show by counterexample that this statement fails if \(X\) is not compact.
\end{ex}

\begin{proof}
  Suppose for sake of contradiction that \(\bigcap_{\alpha \in I} K_\alpha = \emptyset\).
  Since
  \[
    X = X \setminus \emptyset = X \setminus \bigcap_{\alpha \in I} K_\alpha = \bigcup_{\alpha \in I} (X \setminus K_\alpha),
  \]
  we know that
  \begin{align*}
             & \forall \alpha \in I, K_\alpha \text{ is closed in } (X, d)                                                                                      \\
    \implies & \forall \alpha \in I, X \setminus K_\alpha \text{ is open in } (X, d)                                             & \text{(by \cref{1.2.15}(e))} \\
    \implies & X = \bigcup_{\alpha \in I} (X \setminus K_\alpha) \text{ is an open cover of } X \text{ in } (X, d)                                              \\
    \implies & \exists F \subseteq I : (F \text{ is finite}) \land \bigg(X = \bigcup_{\alpha \in F} (X \setminus K_\alpha)\bigg) & \text{(by \cref{1.5.8})}     \\
    \implies & \exists F \subseteq I : (F \text{ is finite}) \land \bigg(X = X \setminus \bigcap_{\alpha \in F} K_\alpha\bigg)                                  \\
    \implies & \exists F \subseteq I : (F \text{ is finite}) \land \bigg(\bigcap_{\alpha \in F} K_\alpha = \emptyset\bigg).
  \end{align*}
  But by hypothesis we know such \(F\) does not exist, a contradiction.
  Thus we must have \(\bigcap_{\alpha \in I} K_\alpha \neq \emptyset\).

  Now we show an counterexample when \(X, d\) is not compact.
  Let \(X = I = \R^+\) and let \(d = d_{l^1}|_{\R \times \R}\).
  We know that the sequence \((n)_{n = 1}^\infty\) in \(\R^+\) has no convergent subsequence with respect to \(d_{l^1}|_{\R \times \R}\), thus by \cref{1.5.1} \((\R^+, d_{l^1}|_{\R \times \R})\) is not compact.
  For each \(\varepsilon \in \R^+\), we define \(K_\varepsilon = [\varepsilon, \infty)\) (which is an interval in \(\R\)).
  Then we have
  \begin{align*}
             & \forall \varepsilon \in \R^+, \overline{K_\varepsilon}_{(\R^+, d_{l^1}|_{\R \times \R})} = \{\varepsilon\} & \text{(by \cref{1.2.5})}     \\
    \implies & \forall \varepsilon \in \R^+, K_\varepsilon \text{ is closed in } (\R^+, d_{l^1}|_{\R \times \R})          & \text{(by \cref{1.2.15}(b))}
  \end{align*}
  Let \(F \subseteq \R^+\) such that \(F \neq \emptyset\) and \(F\) is finite.
  Since \(F\) is finite, we know that \(\max(F)\) is well-defined.
  Then we have
  \begin{align*}
             & \forall i \in F, i \leq \max(F)                                                           \\
    \implies & \forall i \in F, [\max(F), \infty) \subseteq [i, \infty)                                  \\
    \implies & \forall i \in F, K_i \subseteq K_{\max(F)}                                                \\
    \implies & \bigcap_{i \in F} K_i = K_{\max(F)}.                     & \text{(since \(F\) is finite)}
  \end{align*}
  Since \(F\) is arbitrary, we know that every finite subcollection of \((K_\varepsilon)_{\varepsilon \in \R^+}\) has non-empty intersection.
  Suppose for sake of contradiction that \(\bigcap_{\varepsilon \in \R^+} K_\varepsilon \neq \emptyset\).
  Let \(r \in \bigcap_{\varepsilon \in \R^+} K_\varepsilon\).
  Then we have
  \begin{align*}
             & \forall \varepsilon \in \R^+, r \in K_\varepsilon \\
    \implies & r \in K_{2r}                                      \\
    \implies & r \in [2r, \infty)                                \\
    \implies & r = 0.
  \end{align*}
  But we know that \(0 \notin K_\varepsilon\) for every \(\varepsilon \in \R^+\).
  Thus \(0 \notin \bigcap_{\varepsilon \in \R^+} K_\varepsilon\), a contradiction.
  So we must have \(\bigcap_{\varepsilon \in \R^+} K_\varepsilon = \emptyset\).
\end{proof}