\section{Pointwise and uniform convergence}\label{sec 3.2}

\begin{definition}[Pointwise convergence]\label{3.2.1}
    Let \((f^{(n)})_{n = 1}^\infty\) be a sequence of functions from one metric space \((X, d_X)\) to another \((Y, d_Y)\), and let \(f : X \to Y\) be another function.
    We say that \emph{\((f^{(n)})_{n = 1}^\infty\) converges pointwise to \(f\) on \(X\)} if we have
    \[
        \lim_{n \to \infty} f^{(n)}(x) = f(x)
    \]
    for all \(x \in X\), i.e.,
    \[
        \lim_{n \to \infty} d_Y\big(f^{(n)}(x), f(x)\big) = 0.
    \]
    Or in other words, for every \(x\) and every \(\varepsilon > 0\) there exists \(N > 0\) such that \(d_Y\big(f^{(n)}(x), f(x)\big) < \varepsilon\) for every \(n > N\).
    We call the function \(f\) the \emph{pointwise limit} of the functions \(f^{(n)}\).
\end{definition}

\begin{remark}\label{3.2.2}
    Note that \(f^{(n)}(x)\) and \(f(x)\) are points in \(Y\), rather than functions, so we are using our prior notion of convergence in metric spaces to determine convergence of functions.
    Also note that we are not really using the fact that \((X, d_X)\) is a metric space
    (i.e., we are not using the metric \(d_X\));
    for this definition it would suffice for \(X\) to just be a plain old set with no metric structure.
    However, later on we shall want to restrict our attention to \emph{continuous} functions from \(X\) to \(Y\), and in order to do so we need a metric on \(X\) (and on \(Y\)), or at least a topological structure.
    Also when we introduce the concept of \emph{uniform convergence}, then we will definitely need a metric structure on \(X\) and \(Y\);
    there is no comparable notion for topological spaces.
\end{remark}

\begin{note}
    From Proposition \ref{1.1.20} we see that a sequence \((f^{(n)})_{n = 1}^\infty\) of functions from one metric space \((X, d_X)\) to another \((Y, d_Y)\) can have at most one pointwise limit \(f\)
    (this explains why we can refer to \(f\) as \emph{the} pointwise limit).
    However, it is of course possible for a sequence of functions to have no pointwise limit, just as a sequence of points in a metric space do not necessarily have a limit.
\end{note}

\begin{note}
    Pointwise convergence is a very natural concept, but it has a number of disadvantages:
    it does not preserve continuity, derivatives, limits, or integrals.
    The problem is that while \(f^{(n)}(x)\) converges to \(f(x)\) for each \(x\), the \emph{rate} of that convergence varies substantially with \(x\).
    To put things another way, the convergence of \(f^{(n)}\) to \(f\) is not \emph{uniform} in \(x\)
    - the \(N\) that one needs to get \(f^{(n)}(x)\) within \(\varepsilon\) of \(f\) depends on \(x\) as well as on \(\varepsilon\).
    This motivates a stronger notion of convergence.
\end{note}

\setcounter{theorem}{6}
\begin{definition}[Uniform convergence]\label{3.2.7}
    Let \((f^{(n)})_{n = 1}^\infty\) be a sequence of functions from one metric space \((X, d_X)\) to another \((Y, d_Y)\), and let \(f : X \to Y\) be another function.
    We say that \emph{\((f^{(n)})_{n = 1}^\infty\) converges uniformly to \(f\) on \(X\)} if for every \(\varepsilon > 0\) there exists \(N > 0\) such that \(d_Y\big(f^{(n)}(x), f(x)\big) < \varepsilon\) for every \(n > N\) and \(x \in X\).
    We call the function \(f\) the \emph{uniform limit} of the functions \(f^{(n)}\).
\end{definition}

\begin{remark}\label{3.2.8}
    Note that Definition \ref{3.2.7} is subtly different from the definition for pointwise convergence in Definition \ref{3.2.1}.
    In the definition of pointwise convergence, \(N\) was allowed to depend on \(x\);
    now it is not.
    The reader should compare this distinction to the distinction between continuity and uniform continuity
    (i.e., between Definition \ref{2.1.1} and Definition \ref{2.3.4}).
\end{remark}